### 2025-03-01 tutorials/getstarted.qmd
Start tutorial.

# 🚀 Get Started with LieGroups.jl

## Introduction

This tutorial introduces both a few basics of [Lie groups](https://en.wikipedia.org/wiki/Lie_group) as well as how to use [LieGroups.jl](https://juliamanifolds.github.io/LieGroups.jl/stable/). The Lie groups we consider are the rotations in the plane,
or [special orthogonal group](../groups/special_orthogonal_group.md) as well as rigid body motions,
or in other words the [special euclidean group](../groups/special_euclidean_group.md).

In a nutshell a Lie group ``𝒢`` is a [manifold](https://en.wikipedia.org/wiki/Manifold) that also is a group a group operation.
A manifold informally be described as a set that “locally looks like a Euclidean space”
and has some way to measure angles at every point, to be precise an inner product, that is allowed to (smoothly)
depend on the point we are at.
A [group](https://en.wikipedia.org/wiki/Group_(mathematics)) means that on the set there also
exists an operation, within this package usually denoted by ``∘: ``,
`LieGroups.jl` uses the manifolds defined in [`Manifolds.jl`](https://juliamanifolds.github.io/Manifolds.jl/stable/). For more details on that, see also the [introductory tutorial](https://juliamanifolds.github.io/Manifolds.jl/stable/tutorials/getstarted/) there.

For more theoretical background, see for example [HilgertNeeb:2012](@cite), especially their Chapter 9.

## Rotations on the plane ``ℝ^2`` and in ``ℝ^3``.

This first part considers rotations in the plane. These can be represented by rotation matrices
```math
R_α = \begin{pmatrix} \cos(α),& -\sin(α)\\ \sin(α) & \cos(α) \end{pmatrix},\qquad α ∈ ℝ,
```

where we already represent all possible rotations when restricting ``α`` to ``[0,2π]``.
Adding two matrices ``R_α + R_β`` does not yield a valid rotation matrix again, but we
can have these matrices in mind as being “locally like the real line”, since changing the rotation angle
slightly yields rotation matrices that are “close by” in the sense that they perform nearly the same rotation
when applied to a point ``x ∈ ℝ^2`` by computing ``R_αx```.
Matrix multiplication ``R_β = R_{α_1}*R_{α_1}`` does yield a new rotation and also fulfils all other
properties to yield a group operation. Using trigonometric identities we observe that ``β = α_1 + α_2``.

The set of rotation matrices is hence a Lie group and called the [special orthogonal group](../groups/special_orthogonal_group.md) ``\mathrm{SO}(2)``. In `LieGroups.jl` we initialise this as

```{julia}
using LieGroups, RecursiveArrayTools
G = SpecialOrthogonalGroup(2)
```

Elements ``g, h ∈ \mathrm{SO}(2)`` we generate as

```{julia}
g = [1.0 0.0; 0.0 1.0]
h = 1/sqrt(2) .* [1.0 -1.0; 1.0 1.0]
```

A first thing to do is to use [`is_point`](@ref) to verify they are valid.

```{julia}
[is_point(G,g), is_point(G,h)]
```

# TODO
* compose
* lie group exponential & logarithm
* Lie algebra

SO(3)
* hat/vee to get to vectors
* group action and apply

(towards) SE(3)
* product of two Lie groups
* semidirect product

## Technical details

This tutorial is cached. It was last run on the following package versions.

```{julia}
#| code-fold: true
#| echo: false
using Pkg
Pkg.status()
```
```{julia}
#| code-fold: true
#| echo: false
using Dates
println("This tutorial was last rendered $(now())");
```

## Literature

````{=commonmark}
```@bibliography
Pages = ["getstarted.md"]
Canonical=false
```
````

### 2025-03-01 tutorials/getstarted.qmd
Improve technical details print.

@@
 #| echo: false
+#| output: asis
 using Dates
-println("This tutorial was last rendered $(now())");
+println("This tutorial was last rendered $(Dates.format(now(), "U d, Y, H:M:S")).");
 ```

### 2025-03-02 tutorials/getstarted.qmd
Continue the tutorial.

@@
 
-In a nutshell a Lie group ``𝒢`` is a [manifold](https://en.wikipedia.org/wiki/Manifold) that also is a group a group operation.
+In a nutshell a Lie group $𝒢$ is a [manifold](https://en.wikipedia.org/wiki/Manifold) that also is a group a group operation.
 A manifold informally be described as a set that “locally looks like a Euclidean space”
@@
 A [group](https://en.wikipedia.org/wiki/Group_(mathematics)) means that on the set there also
-exists an operation, within this package usually denoted by ``∘: ``,
+exists an operation, within this package usually denoted by $∘: 𝒢 × 𝒢 → 𝒢$$,
 `LieGroups.jl` uses the manifolds defined in [`Manifolds.jl`](https://juliamanifolds.github.io/Manifolds.jl/stable/).
@@
 
-## Rotations on the plane ``ℝ^2`` and in ``ℝ^3``.
+## Rotations on the plane $ℝ^2$ and in $ℝ^3$.
 
@@
 
-where we already represent all possible rotations when restricting ``α`` to ``[0,2π]``.
-Adding two matrices ``R_α + R_β`` does not yield a valid rotation matrix again, but we
+where we already represent all possible rotations when restricting $α$ to $[0,2π]$.
+Adding two matrices $R_α + R_β$ does not yield a valid rotation matrix again, but we
 can have these matrices in mind as being “locally like the real line”, since changing the rotation angle
 slightly yields rotation matrices that are “close by” in the sense that they perform nearly the same rotation
-when applied to a point ``x ∈ ℝ^2`` by computing ``R_αx```.
-Matrix multiplication ``R_β = R_{α_1}*R_{α_1}`` does yield a new rotation and also fulfils all other
-properties to yield a group operation. Using trigonometric identities we observe that ``β = α_1 + α_2``.
+when applied to a point $x ∈ ℝ^2$ by computing $R_αx$.
+Matrix multiplication $R_β = R_{α_1}*R_{α_1}$ does yield a new rotation and also fulfils all other
+properties to yield a group operation. Using trigonometric identities we observe that $β = α_1 + α_2$.
 
-The set of rotation matrices is hence a Lie group and called the [special orthogonal group](../groups/special_orthogonal_group.md) ``\mathrm{SO}(2)``. In `LieGroups.jl` we initialise this as
+The set of rotation matrices is hence a Lie group and called the [special orthogonal group](../groups/special_orthogonal_group.md) $\mathrm{SO}(2)$. In `LieGroups.jl` we initialise this as
 
 ```{julia}
-using LieGroups, RecursiveArrayTools
-G = SpecialOrthogonalGroup(2)
+using LieGroups, LinearAlgebra, RecursiveArrayTools
+SO2 = SpecialOrthogonalGroup(2)
 ```
 
-Elements ``g, h ∈ \mathrm{SO}(2)`` we generate as
+Elements $g, h ∈ \mathrm{SO}(2)$ we generate as
 
@@
 ```{julia}
-[is_point(G,g), is_point(G,h)]
+(is_point(SO2, g), is_point(SO2, h))
 ```
 
+The already mentioned group operation as the matrix multiplication, there is also a generic
+function available, [`compose`](@ref) as well as its in-place variant [`compose!`](@ref).
+The following two yield the same matrix
+
+```{julia}
+[ compose(SO2, g, h), g*h ]
+```
+
+Furthermore a lot of of functions are “passed down” tp the manifold, which is stored within
+the [`LieGroup`](@ref). For example the dimension of the manifold, or the number of degrees of freedom
+can be accessed via the ``[`manifold_dimension`](@extref `ManifoldsBase.manifold_dimension-Tuple{AbstractManifold}`)``{=commonmark}
+
+```{julia}
+manifold_dimension(SO2)
+```
+
+## The Lie algebra, and the Lie group exponential function.
+
+For the following investigations, we consider the special orthogonal group $\mathrm{SO}(3)$,
+that is rotations in a 3-dimensional space.
+Besides one special rotation in the following code for completeness, a prominent, at first glace
+maybe a bit “dull” point on this Lie group is the identity matrix `e`.
+
+```{julia}
+SO3 = SpecialOrthogonalGroup(3)
+g = [cos(π/3) -sin(π/3) 0.0; sin(π/3) cos(π/3) 0.0; 0.0 0.0 1.0]* [cos(π/4) 0.0 -sin(π/4); 0.0 1.0 0.0; sin(π/4) 0.0 cos(π/4)]
+e = Matrix{Float64}(I,3,3)
+g
+```
+The element `g` can be seen as a rotation by $π/3$ in the x-y-plane combines with a rotation by $π/4$ in the x-z plane.
+
+The tangent space $T_e𝒢$ plays a special role and is called the [`LieAlgebra`](@ref) $𝔤$.
+
+Similar to the ``[Riemannian exponential map](@extref `Base.exp-Tuple{AbstractManifold, Any, Any}`)``{=commonmark}
+The exponential function $\exp_𝒢: 𝔤 → 𝒢$ maps Lie algebra tangent vectors $X ∈ T_e𝒢$
+to a point on the Lie group. This is implemented in [`exp(G::LieGroup, X)`](@ref). It's inverse
+is the Lie group logarithmic function [`log(G::LieGroup, g)`](@ref).
+
+```{julia}
+X = log(SO3,g)
+is_point(LieAlgebra(SO3), X)
+```
+
+```{julia}
+exp(SO3,X)
+```
+
+The term “exponential” has at least three different meanings throughout Lie groups and Riemannian manifolds.
+To be precise, we call the just introduced one “exponential function”, since it often agrees with the [exponential](https://en.wikipedia.org/wiki/Exponential_function).
+
+Taking a closed look at `X` we see
+
+```{julia}
+X
+```
+
+an example that all elements from the Lie algebra are skew symmetric matrices.
+This allows for one way to see that we have three degrees of freedom, cf
+
+```{julia}
+manifold_dimension(SO3)
+```
+
+As the Lie algebra was introduced via a tangent space, we also know that it is a vector space.
+To “turn” the `X` into a vector with three elements, we need a basis of the tangent space
+to decompose `X` into and “read off” its components.
+While in general this can be done using the [bases of a tangent space](https://juliamanifolds.github.io/ManifoldsBase.jl/stable/bases/) from [`ManifoldsBase.jl`](https://juliamanifolds.github.io/ManifoldsBase.jl/stable/),
+a very specific one is [`vee(::LieAlgebra, X)`](@ref)
+
+```{julia}
+c = vee(LieAlgebra(SO3), X)
+```
+
+and its inverse, [`hat(G::LieAlgebra, c)`](@ref)
+
+```{julia}
+X_ = hat(LieAlgebra(SO3),c)
+```
+
+## Representing tangent vectors
+
+Consider the function of left group composition $λ_g(h) = g∘h$.
+Its differential (or push forward) $Dλ_g(e)$ maps from $T_e𝒢$ to $T_g𝒢$
+and is a linear bijection between both spaces.
+Its adjoint differential (or pullback) $D^*λ_g(e)$ maps back from $T_g𝒢$ to $T_e𝒢$.
+
+This allows to introduce an inner product on all tangent spaces, that smoothly varies with
+the base point, we obtain the left-invariant metric
+
+```math
+⟨X,Y⟩_g = ⟨D^*λ_g(e)[X],D^*λ_g(e)⟩_e,
+```
+
+since it is invariant if you use the differential of the left composition to identify tangent vectors.
+We can even “spare” the pullbacks in this definition, when spring tangent vectors as their representants in the Lie algebra.
+This is the default in `LieGroups.jl`.
+
+With respect to this metric on the manifold the exponential and logarithmic maps are given by
+
+```math
+\exp_g(X) = g ∘ \exp_𝒢(X)
+\quad\text{with its inverse}\quad
+\log_g(h) = \log_𝒢(g^{-1}∘h)
+```
+
+With respect to these geodesics and the representation in the Lie algebra, parallel transport
+simplifies to the identity. To still access the Riemannian exponential map with recpect to the
+metric (compatible to the Levi-cevita connection) on the underlying Riemannian manifold, use
+`exp(base_manifold(G), g, X)`.
+
+## The Special Euclidean group $\mathrm{SE}(3)$ of rigid body motions
+
 # TODO
-* compose
-* lie group exponential & logarithm
-* Lie algebra
-
-SO(3)
-* hat/vee to get to vectors
 * group action and apply
-
-(towards) SE(3)
 * product of two Lie groups

### 2025-03-02 src/interface.jl

@@ -380,7 +380,7 @@ _doc_exponential = """
     exp(G::LieGroup, X::T)
     exp!(G::LieGroup, g, X)
 
-    Compute the (Lie group) exponential function
+Compute the (Lie group) exponential function
 
 ```math
 $(_tex(:exp))_{$(_math(:G))}: $(_math(:𝔤)) → $(_math(:G)),$(_tex(:qquad)) $(_tex(:exp))_{$(_math(:G))}(X) = γ_X(1),

### 2025-03-02 docs/make.jl
added supporting tests for only the complex circle group (exp, compose, is_identity)
added scalar_multiplication_operation.jl to improve readability

todo:
 -> add support for testing references as input
 -> further functions

many of the problems regarding the implementation of the scalar case root in a lack of functionality in ManifoldsBase.copy_to! and ManifoldsBase.allocate_result() for manifolds with an empty representation_size therefore a scalar representation. for a cleaner and more general implementation those methods need to be generalized

@@ -114,7 +114,7 @@ links = InterLinks(
 makedocs(;
     format=Documenter.HTML(;
         prettyurls=(get(ENV, "CI", nothing) == "true") || ("--prettyurls" ∈ ARGS),
-        assets=["assets/favicon.ico", "assets/citations.css", "assets/link-icons.css"],
+        assets=["assets/favicon.ico", "assets/citations.css"],
     ),
     modules=[LieGroups],
     authors="Seth Axen, Mateusz Baran, Ronny Bergmann, Olivier Verdier, and contributors",

### 2025-03-02 src/LieGroups.jl

@@ -32,6 +32,7 @@ include("Lie_algebra/Lie_algebra_interface.jl")
 # Generic Operations
 include("group_operations/addition_operation.jl")
 include("group_operations/multiplication_operation.jl")
+include("group_operations/scalar_multiplication_operation.jl")
 
 # Actions
 include("group_actions/group_action_interface.jl")

### 2025-03-02 src/group_operations/multiplication_operation.jl

@@ -14,13 +14,6 @@ A group operation that is realised by a matrix multiplication.
 """
 struct MatrixMultiplicationGroupOperation <: AbstractMultiplicationGroupOperation end
 
-"""
-    ScalarMultiplicationGroupOperation <: AbstractMultiplicationGroupOperation
-
-A group operation that is realised by a scalar multiplication.
-"""
-struct ScalarMultiplicationGroupOperation <: AbstractMultiplicationGroupOperation end
-
 Base.:*(::Identity{MatrixMultiplicationGroupOperation}, p::Union{AbstractMatrix,Number}) = p
 function Base.:*(
     p::Union{AbstractMatrix,Number}, ::Identity{MatrixMultiplicationGroupOperation}
@@ -74,14 +67,13 @@ This can be computed in-place of `k`.
 """
 
 @doc "$(_doc_compose_mult)"
-compose(::LieGroup{𝔽,<:AbstractMultiplicationGroupOperation}, ::Any, ::Any) where {𝔽}
-
+compose(::LieGroup{𝔽,<:MatrixMultiplicationGroupOperation}, ::Any, ::Any) where {𝔽}
 @doc "$(_doc_compose_mult)"
 compose!(
     ::LieGroup{𝔽,<:AbstractMultiplicationGroupOperation}, ::Any, ::Any, ::Any
 ) where {𝔽}
 
-function _compose!(::LieGroup{𝔽,<:AbstractMultiplicationGroupOperation}, k, g, h) where {𝔽}
+function _compose!(::LieGroup{𝔽,<:MatrixMultiplicationGroupOperation}, k, g, h) where {𝔽}
     # perform the multiplication “safe”, that is, even when providing
     # one of the inputs `g,h`` and as output `k`
     (k === g || k === h) ? copyto!(k, g * h) : mul!(k, g, h)
@@ -235,53 +227,6 @@ function identity_element!(
     return copyto!(e, LinearAlgebra.I)
 end
 
-_doc_identity_element_scalar_mult = """
-    identity_element(G::LieGroup{𝔽,ScalarMultiplicationGroupOperation})
-    identity_element!(G::LieGroup{𝔽,ScalarMultiplicationGroupOperation}, e)
-
-Return the a point representation of the [`Identity`](@ref),
[…]

### 2025-03-02 src/group_operations/scalar_multiplication_operation.jl

@@ -1,0 +1,79 @@
+"""
+    ScalarMultiplicationGroupOperation <: AbstractMultiplicationGroupOperation
+
+A group operation that is realised by a scalar multiplication.
+"""
+struct ScalarMultiplicationGroupOperation <: AbstractMultiplicationGroupOperation end
+
+function _compose!(
+    ::LieGroup{𝔽,<:ScalarMultiplicationGroupOperation}, k::Union{Ref{T}, Array{T,0}}, g::T1, h::T2
+) where {𝔽, T1<:Number, T2<:Number, T<:Number}
+    k[] = ((k === g || k === h) ? copy(g * h) : g * h)
+    return k
+end
+
+function _compose!(
+    ::LieGroup{𝔽,<:ScalarMultiplicationGroupOperation}, k::Union{Ref{T}, Array{T,0}}, g::Union{Ref{T1}, Array{T1,0}}, h::T2
+) where {𝔽, T1<:Number, T2<:Number, T<:Number}
+    k[] = ((k === g || k === h) ? copy(g[] * h) : g[] * h)
+    return k
+end
+
+function _compose!(
+    ::LieGroup{𝔽,<:ScalarMultiplicationGroupOperation}, k::Union{Ref{T}, Array{T,0}}, g::T1, h::Union{Ref{T2}, Array{T2,0}}
+) where {𝔽, T1<:Number, T2<:Number, T<:Number}
+    k[] = ((k === g || k === h) ? copy(g * h[]) : g * h[])
+    return k
+end
+
+function _compose!(
+    ::LieGroup{𝔽,<:ScalarMultiplicationGroupOperation}, k::Union{Ref{T}, Array{T,0}}, g::Union{Ref{T1}, Array{T1,0}}, h::Union{Ref{T2}, Array{T2,0}}
+) where {𝔽, T1<:Number, T2<:Number, T<:Number}
+    k[] = ((k === g || k === h) ? copy(g[] * h[]) : g[] * h[])
+    return k
+end
+
+_doc_exp_scalar_mult = """
+    exp(G::LieGroup{𝔽,ScalarMultiplicationGroupOperation}, e::Identity{ScalarMultiplicationGroupOperation}, X, t::Number=1)
+    exp!(G::LieGroup{𝔽,ScalarMultiplicationGroupOperation}, g, e::Identity{ScalarMultiplicationGroupOperation}, X, t::Number=1)
+
[…]

### 2025-03-02 src/groups/circle_group.jl

@@ -82,12 +82,13 @@ Computes the Lie group exponential on the complex [`CircleGroup`](@ref), which c
 
 The Lie algebra is precisely the imaginary axis of the complex plane.
 
-This can be computed in-place of `X`.
+This can be computed in-place of `g`.
 ```math
 $(_tex(:exp)) ($(_math(:i))t) = $(_tex(:cos))(t) + $(_math(:i))$(_tex(:sin))(t)
 ```
 
 """
+
 
 @doc "$(_doc_exp_complex_circ)"
 exp(::CircleGroup, X)

### 2025-03-02 src/interface.jl

@@ -194,8 +194,29 @@ function compose(
     return g
 end
 
-function _compose(G::LieGroup, g, h)
-    k = ManifoldsBase.allocate_result(G, compose, g, h)
+function get_number_type(x::Number)
+    return typeof(x)
+end
+
+function get_number_type(x::Array{<:Number, 0})
+    return typeof(x[])
+end
+
+function get_number_type(x::Ref{<:Number})
+    return typeof(x[])
+end
+
+function _compose(G::LieGroup{𝔽,O}, g, h) where {𝔽,O<:AbstractGroupOperation}
+    if ManifoldsBase.representation_size(G) == ()
+        T = promote_type(get_number_type(g), get_number_type(h))
+        if T<:Number
+            k = fill(zero(T)) # Allocate a scalar zero value for representation size ()
+        else
+            throw(ArgumentError("Unsupported type for zero-dimensional array:  " * string(g) * ", "*  string(h)))
+        end
+    else
+        k = ManifoldsBase.allocate_result(G, compose, g, h)
+    end
     return _compose!(G, k, g, h)
 end
 

### 2025-03-02 test/groups/test_circle_group.jl

@@ -6,86 +6,86 @@ s = joinpath(@__DIR__, "..", "LieGroupsTestSuite.jl")
 !(s in LOAD_PATH) && (push!(LOAD_PATH, s))
 using LieGroupsTestSuite
 
-@testset "(complex) Circle Group" begin
+@testset "Complex Circle" begin
     C1 = CircleGroup()
-    z1, z2, z3 = 1, 1im, -1im
+
+
+
+    z1, z2, z3 = 1., 1.0im , -1.0im
     properties = Dict(
-        :Name => "The Circle group",
-        :Points => [z1, z2, z3],
+        :Name => "Array Points",
+        :Points => [fill(z1), fill(z2), fill(z3)],
+        :Mutating => true,
         :Rng => Random.MersenneTwister(),
         :Functions => [
-        # compose,
-        # conjugate,
-        # diff_conjugate,
-        # diff_inv,
-        # diff_left_compose,
-        # diff_right_compose,
-        # exp,
-        # hat,
-        # inv,
-        # inv_left_compose,
-        # inv_right_compose,
-        # is_identity,
-        # lie_bracket,
-        # log,
-        # rand,
-        # show,
-        # vee,
-    ],
+                compose,
+                # conjugate,
+                # diff_conjugate,
+                # diff_inv,
+                # diff_left_compose,
+                # diff_right_compose,
+                exp,
+                # hat,
+                #
+                # inv,
+                # inv_left_compose,
+                # inv_right_compose,
+                is_identity,
+                # lie_bracket,
+                # log,
+                # rand,
+                # show,
+                # vee,
+            ],
     )
-    expectations = Dict(:repr => "CircleGroup()")
+    expectations = Dict(
+        :repr => "CircleGroup()"
+    )
     test_lie_group(C1, properties, expectations)
 
     @testset "complex" begin
         C1 = CircleGroup()
         @test manifold_dimension(C1) == 1
-        e = Identity(C1)
-        s = 1im
-        t = 2 + 1im
-        @test compose(C1, e, s) == s
+        e=Identity(C1)
+        s = 1.0im
+        t = 2.0+1im     
+        @test compose(C1, e, s) == s   
         @test is_point(C1, s; error=:error)
         @test_throws DomainError is_point(C1, t; error=:error)
     end
 end
 
-@testset "real Circle Group" begin
-    C2 = RealCircleGroup()
-    x1, x2, x3 = 0, 1, -π
-    properties = Dict(
[…]

### 2025-03-02 test/runtests.jl

@@ -39,6 +39,7 @@ end
         include_test("groups/test_special_euclidean_group.jl")
         include_test("groups/test_special_unitary_group.jl")
         include_test("groups/test_translation_group.jl")
+        include_test("groups/test_circle_group.jl")
     end
     include("test_utils.jl")
     include("test_aqua.jl")

### 2025-03-02 tutorials/getstarted.qmd
add a final code example for the second section.
First full draft of the tutorial.

@@
 
-In a nutshell a Lie group $𝒢$ is a [manifold](https://en.wikipedia.org/wiki/Manifold) that also is a group a group operation.
+In a nutshell a Lie group $𝒢$ is a [manifold](https://en.wikipedia.org/wiki/Manifold) $ℳ$ that also is a group a group operation.
 A manifold informally be described as a set that “locally looks like a Euclidean space”
@@
 For more details on that, see also the [introductory tutorial](https://juliamanifolds.github.io/Manifolds.jl/stable/tutorials/getstarted/) there.
+A Lie group is usually written as a tuple $(ℳ,∘)$ os the manifold and the group operation thereon.
 
@@
 
+As an example, we compute
+
+```{julia}
+h = exp(SO3, g, X)
+```
+and its inverse is returning `X` as well.
+```{julia}
+log(SO3, g, h)
+```
+
 ## The Special Euclidean group $\mathrm{SE}(3)$ of rigid body motions
 
-# TODO
-* group action and apply
-* product of two Lie groups
-* semidirect product
+There are two further ingredients this tutorial needs to get towards rigid body motions.
+
+### Group actions
+
+For a group like $\mathrm{SO}(3)$ an action describes how points $g ∈ 𝒢$ can act on
+points from a Riemannian manifold $p ∈ ℳ$ – in short an action of a certain [`AbstractGroupActionType`](@ref)
+”combines“ $p$ and $g$ into a new element $q = σ_g(p)$.
+In the example of $𝒢=\mathrm{SO}(3)$ and $ℳ=ℝ^3$ the action is (just) the application of the
+rotation or the matrix-vector multiplication $σ_g(p)=gp$.
+A special case is, when $ℳ$ itself is also a Lie group.
+In the example this is the case, since together with vector addition $p+q$ we get the [`TranslationGroup`](@ref) $T(3) = (ℝ^3, +)$
+Depending on how the concatenation of two types, the [`AbstractLeftGroupActionType`](@ref) like the one here
+and a [`AbstractRightGroupActionType`](@ref).
+
+```{julia}
+T3 = TranslationGroup(3)
+```
+### Products and semidirect products of Lie groups
+
+Similar to [product manifolds](https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/#ProductManifold)
+(direct) [product Lie groups](../groups/product_group.md) are formed the same way. For two Lie groups
+$𝒢 = 𝒢_1 × 𝒢_2$ is the product of the two manifolds together with the component wise application of the group operations:
+the first group operation (of $𝒢_1$) to the first component and that of $𝒢_2$ to the second.
+
+Instead of the component wise or “not interacting” variant of the (direct) product Lie groups,
+for the [semidirect product Lie groups](../groups/semidirect_product_group.md) $𝒢 = 𝒢_1 ⋉ 𝒢_2$
+we also require an action of how $𝒢_1$ acts on $𝒢_2$.
+Semidirect here means that the first component of $𝒢$ is the same as for the direct product,
+but before applying the second group operation on $𝒢_2$ one of the elements is “acted upon”
+from an element of $𝒢_1$. The group operation reads for $(g_1,g_2), (h_1,h_2) ∈ 𝒢$
+
+```math
+(g_1,g_2) ∘ (h_1,h_2) := (g_1 ⋆ h_1, σ_{h_1}(g_2) ⋄ h_2).
+```
+
+### Rigid body motions
+
+We obtain the [special Euclidean group](../groups/special_euclidean_group.md)
+$\mathrm{SE}(3) = \mathrm{SO}(3) ⋉ T(3)`` where the group action is the one discussed as an example before.
+
+```{julia}
+SE3 = SpecialEuclideanGroup(3)
+```
+
+which we could also generate with `SO3` and `T3` from before as
+
+```{julia}
+SO3 ⋉ T3
+```
+
+This call employs the [`default_left_action`](@ref)`(SO3,T3)` to determine the action
+for the semidirect product Lie group. This is defined whenever for two Lie groups, their (left)
+action is clear, because there exists a reasonable default. Otherwise the full form
+
+```{julia}
+LeftSemidirectProductLieGroup(SO3, T3, LeftGroupOperationAction())
+```
+
+is necessary. Here, the first `Left` for the semidirect product refers to the fact that
+the left group acts on the right one before the right group operation is performed.
+[`LeftGroupOperationAction`](@ref) refers to that the group operation – left multiplication with a matrix – is applied here.
+
+For this case, any point $\mathrm{SE}(3)$ is a tuple of a rotation matrix and a vector.
+We model this using [RecursiveArrayTools.jl](https://docs.sciml.ai/RecursiveArrayTools/stable/),
+namely its `ArrayPartition`.
+For example with the rotation matrix `g` from before we have
+
+```{julia}
+g1 = ArrayPartition(g, [1.0, 0.0, 2.0])
+```
+
+and for a pure translation we can reuse `e` as in
+
+```{julia}
+h1 = ArrayPartition(e, [0.0, 3.0, 0.0])
+```
+
+and we can summarize a few operations from before: both are valid points on `SE3`
+
+```{julia}
+(is_point(SE3, g1), is_point(SE3, h1))
+```
+
+we can perform the group operation
+
+```{julia}
+gh1 = compose(SE3, g1, h1)
+```
+
+apply the Lie group exponential
+
+```{julia}
+Y = log(SE3, gh1)
+```
+
+and look at the manifold dimension, it is the sum of the single ones
+
+```{julia}
+manifold_dimension(SE3)
+```
+
+and its coordinates as
+
+```{julia}
+vee(LieAlgebra(SE3), gh1)
+```
+
+which actually is component wise again, the first 3 values refer to the rotation part,
+the second three to the translation part.
 

### 2025-03-04 docs/make.jl
Write a bit of a nicer start page.

@@ -58,7 +58,7 @@ if "--quarto" ∈ ARGS
     end
 else
     # fallback to at least create empty files for tutorials that are directly linked from the docs
-    #    touch(joinpath(@__DIR__, "src/tutorials/Optimize.md"))
+    touch(joinpath(@__DIR__, "src/tutorials/getstarted.md"))
 end
 
 tutorials_in_menu = true

### 2025-03-04 src/groups/validation_group.jl
Refine ValidationLieGroup doc string.

@@ -18,7 +18,13 @@ Current contexts are
 # Fields
 
 * `lie_group::L` the [`AbstractLieGroup`](@ref) to be decorated
-* ``
+* `mode::Symbol`: The mode to be used for error handling, either `:error` or `:warn`
+* `ignore_contexts::AbstractVector{Symbol}`: store contexts to be ignored of validation.
+* `ignore_functions::Dict{<:Function,<:Union{Symbol,<:AbstractVector{Symbol}}`:
+  store contexts to be ignored with in a function or its mutating variant.
+
+where all but the first field are analogous to the setups of the [`ValidationManifold`](@extref `ManifoldsBase.ValidationManifold`).
+We refer to those docs for more examples on their meaning.
 
 # Constructor
     ValidationLieGroup(L::AbstractLieGroup, check_manifold=true; kwargs...)

### 2025-03-04 src/groups/validation_group.jl
Add a ValidationLieGroup (#22)
* refactor the interface to be based on an AbstractLieGroup
* Initial implementation of the interface for the ValidationLieGroup
* add (dummy for now) test file.
* Add changelog entry.
* unifications of interfaces and resolution of ambiguities.
* Setup the first tests.
* Finish most LieGroup functions, the LieAlgebra ones are still missing.
* Leave Identity as is, the other one would change too many types as well; implement norm
* Refine ValidationLieGroup doc string.

@@ -1,0 +1,766 @@
+"""
+    ValidationLieGroup{L<:AbstractLieGroup} <: AbstractLieGroup
+
+A Lie group to add tests to input parameters and output values of functions defined
+for [`LieGroups`](@ref).
+
+
+Using the `ignore_contexts` keyword allows to specify a single `Symbol` or a vector of `Symbols`
+Of which contexts to ignore.
+
+Current contexts are
+* `:All`: disable all checks
+* `:Point`: checks for points
+* `:Algebra`: checks related to the [`LieAlgebra`](@ref)
+* `:Output`: checks for output
+* `:Input`: checks for input variables
+
+# Fields
+
+* `lie_group::L` the [`AbstractLieGroup`](@ref) to be decorated
+* `mode::Symbol`: The mode to be used for error handling, either `:error` or `:warn`
+* `ignore_contexts::AbstractVector{Symbol}`: store contexts to be ignored of validation.
+* `ignore_functions::Dict{<:Function,<:Union{Symbol,<:AbstractVector{Symbol}}`:
+  store contexts to be ignored with in a function or its mutating variant.
+
+where all but the first field are analogous to the setups of the [`ValidationManifold`](@extref `ManifoldsBase.ValidationManifold`).
+We refer to those docs for more examples on their meaning.
+
+# Constructor
+    ValidationLieGroup(L::AbstractLieGroup, check_manifold=true; kwargs...)
+
+Generate the Validation Lie Group for the given [`AbstractLieGroup`](@ref) `L`.
+If `check_manifold` is set to `true` the inner manifold is additionally wrapped
+in a [`ValidationManifold`](@extref `ManifoldsBase.ValidationManifold`).
+All suitable keywords are passed to the constructor of the validation manifold as well.
+
+# Keyword arguments
+
+* `error::Symbol=:error`: specify how errors in the validation should be reported.
+  this is passed to [`is_point`](@extref `ManifoldsBase.is_point`) and [`is_vector`](@extref `ManifoldsBase.is_vector`) as the `error` keyword argument.
+  Available values are `:error`, `:warn`, `:info`, and `:none`. Every other value is treated as `:none`.
[…]

### 2025-03-04 test/LieGroupsTestSuite.jl/LieGroupsTestSuite.jl

@@ -53,7 +53,7 @@ end
 #
 # --- A
 """
-    test_adjoint(G::LieGroup, g, X; kwargs...)
+    test_adjoint(G::AbstractLieGroup, g, X; kwargs...)
 
 Test  `adjoint` function for a given Lie group element `g` and a Lie Algebra vector `X`
 
@@ -62,7 +62,7 @@ Test  `adjoint` function for a given Lie group element `g` and a Lie Algebra vec
   default from `diff_conjugate` is used
 * `test_mutating::Bool=true`: test the mutating functions
 """
-function test_adjoint(G::LieGroup, g, X; expected=missing, test_mutating::Bool=true)
+function test_adjoint(G::AbstractLieGroup, g, X; expected=missing, test_mutating::Bool=true)
     @testset "adjoint" begin
         v = if ismissing(expected)
             diff_conjugate(G, g, identity_element(G, typeof(X)), X)
@@ -108,7 +108,7 @@ end
 #
 # --- C
 """
-    test_compose(G::LieGroup, g, h; kwargs...)
+    test_compose(G::AbstractLieGroup, g, h; kwargs...)
 
 Test  `compose` for given Lie group elements `g`, `h`.
 
@@ -121,7 +121,7 @@ Test  `compose` for given Lie group elements `g`, `h`.
 * `test_mutating::Bool=true`: test the mutating functions
 """
 function test_compose(
-    G::LieGroup,
+    G::AbstractLieGroup,
     g,
     h;
     atol::Real=0,
@@ -167,7 +167,7 @@ function test_compose(
 end
 
 """
-    test_conjugate(G::LieGroup, g, h; expected=missing)
+    test_conjugate(G::AbstractLieGroup, g, h; expected=missing)
 
 Test  `conjugate`.
 
@@ -178,7 +178,12 @@ Test  `conjugate`.
 * `test_mutating::Bool=true`: test the mutating functions
 """
 function test_conjugate(
-    G::LieGroup, g, h; expected=missing, test_default::Bool=true, test_mutating::Bool=true
+    G::AbstractLieGroup,
+    g,
+    h;
+    expected=missing,
+    test_default::Bool=true,
+    test_mutating::Bool=true,
 )
     @testset "conjugate" begin
         k1 = conjugate(G, g, h)
@@ -255,7 +260,7 @@ function test_diff_group_apply(
 end
 
 """
[…]

### 2025-03-04 test/groups/test_translation_group.jl

@@ -17,6 +17,7 @@ begin
             adjoint,
             compose,
             conjugate,
+            diff_conjugate,
             diff_inv,
             diff_left_compose,
             diff_right_compose,

### 2025-03-04 src/interface.jl
Update src/interface.jl
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@ -667,8 +667,7 @@ function jacobian_conjugate!(
         c[j] = 1
         get_vector!(𝔤, X, c, B)         # store ``X_j`` in X
         diff_conjugate!(G, X, g, h, X) # compute the differential in-place
-        get_coordinates!(𝔤, c, X, B)   # compute its coordinates in B
-        J[:, j] .= c                    # store the result in J
+        get_coordinates!(𝔤, view(J, :, j), X, B)   # compute its coordinates in B and store the result in J
     end
     return J
 end

### 2025-03-04 src/LieGroups.jl
Fix a hat/vee confusion
it is already defined in ManifoldsBase so let's extend that function with methods instead of coming up with our own.

@@ -104,6 +104,7 @@ export CompositeManifoldError
 export adjoint, adjoint!, apply, apply!
 export base_lie_group, base_manifold
 export compose, compose!
+export conjugate, conjugate!, diff_conjugate, diff_conjugate!
 export default_left_action,
     default_right_action,
     diff_apply,
@@ -118,7 +119,6 @@ export get_coordinates, get_coordinates!, get_vector, get_vector!
 export hat, hat!
 export inv, inv!, inv_left_compose, inv_left_compose!, inv_right_compose, inv_right_compose!
 export isapprox, is_point, is_vector
-export conjugate, conjugate!, diff_conjugate, diff_conjugate!
 export exp, exp!
 export identity_element, identity_element!, is_identity, inv, inv!, diff_inv, diff_inv!
 export jacobian_conjugate, jacobian_conjugate!

### 2025-03-04 src/groups/validation_group.jl

@@ -365,7 +365,7 @@ function ManifoldsBase.exp!(G::ValidationLieGroup, h, g, X; kwargs...)
     return g
 end
 
-function hat(
+function ManifoldsBase.hat(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, c; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     G = base_lie_group(𝔤).lie_group
@@ -373,7 +373,7 @@ function hat(
     is_point(𝔤, X; widthin=hat, context=(:Output,), kwargs...)
     return ValidationLieAlgebraTangentVector(X)
 end
-function hat(
+function ManifoldsBase.hat(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, c, T::Type; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     G = base_lie_group(𝔤).lie_group
@@ -381,7 +381,7 @@ function hat(
     is_point(𝔤, X; widthin=hat, context=(:Output,), kwargs...)
     return ValidationLieAlgebraTangentVector(X)
 end
-function hat!(
+function ManifoldsBase.hat!(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, X, c; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     G = base_lie_group(𝔤).lie_group
@@ -722,14 +722,14 @@ function Base.show(io::IO, G::ValidationLieGroup)
     return print(io, s)
 end
 
-function vee(
+function ManifoldsBase.vee(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, X; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     is_point(𝔤, X; widthin=vee, context=(:Input,), kwargs...)
     G = base_lie_group(𝔤).lie_group
     return vee(LieAlgebra(G), unwrap_validation(X))
 end
-function vee!(
+function ManifoldsBase.vee!(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, c, X; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     is_point(𝔤, X; widthin=vee, context=(:Input,), kwargs...)

### 2025-03-04 test/LieGroupsTestSuite.jl/LieGroupsTestSuite.jl
Fix another thing I missed in the merge.

@@ -211,7 +211,7 @@ Test that `copyto!` works also when copying over an `Identity`.
 The point `g` can be any point _but_ the `identity_element`.
 The group has to be a mutating one, that is, not work on isbit types.
 """
-function test_copyto(G::LieGroup, g)
+function test_copyto(G::AbstractLieGroup, g)
     @testset "copyto!" begin
         k = copy(G, g)
         e = Identity(G)

### 2025-03-04 tutorials/getstarted.qmd
Apply suggestions from code review
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@
 
-In a nutshell a Lie group $𝒢$ is a [manifold](https://en.wikipedia.org/wiki/Manifold) $ℳ$ that also is a group a group operation.
-A manifold informally be described as a set that “locally looks like a Euclidean space”
+In a nutshell a Lie group $𝒢$ is a [manifold](https://en.wikipedia.org/wiki/Manifold) $ℳ$ that also is a group with a group operation.
+A manifold can informally be described as a set that “locally looks like a Euclidean space”
 and has some way to measure angles at every point, to be precise an inner product, that is allowed to (smoothly)
@@
 A [group](https://en.wikipedia.org/wiki/Group_(mathematics)) means that on the set there also
-exists an operation, within this package usually denoted by $∘: 𝒢 × 𝒢 → 𝒢$$,
+exists an operation, within this package usually denoted by $∘: 𝒢 × 𝒢 → 𝒢$,
 `LieGroups.jl` uses the manifolds defined in [`Manifolds.jl`](https://juliamanifolds.github.io/Manifolds.jl/stable/).
 For more details on that, see also the [introductory tutorial](https://juliamanifolds.github.io/Manifolds.jl/stable/tutorials/getstarted/) there.
-A Lie group is usually written as a tuple $(ℳ,∘)$ os the manifold and the group operation thereon.
+A Lie group is usually written as a tuple $(ℳ,∘)$ with the manifold and the group operation thereon.
 
@@
 ```math
-R_α = \begin{pmatrix} \cos(α),& -\sin(α)\\ \sin(α) & \cos(α) \end{pmatrix},\qquad α ∈ ℝ,
+R_α = \begin{pmatrix} \cos(α)& -\sin(α)\\ \sin(α) & \cos(α) \end{pmatrix},\qquad α ∈ ℝ,
 ```
@@
 when applied to a point $x ∈ ℝ^2$ by computing $R_αx$.
-Matrix multiplication $R_β = R_{α_1}*R_{α_1}$ does yield a new rotation and also fulfils all other
+Matrix multiplication $R_β = R_{α_1} R_{α_2}$ does yield a new rotation and also fulfills all other
 properties to yield a group operation. Using trigonometric identities we observe that $β = α_1 + α_2$.
@@
 ```
-The element `g` can be seen as a rotation by $π/3$ in the x-y-plane combines with a rotation by $π/4$ in the x-z plane.
+The element `g` can be seen as a rotation by $π/3$ in the x-y-plane combined with a rotation by $π/4$ in the x-z plane.
 
@@
 The exponential function $\exp_𝒢: 𝔤 → 𝒢$ maps Lie algebra tangent vectors $X ∈ T_e𝒢$
-to a point on the Lie group. This is implemented in [`exp(G::LieGroup, X)`](@ref). It's inverse
+to a point on the Lie group. This is implemented in [`exp(G::LieGroup, X)`](@ref). Its inverse
 is the Lie group logarithmic function [`log(G::LieGroup, g)`](@ref).
@@
 With respect to these geodesics and the representation in the Lie algebra, parallel transport
-simplifies to the identity. To still access the Riemannian exponential map with recpect to the
-metric (compatible to the Levi-cevita connection) on the underlying Riemannian manifold, use
+simplifies to the identity. To still access the Riemannian exponential map with respect to the
+metric (compatible to the Levi-Civita connection) on the underlying Riemannian manifold, use
 `exp(base_manifold(G), g, X)`.

### 2025-03-05 src/LieGroups.jl
Implement a default for `jacobian_conjugate` (#23)
* Implement a default for `jacobian_conjugate`

---------

Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@ -21,8 +21,7 @@ using ManifoldsBase: RealNumbers, ComplexNumbers, ℝ, ℂ, internal_value
 # = Compatibility (and a bit of type piracy for now)
 # The following imports are necessary to use Manifolds.jl 0.10 with Lie groups
 # The line is removed when the Groups are removed from possibly 0.11
-import Manifolds:
-    apply, apply!, compose, identity_element, is_identity, hat, hat!, vee, vee!
+import Manifolds: apply, apply!, compose, identity_element, is_identity
 # Both define the following structs, so these for now lead to asking for explicit prefixes
 # Manifolds: Identity, TranslationGroup
 include("documentation_glossary.jl")
@@ -105,6 +104,7 @@ export CompositeManifoldError
 export adjoint, adjoint!, apply, apply!
 export base_lie_group, base_manifold
 export compose, compose!
+export conjugate, conjugate!, diff_conjugate, diff_conjugate!
 export default_left_action,
     default_right_action,
     diff_apply,
@@ -119,13 +119,12 @@ export get_coordinates, get_coordinates!, get_vector, get_vector!
 export hat, hat!
 export inv, inv!, inv_left_compose, inv_left_compose!, inv_right_compose, inv_right_compose!
 export isapprox, is_point, is_vector
-export conjugate, conjugate!, diff_conjugate, diff_conjugate!
 export exp, exp!
 export identity_element, identity_element!, is_identity, inv, inv!, diff_inv, diff_inv!
 export jacobian_conjugate, jacobian_conjugate!
 export lie_bracket, lie_bracket!, log, log!
 export manifold_dimension
-export norm
+export norm, number_of_coordinates
 export injectivity_radius
 export rand, rand!
 export switch

### 2025-03-05 src/Lie_algebra/Lie_algebra_interface.jl

@@ -195,16 +195,18 @@ function ManifoldsBase._get_vector!(
     return ManifoldsBase.get_vector!(base_manifold(G), X, identity_element(G, T), c, B)
 end
 
-@inline function get_vector_lie(𝔤::LieAlgebra, c, N)
+@inline function get_vector_lie(𝔤::LieAlgebra, c, B::DefaultLieAlgebraOrthogonalBasis)
     X = zero_vector(𝔤)
-    return get_vector_lie!(𝔤, X, c, N)
+    return get_vector_lie!(𝔤, X, c, B)
 end
-@inline function get_vector_lie(𝔤::LieAlgebra, c, N, T::Type)
+@inline function get_vector_lie(
+    𝔤::LieAlgebra, c, B::DefaultLieAlgebraOrthogonalBasis, T::Type
+)
     X = zero_vector(𝔤, T)
-    return get_vector_lie!(𝔤, X, c, N)
+    return get_vector_lie!(𝔤, X, c, B)
 end
 @inline function get_vector_lie!(
-    𝔤::LieAlgebra, X::T, c, B::DefaultLieAlgebraOrthogonalBasis{𝔽}
+    𝔤::LieAlgebra, X::T, c, ::DefaultLieAlgebraOrthogonalBasis{𝔽}
 ) where {T,𝔽}
     G = base_lie_group(𝔤)
     return get_vector!(
@@ -237,18 +239,16 @@ Technically, `hat` is a specific case of [`get_vector`](@ref) and is implemented
 [`DefaultLieAlgebraOrthogonalBasis`](@ref).
 """
 
-# function hat end
 @doc "$(_doc_hat)"
-function hat(𝔤::LieAlgebra, c)
+function ManifoldsBase.hat(𝔤::LieAlgebra, c)
     return get_vector(𝔤, c, DefaultLieAlgebraOrthogonalBasis())
 end
-function hat(𝔤::LieAlgebra, c, T::Type)
+function ManifoldsBase.hat(𝔤::LieAlgebra, c, T::Type)
     return get_vector(𝔤, c, DefaultLieAlgebraOrthogonalBasis(); tangent_vector_type=T)
 end
 
-# function hat! end
 @doc "$(_doc_hat)"
-function hat!(𝔤::LieAlgebra, X, c)
+function ManifoldsBase.hat!(𝔤::LieAlgebra, X, c)
     get_vector!(𝔤, X, c, DefaultLieAlgebraOrthogonalBasis())
[…]

### 2025-03-05 src/groups/validation_group.jl

@@ -365,7 +365,7 @@ function ManifoldsBase.exp!(G::ValidationLieGroup, h, g, X; kwargs...)
     return g
 end
 
-function hat(
+function ManifoldsBase.hat(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, c; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     G = base_lie_group(𝔤).lie_group
@@ -373,7 +373,7 @@ function hat(
     is_point(𝔤, X; widthin=hat, context=(:Output,), kwargs...)
     return ValidationLieAlgebraTangentVector(X)
 end
-function hat(
+function ManifoldsBase.hat(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, c, T::Type; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     G = base_lie_group(𝔤).lie_group
@@ -381,7 +381,7 @@ function hat(
     is_point(𝔤, X; widthin=hat, context=(:Output,), kwargs...)
     return ValidationLieAlgebraTangentVector(X)
 end
-function hat!(
+function ManifoldsBase.hat!(
     𝔤::LieAlgebra{𝔽,O,<:ValidationLieGroup}, X, c; kwargs...
 ) where {𝔽,O<:AbstractGroupOperation}
     G = base_lie_group(𝔤).lie_group
@@ -573,7 +573,7 @@ function jacobian_conjugate(
 )
     is_point(G, g; widthin=jacobian_conjugate, context=(:Input,), kwargs...)
     is_point(G, h; widthin=jacobian_conjugate, context=(:Input,), kwargs...)
-    J = jacobian_conjugate(G.lie_group, unwrap_validation(g), unwrap_validation(h), args...)
+    J = jacobian_conjugate(G.lie_group, unwrap_validation(g), unwrap_validation(h))
     return J
 end
 function jacobian_conjugate!(
@@ -586,7 +586,7 @@ function jacobian_conjugate!(
 )
     is_point(G, g; widthin=jacobian_conjugate, context=(:Input,), kwargs...)
     is_point(G, h; widthin=jacobian_conjugate, context=(:Input,), kwargs...)
-    jacobian_conjugate!(G.lie_group, J, unwrap_validation(g), unwrap_validation(h), args...)
[…]

### 2025-03-05 src/interface.jl

@@ -21,7 +21,7 @@ at the [`identity_element`](@ref) on the [`base_manifold](@ref base_manifold(::A
 
 !!! note
     In order to implement the corresponding [`get_coordinates`](@ref) and [`get_vector`](@ref) functions,
-    define `get_coordinates_lie(::AbstractLieGroup, p, X, B)` and `get_vector_lie(::AbstractLieGroup, p, X, B)`, resp.
+    define `get_coordinates_lie(::AbstractLieAlgebra, X, B)` and `get_vector_lie(::AbstractLieAlgebra, X, B)`, resp.
 """
 struct DefaultLieAlgebraOrthogonalBasis{𝔽} <:
        ManifoldsBase.AbstractOrthogonalBasis{𝔽,ManifoldsBase.TangentSpaceType} end
@@ -644,20 +644,26 @@ function ManifoldsBase.isapprox(
     return false
 end
 _doc_jacobian_conjugate = """
-    jacobian_conjugate(G::AbstractLieGroup, g, h, B::AbstractBasis=DefaultLieAlgebraOrthogonalBasis())
-    jacobian_conjugate!(G::AbstractLieGroup, J, g, h, B::AbstractBasis=DefaultLieAlgebraOrthogonalBasis())
+    jacobian_conjugate(G::AbstractLieGroup, g, h; basis::AbstractBasis=DefaultLieAlgebraOrthogonalBasis(); X=zero_vector(LieAlgebar(G)))
+    jacobian_conjugate!(G::AbstractLieGroup, J, g, h; basis::AbstractBasis=DefaultLieAlgebraOrthogonalBasis(); X=zero_vector(LieAlgebar(G)))
 
 Compute the Jacobian of the [`conjugate`](@ref) ``c_g(h) = g$(_math(:∘))h$(_math(:∘))g^{-1}``,
-with respect to an [`AbstractBasis`](@extref `ManifoldsBase.AbstractBasis`).
+with respect to an [`AbstractBasis`](@extref `ManifoldsBase.AbstractBasis`) of the [`LieAlgebra`](@ref).
 
-This can be seen as a matrix representation of the [`diff_conjugate`](@ref) ``D(c_g(h))[X]``
-with respect to the given basis.
+A default is implemented using [`diff_conjugate`](@ref) ``D(c_g(h))[X]``:
+the ``j``th column of of the Jacobian matrix ``J`` are given by the coefficients of
+the tangent vector `D(c_g(h))[X_j]`` with respect to the basis ``B``,
[…]

### 2025-03-05 test/LieGroupsTestSuite.jl/LieGroupsTestSuite.jl

@@ -203,6 +203,36 @@ function test_conjugate(
     return nothing
 end
 
+"""
+    test_copyto(G::LieGroup, g)
+
+Test that `copyto!` works also when copying over an `Identity`.
+
+The point `g` can be any point _but_ the `identity_element`.
+The group has to be a mutating one, that is, not work on isbit types.
+"""
+function test_copyto(G::AbstractLieGroup, g)
+    @testset "copyto!" begin
+        k = copy(G, g)
+        e = Identity(G)
+        copyto!(G, k, e)
+        # also check that this holds both ways
+        @test isapprox(G, k, Identity(G))
+        @test isapprox(G, Identity(G), k)
+        # Test that copying back also works
+        copyto!(G, k, g)
+        @test isapprox(G, k, g)
+        # copy into identity only works if provided the identity
+        @test copyto!(G, e, Identity(G)) == e
+        # but then also always returns e
+        copyto!(G, k, e)
+        @test copyto!(G, e, k) == e
+        # and fails if the point to copy in is not (numerically) e
+        @test_throws DomainError copyto!(G, e, g)
+        return nothing
+    end
+end
+
 #
 #
 # --- D
@@ -346,36 +376,6 @@ function test_diff_right_compose(
         if !ismissing(expected)
             @test isapprox(𝔤, Y1, expected)
         end
-    end
-end
-
-"""
-    test_copyto(G::AbstractLieGroup, g)
-
-Test that `copyto!` works also when copying over an `Identity`.
-
-The point `g` can be any point _but_ the `identity_element`.
-The group has to be a mutating one, that is, not work on isbit types.
-"""
-function test_copyto(G::AbstractLieGroup, g)
-    @testset "copyto!" begin
-        k = copy(G, g)
-        e = Identity(G)
-        copyto!(G, k, e)
-        # also check that this holds both ways
-        @test isapprox(G, k, Identity(G))
-        @test isapprox(G, Identity(G), k)
-        # Test that copying back also works
-        copyto!(G, k, g)
-        @test isapprox(G, k, g)
-        # copy into identity only works if provided the identity
-        @test copyto!(G, e, Identity(G)) == e
-        # but then also always returns e
-        copyto!(G, k, e)
-        @test copyto!(G, e, k) == e
-        # and fails if the point to copy in is not (numerically) e
-        @test_throws DomainError copyto!(G, e, g)
[…]

### 2025-03-05 test/groups/test_translation_group.jl

@@ -28,6 +28,7 @@ begin
             inv_left_compose,
             inv_right_compose,
             is_identity,
+            jacobian_conjugate,
             lie_bracket,
             log,
             rand,

### 2025-03-05 test/groups/test_validation_group.jl

@@ -27,6 +27,7 @@ using LieGroupsTestSuite
         inv_right_compose,
         is_identity,
         lie_bracket,
+        jacobian_conjugate,
         log,
         rand,
         show,

### 2025-03-05 tutorials/getstarted.qmd
Rephrase the inner product point.

@@
 In a nutshell a Lie group $𝒢$ is a [manifold](https://en.wikipedia.org/wiki/Manifold) $ℳ$ that also is a group with a group operation.
+
 A manifold can informally be described as a set that “locally looks like a Euclidean space”
-and has some way to measure angles at every point, to be precise an inner product, that is allowed to (smoothly)
-depend on the point we are at.
+and has some way to measure angles at every point.
+Formally, this yields an inner product that depends (smoothly) on the point one is at.
+For Lie groups this will be a bit easier.
+
 A [group](https://en.wikipedia.org/wiki/Group_(mathematics)) means that on the set there also

### 2025-03-05 tutorials/getstarted.qmd
avoid the modulo propsed in the PR but still mention the periodicity.

@@
 Matrix multiplication $R_β = R_{α_1} R_{α_2}$ does yield a new rotation and also fulfills all other
-properties to yield a group operation. Using trigonometric identities we observe that $β = α_1 + α_2$.
+properties to yield a group operation.
+Using trigonometric identities we observe that $R_β = R_{α_1+α_2}$.
+Since the angle of rotation is $2π$-periodic, we also see here, that e.g. for small angles around $0$,
+this behaves like the real line, but globally it differs, since $R_{α} = R_{α+2π}$
 

### 2025-03-05 tutorials/getstarted.qmd
Adress the remaining points.

@@
 ```{julia}
-using LieGroups, LinearAlgebra, RecursiveArrayTools
+using LieGroups, LinearAlgebra, RecursiveArrayTools, Rotations
 SO2 = SpecialOrthogonalGroup(2)
@@
 SO3 = SpecialOrthogonalGroup(3)
-g = [cos(π/3) -sin(π/3) 0.0; sin(π/3) cos(π/3) 0.0; 0.0 0.0 1.0]* [cos(π/4) 0.0 -sin(π/4); 0.0 1.0 0.0; sin(π/4) 0.0 cos(π/4)]
+g = RotZ(π/3)*RotY(π/4)
 e = Matrix{Float64}(I,3,3)
@@
 Consider the function of left group composition $λ_g(h) = g∘h$.
-Its differential (or push forward) $Dλ_g(e)$ maps from $T_e𝒢$ to $T_g𝒢$
+Its ``[differential](@ref diff_left_compose(G::AbstractLieGroup, g, h, X))``{=commonmark} (or push forward) $Dλ_g(e)$ maps from $T_e𝒢$ to $T_g𝒢$
 and is a linear bijection between both spaces.
@@
 since it is invariant if you use the differential of the left composition to identify tangent vectors.
-We can even “spare” the pullbacks in this definition, when spring tangent vectors as their representants in the Lie algebra.
+We can even “spare” the pullbacks in this definition, when representing tangent vector alsways by their correspnding representants in the Lie algebra.
 This is the default in `LieGroups.jl`.
+
+Alternatively one can start with the right composition and its ``[differential](@ref diff_right_compose(G::AbstractLieGroup, h, g, X))``{=commonmark} to arrive an right-invariant vector fields and a right-invariant metric.
+
+The left-invariant metric is in general not right-invariant, except for compact Lie groups and their direct products with Translations.
 
@@
 ```{julia}
-vee(LieAlgebra(SE3), gh1)
+vee(LieAlgebra(SE3), Y)
 ```

### 2025-03-06 tutorials/getstarted.qmd
Update tutorials/getstarted.qmd
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@
 The term “exponential” has at least three different meanings throughout Lie groups and Riemannian manifolds.
-To be precise, we call the just introduced one “exponential function”, since it often agrees with the [exponential](https://en.wikipedia.org/wiki/Exponential_function).
+To be precise, we call the just introduced one “exponential function”, since it often agrees with the [matrix exponential](https://en.wikipedia.org/wiki/Matrix_exponential).
 

### 2025-03-06 tutorials/getstarted.qmd
Update tutorials/getstarted.qmd
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@
 We obtain the [special Euclidean group](../groups/special_euclidean_group.md)
-$\mathrm{SE}(3) = \mathrm{SO}(3) ⋉ T(3)`` where the group action is the one discussed as an example before.
+$\mathrm{SE}(3) = \mathrm{SO}(3) ⋉ T(3)$ where the group action is the one discussed as an example before.
 

### 2025-03-06 tutorials/getstarted.qmd
Update tutorials/getstarted.qmd
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>
A get-started tutorial (#24)
* Start tutorial.
* Improve technical details print.
* Continue the tutorial.

---------

Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@
 
-The left-invariant metric is in general not right-invariant, except for compact Lie groups and their direct products with Translations.
+The left-invariant metric is in general not right-invariant, except for compact Lie groups and their direct products with translations, see for example [LatifiToomanian:2013](@cite) for an even more general proof.
 
@@
 
-and look at the manifold dimension, it is the sum of the single ones
+and look at the manifold dimension, it is the sum of dimensions of its components
 
@@
 
-and its coordinates as
+The coordinates of `Y` are
 

### 2025-03-06 docs/make.jl

@@ -93,6 +93,9 @@ if run_quarto || run_on_CI
         run(`quarto render $(tutorials_folder)`)
         return nothing
     end
+else
+    # fallback to at least create empty files for tutorials that are directly linked from the docs
+    touch(joinpath(@__DIR__, "src/tutorials/getstarted.md"))
 end
 
 # (d) load necessary packages for the docs

### 2025-03-06 src/interface.jl

@@ -394,7 +394,7 @@ _doc_exponential = """
     exp(G::AbstractLieGroup, X::T)
     exp!(G::AbstractLieGroup, g, X)
 
-    Compute the (Lie group) exponential function
+Compute the (Lie group) exponential function
 
 ```math
 $(_tex(:exp))_{$(_math(:G))}: $(_math(:𝔤)) → $(_math(:G)),$(_tex(:qquad)) $(_tex(:exp))_{$(_math(:G))}(X) = γ_X(1),

### 2025-03-06 src/LieGroups.jl
Initial sketch

@@ -36,7 +36,7 @@ include("group_operations/multiplication_operation.jl")
 # Actions
 include("group_actions/group_action_interface.jl")
 include("group_actions/group_operation_action.jl")
-
+include("group_actions/linear_representation.jl")
 # Meta Lie groups
 include("groups/power_group.jl")
 include("groups/product_group.jl")

### 2025-03-06 src/group_actions/linear_representation.jl

@@ -1,0 +1,17 @@
+"""
+   LinearRepresentation{T<:AbstractGroupActionType} <: AbstractGroupActionType
+
+Indicate that a certain group action type has a linear representation.
+
+# Fields
+* ` type` – the [`AbstractGroupActionType`](@ref) that has a linear representation
+"""
+struct LinearRepresentation{T<:AbstractGroupActionType} <: AbstractGroupActionType
+    type::T
+end
+
+function diff_apply!(
+    A::GroupAction{<:LinearRepresentation{<:AbstractGroupActionType},L,M}, Y, g, p, X
+)
+    return apply!(A, Y, g, X)
+end

### 2025-04-30 tutorials/getstarted.qmd
Fix date in tutorial.

@@
 using Dates
-println("This tutorial was last rendered $(Dates.format(now(), "U d, Y, H:M:S")).");
+println("This tutorial was last rendered $(Dates.format(now(), "U d, Y, HH:MM:SS")).");
 ```

### 2025-10-25 paper/header.tex
A bit of work on the setup

% **************GENERATED FILE, DO NOT EDIT**************

\title{Groups and smooth geometry using LieGroups.jl}

\author[1]{Ronny Bergmann}
\affil[1]{Department of Mathematical Sciences, Norwegian University of Science and Technology, Trondheim, Norway}

\keywords{Julia, Riemannian manifolds, Lie groups, Differential geometry}

\hypersetup{
pdftitle = {Groups and smooth geometry using LieGroups.jl},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {Ronny Bergmann},
pdfkeywords = {Julia, Riemannian manifolds, Lie groups, Differential geometry},
}

### 2025-10-25 paper/paper.tex
A bit of work on the setup

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{booktabs}

\hypersetup{colorlinks=true}

\begin{document}

\input{header}

\maketitle

\begin{abstract}
    \href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{\texttt{LieGroups.jl}} is a
    Julia package that provides both an interface to define as well as work with Lie groups,
    the corresponding Lie algebra as well as group actions.

    This paper presents the main features of the interfaces and how that is integrated within
    the \verb|JuliaManifolds| ecosystem. We further present an overview on existing Lie groups
    implemented in \texttt{LieGroups.jl} as well as how to get started to use the package in practice.
\end{abstract}

\section{Introduction}

\cite{Bergmann:2022}, \cite{AxenBaranBergmannRzecki:2023}.

\section{Notation}

\section{The interface}
\subsection{Lie groups}

\cite{HilgertNeeb:2012}

\subsection{Lie algebras}
\subsection{Group actions}

\section{Implemented Lie groups}
\begin{table}[tbp]
    \centering
    \caption{Implemented Lie groups in \texttt{LieGroups.jl} as of version 0.1.6}
    \begin{tabular}{@{}lll@{}}
        \toprule
        \textbf{Group} & \textbf{Symbol} & comment/code\\
        \midrule
        \verb|CircleGroup()| & \(\mathbb{S}^1\) & 3 representations \\
        \verb|GeneralLinearGroup(n, F)| & \(\mathrm{GL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
        \verb|HeisenbergGroup(n)| & \verb|H(n)| \\
        \verb|OrthogonalGroup(n)| & \(\mathrm{O}(n)\) &\\
        \midrule
        \verb|PowerLieGroup(G, n)| & \(\mathcal G^n\) & \verb|G^n|\\
        \verb|ProductLieGroup(G1, G2,...)| & \(\mathcal G_1 \times \mathcal G_2 \times \ldots\) & \verb|G1| $\times$ \verb|G2| $\times$ \ldots\\
        Semidirect product group & \(\mathcal G_1 \ltimes \mathcal G_2\) & \verb|G1| $\ltimes$ \verb|G2|\\
                                 & \(\mathcal G_1 \rtimes \mathcal G_2\) & \verb|G1| $\rtimes$ \verb|G2|\\
        \midrule
        \verb|SpecialEuclideanGroup(n)| & \(\mathrm{SE}(n)\) & \\
        \verb|SpecialGalileanGroup(n)| & \(\mathrm{SGal}(n)\) &  \\
        \verb|SpecialLinearGroup(n, F)| & \(\mathrm{SL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
        \verb|SpecialOrthogonalGroup(n)| & \(\mathrm{SO}(n)\) &  \\
        \verb|SpecialUnitaryGroup(n)| & \(\mathrm{SU}(n)\) & \\
        \midrule
        \verb|SymplecticGroup(n)| & \(\mathrm{Sp}(2n)\) & \\
        \verb|TranslationGroup(n; field=|$\mathbb{F}$\verb|)| & \(\mathbb{F}^n\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}, \mathbb{H}\}$\\
        \verb|UnitaryGroup(n)| & \(\mathrm{U}(n)\) & \\
        \bottomrule
    \end{tabular}
\end{table}

\section{An example how to use {\texttt{LieGroups.jl}}}

\cite{LieGroups.jl}
\input{bib.tex}
\end{document}
%
% ------------------------------------------------------------------------------------------
%
\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb|\ref| ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb|juliacon.cls|  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb|juliacon|  document class is implemented
as a complete class, not a document style option. In order to
use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
\verb|\documentclass|  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb|\documentclass{article}|  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb|\documentclass{juliacon}|  \vskip 6pt
In general, the following standard document \verb|style|  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb|juliacon|  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
\verb|\def|  and \verb|\newcommand|   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb|juliacon.cls|  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
In general, once you have used the additional \verb|juliacon.cls| facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:exmplg},
    caption={Example Code Block.}
]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:exmplg},
    caption={Example Code Block.}
]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
\verb|\terms|,
\verb|\keywords|  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb|juliacon|  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb|juliacon|  class file:
\begin{itemize}
\item \verb|\section|
\item \verb|\subsection|
\item \verb|\subsubsection|
\item \verb|\paragraph|
\end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb|juliacon|   class file provides unnumbered lists using the
\verb|unnumlist|  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb|juliacon|  class file also provides hyphen list using the
\verb|itemize|  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb|juliacon|  class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb|figure|  environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb|\caption|  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb|\includegraphics|   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb|Figure~\ref{sample-figure}|

 \subsection{Tables}
\label{subsub:sec_Tab}
The \verb|juliacon|   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb|juliacon|   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb|rotating|   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb|juliacon| class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb|juliacon|   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

\input{bib.tex}

\end{document}

% Inspired by the International Journal of Computer Applications template

### 2025-10-26 paper/paper.tex
Start structuring the paper and start the introduction.

@@
 
+In many situations, one encounters data that does not reside in a vector space.
+For example in robotics, the configuration space of a rigid body in three-dimensional
+space is given by the special Euclidean group \(\mathrm{SE}(3)\),
+consisiting of all translations and rotations.
+A subset of these is the space of rotations, given by the special orthogonal group \(\mathrm{SO}(3)\),
+or more generally \(\mathrm{SO}(n)\) in \(n\)-dimensional space.
+
+These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
+
+The package \verb|LieGroups.jl| aims to provide an easy access to using Lie groups within
+the Julia programming language \cite{bezanson2017julia} by providing an interface to define
+and work with Lie groups, as well as providing a library of Lie groups, that can directly be used.
+
+This paper aims to provide an overview of the main features of \verb|LieGroups.jl|.
+After introducing the necessary notation in Section~\ref{sec:Notation},
+we present the interface in Section~\ref{sec:Interface}, split into the Lie group, the Lie algebra, and group actions.
+In Section~\ref{sec:LieGroups}, we present an overview of currently implemented Lie groups.
+Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups|.
+
+\section{Notation}\label{sec:Notation}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.1875\textwidth]{logo.png}
+    \caption{Logo of \texttt{LieGroups.jl}.}
+    \label{fig:liegroups_logo}
+\end{figure}
+
+\section{The interface}\label{sec:Interface}
+
 \cite{Bergmann:2022}, \cite{AxenBaranBergmannRzecki:2023}.
 
-\section{Notation}
-
-\section{The interface}
 \subsection{Lie groups}
@@
 
-\section{Implemented Lie groups}
+\section{Implemented Lie groups}\label{sec:LieGroups}
 \begin{table}[tbp]
@@
 
-\section{An example how to use {\texttt{LieGroups.jl}}}
+\section{An example how to use {\texttt{LieGroups.jl}}}\label{sec:Example}
 

### 2025-10-26 paper/header.tex
Write about the mathematical background.

@@
 
-\keywords{Julia, Riemannian manifolds, Lie groups, Differential geometry}
+\keywords{Julia, Riemannian manifolds, Lie groups, differential geometry, numerical analysis, scientific computing}
 
@@
 pdfauthor = {Ronny Bergmann},
-pdfkeywords = {Julia, Riemannian manifolds, Lie groups, Differential geometry},
+pdfkeywords = {Julia, Riemannian manifolds, Lie groups, differential geometry, numerical analysis, scientific computing},
 }

### 2025-10-26 paper/paper.tex
Write about the mathematical background.

--- 2025-10-25 paper/paper.tex
@@
 \setcounter{page}{1}
-\usepackage{booktabs}
-
+\usepackage{amsmath,booktabs,mathtools,amssymb}
+\newcommand{\term}[1]{\emph{#1}}
+%
 \hypersetup{colorlinks=true}
-
+%
 \begin{document}
@@
 \input{header}
-
+% TODO : How to specify an email-address?
 \maketitle
@@
     \href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{\texttt{LieGroups.jl}} is a
-    Julia package that provides both an interface to define as well as work with Lie groups,
-    the corresponding Lie algebra as well as group actions.
+    Julia package that provides on the one hand an interface to define Lie groups as well as the corresponding Lie algebra and Lie group actions.
+On the other hand, it offers a well-documented, performant, and well-tested library of existing Lie groups, their Algebra and corresponding group actions.
 
@@
     the \verb|JuliaManifolds| ecosystem. We further present an overview on existing Lie groups
-    implemented in \texttt{LieGroups.jl} as well as how to get started to use the package in practice.
+    implemented in \verb|LieGroups.jl| as well as how to get started to use the package in practice.
 \end{abstract}
+%
+\section{Introduction}%
+\label{sec:Introduction}
+%
+In many situations, one encounters data that does not reside in a vector space.
+We can hence not used standard linear algebra tools to work with such data.
+For example in robotics, the configuration space of a rigid body in three-dimensional
+space is given by the special Euclidean group \(\mathrm{SE}(3)\),
+consisting of all translations and rotations.
+A subset of these is the space of rotations, given by the special orthogonal group \(\mathrm{SO}(3)\),
+or more generally \(\mathrm{SO}(n)\) in \(n\)-dimensional space.
 
-\section{Introduction}
+These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
+They have applications e.g.\ in physics, robotics, stochastic processes, information geometry, see e.g.~\cite{Chirikjian:2009,Chirikjian:2012},
+but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
+and their numerical aspects, e.g.\ when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998}.
 
-\cite{Bergmann:2022}, \cite{AxenBaranBergmannRzecki:2023}.
+The package~\href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{\texttt{LieGroups.jl}}\footnote{%
+Available at~\href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{juliamanifolds.github.io/LieGroups.jl/stable/},
+see also the zenodo archive~\cite{LieGroups.jl}.%
+}
+provides an easy access to both defining and using Lie groups within
+the Julia programming language~\cite{bezanson2017julia} by defining an interface of Lie groups, as well as implementing a library of Lie groups, that can directly be used.
 
-\section{Notation}
+This paper provides an overview of the main features of \verb|LieGroups.jl|.
+After introducing the necessary mathematical background and notation in Section~\ref{sec:Notation},
+we present the interface in Section~\ref{sec:Interface}, split into the Lie group, the Lie algebra, and group actions.
+In Section~\ref{sec:LieGroups}, we present an overview of currently implemented Lie groups.
+Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups.jl|.
 
-\section{The interface}
+\section{Mathematical Background}%
+\label{sec:Notation}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.1875\textwidth]{logo.png}
+    \caption*{\ \\[-.5\baselineskip]Logo of \texttt{LieGroups.jl}.}%
+    \label{fig:liegroups_logo}
+\end{figure}
+
+The following notation and definitions follow the text book~\cite{HilgertNeeb:2012},
+for more details on Riemannian manifolds, see also~\cite{DoCarmo:1992}.
+
+We denote a Lie group by \(\mathcal{G} = (\mathcal{M}, \cdot)\) where \(\mathcal{M}\) is a smooth manifold and \(\cdot\) is the group operation.
+A smooth manifold \(\mathcal{M}\) is a topological space that is locally isomorphic to an Euclidean space \(\mathbb{R}^n\) for some \(n \in \mathbb{N}\), but globally may have a different topology.
+We call \(n\) the dimension of the manifold \(\mathcal{M}\), denoted by \(\dim(\mathcal{M}) = n\).
+As an example, take the \(2\)-dimensional sphere \(\mathbb{S}^2 = \{x \in \mathbb{R}^3 \mid \lVert x\rVert = 1\}\),
+which locally looks like \(\mathbb{R}^2\), think of charts in an atlas, but globally it is not.
+Finally we denote the tangent space at a point \(p \in \mathcal{M}\) by \(T_p\mathcal{M}\).
+This can be thought of as all “velocities” (direction and speed) in which a curve can “pass through” a point.
+Formally it is set the equivalence classes of derivatives of smooth curves.
+Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces $T\mathcal M = \dot\bigsqcup_{p \in \mathcal M} T_p\mathcal M$ the \term{tangent bundle} of $\mathcal M$.
+
+As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(x^{-1} \in \mathcal{G}\) for all \(x \in \mathcal{G}\).
+Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, x \mapsto x^{-1}\) have to be smooth maps.
+As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), i.e.\ for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
+For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$, or in other words the circle. % chktex 9ok
+The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose (or an angle $-\alpha$).
+
+The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
+The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$, since we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
+Then, using the differential (or pushforward) $D\lambda_g(h)\colon T_g\mathcal G \to T_h\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.
+Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
+
+Finally, a \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\rho\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\rho(e, p) = p\) and \(\rho(g, \rho(h, p)) = \rho(g \cdot h, p)\).
+Informally a group action describes how elements of the Lie group \(\mathcal{G}\) “act on” points on the manifold \(\mathcal{M}\).
+As an example, think of the special orthogonal group \(\mathrm{SO}(3)\) acting on points on Euclidean space $\mathbb{R}^3$ “moving” them somewhere by rotating them around the origin.
+The same action can also be applied to points from the sphere \(\mathbb{S}^2\).
+
+\section{The interface}\label{sec:Interface}
+
+Since a Lie group \(\mathcal{G}\) consists of two main components, the smooth manifold \(\mathcal{M}\) and the group operation \(\cdot\), we can reuse existing functionality from the existing interface for manifolds provided by \verb|ManifoldsBase.jl|, and later concrete manifolds provided by \verb|Manifolds.jl|~\cite{AxenBaranBergmannRzecki:2023}.
+This is done in a transparent way, i.e.\ the \verb|AbstractLieGroup| itself is a subtype of \verb|AbstractManifold| from \verb|ManifoldsBase.jl| and can hence also be used in all existing places, as for example optimization on manifolds provided by \verb|Manopt.jl|~\cite{Bergmann:2022}.
+
 \subsection{Lie groups}
+{\color{red} TODO 1-1.5 pages on the interfaces and main functions provided for Lie groups and group operations, Lie algebra, and group actions. Maybe take over some story parts from the talk.}
 
@@
         \verb|GeneralLinearGroup(n, F)| & \(\mathrm{GL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
-        \verb|HeisenbergGroup(n)| & \verb|H(n)| \\
+        \verb|HeisenbergGroup(n)| & \(\mathrm{H}(n)\)\\
         \verb|OrthogonalGroup(n)| & \(\mathrm{O}(n)\) &\\
@@
 
-\section{An example how to use {\texttt{LieGroups.jl}}}
-
-\cite{LieGroups.jl}
+\section{An example how to use {\texttt{LieGroups.jl}}}\label{sec:Example}
 \input{bib.tex}
@@
 %
-\section{Introduction}
-
-The \LaTeX{} document preparation system is a special version of the
-\TeX; typesetting program where-in a collection of \TeX; commands
-are added to \LaTeX{} to simplify typesetting. Importantly, it allows
-the author to concentrate on the logical structure of the document
-rather than its visual layout.\vskip 6pt
-Moreover, \LaTeX{} provides a consistent and comprehensive document
-preparation interface. There are simple-to-use commands for
-generating a table of contents, lists of figures and/or tables, and indexes.
-\LaTeX{} can automatically number list entries, equations, figures,
-tables, and footnotes, as well as articles, sections, and subsections.
-Using this numbering system, bibliographic citations, page
-references, and cross references to any other numbered entity (e.g.
-article, section, equation, figure, list entry, etc.) become quite simple
-and straightforward. The use of \LaTeX{} document classes allows
-a simple change of class to transform the appearance of your document.\vskip 6pt
-\LaTeX{} is a powerful tool for managing long and complex documents.
-In particular, partial processing enables long documents to
-be produced article by article without losing sequential information.
-The use of document classes allows a simple change of style
-(or style option) to transform the appearance of your document.
-
-\section{The JuliaCon Article Class}
-\label{sec:documentclass}
-%
-The juliacon class file preserves the standard LATEX{} interface such
-that any document that can be produced using the standard LATEX{}
-article class can also be produced with the class file.\vskip 6pt
-It is likely that the make up will change after file submission. For
-this reason, we ask you to ignore details such as slightly long lines,
-page stretching, or figures falling out of synchronization, as these
-details can be dealt with at a later stage.\vskip 6pt
-Use should be made of symbolic references (\verb|\ref| ) in order to
-protect against late changes of order, etc.
-
-\section{USING THE JuliaCon Article CLASS FILE}
-
-If the file \verb|juliacon.cls|  is not already in the appropriate system directory
-for \LaTeX{} files, either arrange for it to be put there or copy
-it to your working directory. The \verb|juliacon|  document class is implemented
-as a complete class, not a document style option. In order to
-use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
-\verb|\documentclass|  command at the beginning of your document:
-\vskip 6pt
-\begin{centering}
-    \verb|\documentclass{article}|  \end{centering}
-\vskip 6pt
-replace by
-\vskip 6pt
- \verb|\documentclass{juliacon}|  \vskip 6pt
-In general, the following standard document \verb|style|  options should
-{ \itshape not} be used with the {\footnotesize \itshape article} class file:
-\begin{enumerate}
-\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
-\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
-\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
-\end{enumerate}
-
-\section{Additional Document Style Options}
-\label{sec:additional_doc}
-%
-The following additional style option is available with the \verb|juliacon|  class file:
-\vskip 6pt
-Please place any additional command definitions at the very start of
-the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
-\verb|\def|  and \verb|\newcommand|   commands that define macros for
-technical expressions should be placed here. Other author-defined
-macros should be kept to a minimum.
-\vskip 6pt
-Commands that differ from the standard \LaTeX{} interface, or that
-are provided in addition to the standard interface, are explained in
-this guide. This guide is not a substitute for the \LaTeX{} manual itself.
-Authors planning to submit their papers in \LaTeX{} are advised to use
-\verb|juliacon.cls|  as early as possible in the creation of their files.
-
-%
-%
-%
-%
-\begin{table*}[t]
-\tabcolsep22pt
-\tbl{If necessary, the tables can be extended both columns.}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\label{tab:symbols}
-\begin{tabnote}
-This is an example of table footnote.
-\end{tabnote}
-\end{table*}
-% \begin{figure*}[t]
-% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
-% \caption{If necessary, the images can be extended both columns.}
-%   \label{fig:sample_image}
-% \end{figure*}
-
-\section{Additional features}
-\label{sec:additional_faci}
-In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
-In general, once you have used the additional \verb|juliacon.cls| facilities
-in your document, do not process it with a standard \LaTeX{} class
-file.
-
-\subsection{Titles, Author's Name, and Affiliation}
-\label{subsub:title_auth}
-The title of the article, author's name, and affiliation are used at the
-beginning of the article (for the main title). These can be produced
-using the following code:
-
-\begin{verbatim}
-\title{ This is an example of article title} }
-\author{
-   \large 1st Author \\[-3pt]
-   \normalsize 1st author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	1st author's email address \\[-3pt]
-  \and
-   \large 2nd Author \\[-3pt]
-   \normalsize 2nd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	2nd author's email address \\[-3pt]
-\and
-   \large 3rd Author \\[-3pt]
-   \normalsize 3rd author's affiliation  \\[-3pt]
-    \normalsize 1st line of address \\[-3pt]
-    \normalsize 2nd line of address \\[-3pt]
-    \normalsize	3rd author's email address \\[-3pt]
-}
-\maketitle
-\end{verbatim}
 
@@
 
-\subsection{Abstracts, Key words, term etc...}
-\label{subsub:abs_key_etc}
-
-At the beginning of your article, the title should be generated
-in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
-\verb|\terms|,
-\verb|\keywords|  commands respectively. The abstract should be enclosed
-within an abstract environment, All these environment
-can be produced using the following code:
-\begin{verbatim}
-\terms{Experimentation, Human Factors}
-
-\keywords{Face animation, image-based modelling...}
-
-\begin{abstract}
-In this paper, we propose a new method for the
-systematic determination of the model's base of
-time varying delay system. This method based on
-the construction of the classification data related
-to the considered system. The number, the orders,
-the time delay and the parameters of the local
-models are generated automatically without any
-knowledge about the full operating range of the
-process. The parametric identification of the local
-models is realized by a new recursive algorithm for
-on line identification of systems with unknown time
-delay. The proposed algorithm allows simultaneous
-estimation of time delay and parameters of
-discrete-time systems. The effectiveness of
-the new method has been illustrated through
-simulation.
-\end{abstract}
-
-\end{verbatim}
-
-\section{Some guidelines}
-\label{sec:some_guide}
-The following notes may help you achieve the best effects with the
-\verb|juliacon|  class file.
-
-\subsection{Sections}
-\label{subsub:sections}
-\LaTeXe{}  provides four levels of section headings and they are all
-defined in the \verb|juliacon|  class file:
-\begin{itemize}
-\item \verb|\section|
-\item \verb|\subsection|
-\item \verb|\subsubsection|
-\item \verb|\paragraph|
-\end{itemize}
-Section headings are automatically converted to allcaps style.
-\subsection{Lists}
-\label{sec:lists}
-%
-The \verb|juliacon|   class file provides unnumbered lists using the
-\verb|unnumlist|  environment for example,
-
-\begin{unnumlist}
-\item First unnumbered item which has no label and is indented from the
-left margin.
-\item Second unnumbered item.
-\item Third unnumbered item.
-\end{unnumlist}
-The unnumbered list which has no label and is indented from the
-left margin. was produced by:
-\begin{verbatim}
-\begin{unnumlist}
-\item First unnumbered item...
-\item Second unnumbered item...
-\item Third unnumbered item...
-\end{unnumlist}
-\end{verbatim}
-
-The \verb|juliacon|  class file also provides hyphen list using the
-\verb|itemize|  environment for example,
-\begin{itemize}
-\item First unnumbered bulleted item which has no label and is indented
-from the left margin.
-\item Second unnumbered bulleted item.
-\item Third unnumbered bulleted item which has no label and is indented
-from the left margin.
-\end{itemize}
-was produced by:
-\begin{verbatim}
-\begin{itemize}
-\item First item...
-\item Second item...
-\item Third item...
-\end{itemize}
-\end{verbatim}
-
-Numbered list is also provided in acmtog class file using the
-enumerate environment for example,
-\begin{enumerate}
-\item The attenuated and diluted stellar radiation.
-\item Scattered radiation, and
-\item Reradiation from other grains.
-\end{enumerate}
-
-was produced by:
-\begin{verbatim}
-\begin{enumerate}
-\item The attenuated...
-\item Scattered radiation, and...
-\item Reradiation from other grains...
-\end{enumerate}
-\end{verbatim}
-\subsection{Illustrations (or figures)}
-\label{subsub:sec_Illus}
-The \verb|juliacon|  class file will cope with most of the positioning of
-your illustrations and you should not normally use the optional positional
-qualifiers on the \verb|figure|  environment that would override
-these decisions.
-\vskip 6pt
-
-%
-\begin{figure}[t]
-\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
-\caption{This is example of the image in a column.}
-	\label{fig:sample_figure}
-\end{figure}
-
-The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
-organization \footnote{https://github.com/JuliaGraphs}.
-
-Figure captions should be \emph{below} the figure itself, therefore the
-\verb|\caption|  command should appear after the figure or space left for
-an illustration. For example, Figure 1 is produced using the following
-commands:
-
-\begin{verbatim}
-\begin{figure}
-\centerline{\includegraphics[width=20pc]{Graphics.eps}}
-\caption{An example of the testing process for a
-binary tree. The globa null hypothesis is tested
-first at level $\alpha$ (a), and the level of
-individual variables is reached last (d). Note
-that individual hypotheses can be tested at
-level $\alpha/4$ and not $\alpha/8$ as one might
-expect at first.}
-\label{sample-figure_2}
-\end{figure}
-\end{verbatim}
-Figures can be resized using first and second argument of
-\verb|\includegraphics|   command. First argument is used for modifying
-figure height and the second argument is used for modifying
-figure width respectively.
-\vskip 6pt
-Cross-referencing of figures, tables, and numbered, displayed
-equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
-For example, in referencing Figure 1 above, we used
-\verb|Figure~\ref{sample-figure}|
-
- \subsection{Tables}
-\label{subsub:sec_Tab}
-The \verb|juliacon|   class file will cope with most of the positioning of
-your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
-decisions. Table captions should be at the top.
-\begin{verbatim}
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Train70 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test70 & Testing Data I & & 105\\\hline
-Test30 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\end{verbatim}
-
-\begin{table}
-\tbl{Tuning Set and Testing Set}{
-\begin{tabular}{|l|l|c|c|}\hline
-Label & \multicolumn{1}{c|}{Description}
-& Number of Users &
-Number of Queries\\\hline
-Test 1 & Training Data &
-\smash{\raise-7pt\hbox{70}} & 104\\
-\cline{1-2}\cline{4-4}
-Test 2 & Testing Data I & & 105\\\hline
-Test 3 & Testing Data II & 30 & 119\\\hline
-& Total & 100 & 328\\\hline
-\end{tabular}}
-\end{table}
-\subsection{Landscaping Pages}
-\label{subsub:landscaping_pages}
-If a table is too wide to fit the standard measure, it may be turned,
-with its caption, to 90 degrees. Landscape tables cannot be produced
-directly using the \verb|juliacon|   class file because \TeX{} itself cannot
-turn the page, and not all device drivers provide such a facility.
-The following procedure can be used to produce such pages.
-\vskip 6pt
-Use the package \verb|rotating|   in your document and change the coding
-from
-\begin{verbatim}
-\begin{table}...\end{table}
-to
-\begin{sidewaystable}...\end{sidewaystable}
-and for figures
-\begin{figure}...\end{figure}
-to
-\begin{sidewaysfigure}...\end{sidewaysfigure}
-\end{verbatim}
-
-environments in your document to turn your table on the appropriate
-page of your document. For instance, the following code prints
-a page with the running head, a message half way down and the
-table number towards the bottom.
-\begin{verbatim}
-\begin{sidewaystable}
-\tbl{Landscape table caption to go here.}{...}
-\label{landtab}
-\end{sidewaystable}
-\end{verbatim}
-
-\subsection{Double Column Figure and Tables}
-\label{subsub:double_fig_tab}
-For generating the output of figures and tables in double column
-we can use the following coding:
-
-\begin{enumerate}
-\item For Figures:
-\begin{verbatim}
-\begin{figure*}...\end{figure*}
-\end{verbatim}
-\item For landscape figures:
-\begin{verbatim}
-\begin{sidewaysfigure*}...\end{sidewaysfigure*}
-\end{verbatim}
-\item For Tables:
-\begin{verbatim}
-\begin{table*}...\end{table*}
-\end{verbatim}
-\item For landscape tables:
-\begin{verbatim}
-\begin{sidewaystable*}...\end{sidewaystable*}
-\end{verbatim}
-\end{enumerate}
-
-\subsection{Typesetting Mathematics}
-\label{subsub:type_math}
-The \verb|juliacon| class file will set displayed mathematics with center to
-the column width, provided that you use the \LaTeXe{} standard of
-open and closed square brackets as delimiters.
-The equation
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-
-was typeset using the acmtog class file with the commands
-
-\begin{verbatim}
-\[
-\sum_{i=1}^p \lambda_i = (S)
-\]
-\end{verbatim}
-
-For display equations, cross-referencing is encouraged. For example,
-\begin{verbatim}
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-\end{verbatim}
-The following output is generated with the above coding:
-\begin{equation}
-(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
-\label{eq:samplevar}
-\end{equation}
-Equation~(\ref{eq:samplevar}) gives the formula for
-sample variance.
-
-\subsection{Enunciations}
-\label{subsub:enunciation}
-The \verb|juliacon|   class file generates the enunciations with the help of
-the following commands:
-\begin{verbatim}
-\begin{theorem}...\end{theorem}
-\begin{strategy}...\end{strategy}
-\begin{property}...\end{property}
-\begin{proposition}...\end{proposition}
-\begin{lemma}...\end{lemma}
-\begin{example}...\end{example}
-\begin{proof}...\end{proof}
-\begin{definition}...\end{definition}
-\begin{algorithm}...\end{algorithm}
-\begin{remark}...\end{remark}
-\end{verbatim}
-The above-mentioned coding can also include optional arguments
-such as
-\begin{verbatim}
-\begin{theorem}[...]. Example for theorem:
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven ... created equal.
-\end{theorem}
-\end{verbatim}
-
-\begin{theorem}[Generalized Poincare Conjecture]
-Four score and seven years ago our fathers brought forth,
-upon this continent, a new nation, conceived in Liberty,
- and dedicated to the proposition that all men are
-created equal.
-\end{theorem}
-
-\subsection{Extract}
-\label{subsub:extract}
-Extract environment should be coded within
-\begin{verbatim}
-\begin{extract}..\end{extract}
-\end{verbatim}
-
 \subsection{Balancing column at last page}
@@
 at appropriate place in your \TeX{} file or in bibliography file.
-
-\section{Handling references}
-\label{subsub:references}
-References are most easily (and correctly) generated using the
-BIBTEX, which is easily invoked via
-\begin{verbatim}
-\bibliographystyle{juliacon}
-\bibliography{ref}
-\end{verbatim}
-When submitting the document source (.tex) file to external
-parties, the ref.bib file should be sent with it.
-\cite{bezanson2017julia}
-
-\input{bib.tex}
-
-\end{document}
-
-% Inspired by the International Journal of Computer Applications template

### 2025-10-27 paper/paper.tex
Finish first full draft of section 3.1

@@
 \newcommand{\term}[1]{\emph{#1}}
+\newcommand{\e}{\ensuremath{\mathrm{e}}}
+\newcommand{\im}{\ensuremath{\mathrm{i}}}
+\newcommand{\LieGroupsVersion}{0.1.6}
 %
@@
 
-This paper provides an overview of the main features of \verb|LieGroups.jl|.
+This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}.
 After introducing the necessary mathematical background and notation in Section~\ref{sec:Notation},
@@
     \centering
-    \includegraphics[width=0.1875\textwidth]{logo.png}
+    \includegraphics[width=0.17\textwidth]{logo.png}
     \caption*{\ \\[-.5\baselineskip]Logo of \texttt{LieGroups.jl}.}%
@@
 
+\subsection{Lie groups}
 We denote a Lie group by \(\mathcal{G} = (\mathcal{M}, \cdot)\) where \(\mathcal{M}\) is a smooth manifold and \(\cdot\) is the group operation.
@@
 We call \(n\) the dimension of the manifold \(\mathcal{M}\), denoted by \(\dim(\mathcal{M}) = n\).
-As an example, take the \(2\)-dimensional sphere \(\mathbb{S}^2 = \{x \in \mathbb{R}^3 \mid \lVert x\rVert = 1\}\),
+As an example, take the \(2\)-dimensional sphere
+\begin{equation*}
+\mathbb{S}^2 = \{x \in \mathbb{R}^3 \mid \lVert x\rVert = 1\},
+\end{equation*}
+
 which locally looks like \(\mathbb{R}^2\), think of charts in an atlas, but globally it is not.
@@
 Formally it is set the equivalence classes of derivatives of smooth curves.
-Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces $T\mathcal M = \dot\bigsqcup_{p \in \mathcal M} T_p\mathcal M$ the \term{tangent bundle} of $\mathcal M$.
-
-As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(x^{-1} \in \mathcal{G}\) for all \(x \in \mathcal{G}\).
-Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, x \mapsto x^{-1}\) have to be smooth maps.
+Each tangent space $T_p\mathcal M$ is a $n$-dimensional vector space and we call the disjoint union of all tangent spaces
+\begin{equation*}
+    T\mathcal M = \dot\bigcup_{p \in \mathcal M} T_p\mathcal M
+\end{equation*}
+the \term{tangent bundle} of $\mathcal M$.
+\\
+As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
+Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
 As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), i.e.\ for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
@@
 
-The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
+\subsection{Tangent spaces and the Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
 The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$, since we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
-Then, using the differential (or pushforward) $D\lambda_g(h)\colon T_g\mathcal G \to T_h\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.
+Then, using the differential (or pushforward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
+\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields.
+\texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
 Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
+\\
+As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$.
+We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
+Here, a tangent vector $X$ induces the constrant vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
+As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
+The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
+For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$.
+Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
+In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \frak{g}$.
 
-Finally, a \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\rho\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\rho(e, p) = p\) and \(\rho(g, \rho(h, p)) = \rho(g \cdot h, p)\).
+Concerning a metric on the tangent spaces, manifolds are turned into Riemannian manifolds when they are equipped with a Riemannian metric \(\langle \cdot, \cdot \rangle_p\colon T_p\mathcal{M} \times T_p\mathcal{M} \to \mathbb{R}\) for each point \(p \in \mathcal{M}\) that smoothly varies with \(p\).
+For Lie groups, we can use a single inner product \(\langle \cdot, \cdot \rangle\) on the Lie algebra \(\mathfrak{g}\) and use the change in representation as mentioned above to define
+\begin{equation*}
+\langle X, Y \rangle_g = \langle D\lambda_{g^{-1}}(g)[X], D\lambda_{g^{-1}}(g)[Y] \rangle_e
+\end{equation*}
+for all \(X, Y \in T_g\mathcal{G}\) and \(g \in \mathcal{G}\).
+Representing tangent vectors at arbitrary points \(g \in \mathcal{G}\) using the Lie algebra \(\mathfrak{g}\) yields here, that we can use the single inner product directly to evaluate this Riemannian metric.
+Hence representing tangent vectors using the Lie algebra is the default in the following.
+\\
+For the two examples above we obtain the Euclidean inner product \(\langle X, Y \rangle = X^{\mathrm{T}} Y\) for \(X, Y \in \mathbb{R}^n\) for \((\mathbb{R}^n, +)\) and the Frobenius inner product \(\langle X, Y \rangle = \mathrm{trace}(X^{\mathrm{T}} Y)\), \(X, Y \in \mathfrak{so}(n)\) can be used for \(\mathrm{SO}(n)\).
+
+\subsection{Group Actions} A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).
 Informally a group action describes how elements of the Lie group \(\mathcal{G}\) “act on” points on the manifold \(\mathcal{M}\).
 As an example, think of the special orthogonal group \(\mathrm{SO}(3)\) acting on points on Euclidean space $\mathbb{R}^3$ “moving” them somewhere by rotating them around the origin.
-The same action can also be applied to points from the sphere \(\mathbb{S}^2\).
+We obtain the group action \(\sigma\colon \mathrm{SO}(3) \times \mathbb{R}^3 \to \mathbb{R}^3\), \(\sigma(R, x) \coloneqq Rx\).
+
+The same action can also be applied to points from the sphere \(\mathbb{S}^2\), resulting in a similar group action \(\sigma\colon \mathbb{SO}(3) \times \mathbb{S}^2 \to \mathbb{S}^2\), \(\sigma(R, p) \coloneqq Rp\).
 
@@
 This is done in a transparent way, i.e.\ the \verb|AbstractLieGroup| itself is a subtype of \verb|AbstractManifold| from \verb|ManifoldsBase.jl| and can hence also be used in all existing places, as for example optimization on manifolds provided by \verb|Manopt.jl|~\cite{Bergmann:2022}.
+In notation, we use typewriter font to denote functions in the interface, but we keep the same letters of notation as before, i.e.\ $\mathcal G$ for a Lie group is \verb|G| in code, a point $g \in \mathcal G$ is \verb|g| in code, and so on, just that for the Lie algebra we use $\mathfrak g$ in both text and code.
+
+In general, the interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
+Similarly, if a function computes something like a new point on the Lie group or a tangent vector, there also exists a variant, that computes this in-place.
+For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default represention, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.
 
 \subsection{Lie groups}
-{\color{red} TODO 1-1.5 pages on the interfaces and main functions provided for Lie groups and group operations, Lie algebra, and group actions. Maybe take over some story parts from the talk.}
 
-\cite{HilgertNeeb:2012}
+The main type for Lie Groups is the \verb|LieGroup{|$\mathbb {F}$\verb|, O, M} <: AbstractLieGroup|, which contains a manifold \verb|M <: AbstractManifold{|$\mathbb F$\verb|}| as well as the group operation \verb| O <: AbstractGroupOperation|, where $\mathbb {F}$ is the field of scalars used in the representation of the manifold, usually \(\mathbb {F} = \mathbb{R}\) and in some cases \(\mathbb {F} = \mathbb{C}\) or \(\mathbb F = \mathbb H\).
+
+\subsubsection*{Topological functions}
+From a topological viewpoint, one has to first distinguish between points on the Lie group and tangent vectors on the Lie algebra.
+\verb|LieGroups.jl| intrduces a \verb|AbstractLieGroupPoint <: AbstractManifoldPoint|, that is used to represent points on the Lie group.
+This abstract supertype is not necessary, points can be represented by arrays or other types as well, but the type allows to intrduce special types when one Lie groups has \emph{different representations} of points, that have to be distinguished.
+When there are different representations, it is recommended to introdcude a point type for each representation and make the default one fall back to (just) using arrays.
+\\
+Alternatively to points, also different manifolds can be used.
+This is for example the case for the circle group \(\mathbb{S}^1\), which can be represented as angles, points on the unit circle in \(\mathbb{R}^2\) or as complex numbers with unit norm, all three of which are different manifolds to be used internally.
+To access the underlying manifold, the function \verb|base_manifold(G)| is provided.
+
+Beyond that, the following functions are available and pass directly to the underlying manifold interface from \verb|ManifoldsBase.jl|:
+\verb|is_approx(G, g, h)| to check for (approximate) equality of two points,
+\verb|is_point(G, g)| to check if a point is a valid point on the manifold,
+\verb|manifold_dimension(G)| to get the dimension of the manifold,
+\verb|project(G, q)|\footnote{in-place variant \texttt{project!(G, p, q)}.}
+to project a point onto the manifold,
+\verb|rand(G)| and \verb|rand(rng, G)|\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
+to sample a random point from the manifold using a random number generator \verb|rng|.
+
+\subsubsection*{Group operation related functions}
+
+\begin{table}
+    \centering
+    \caption{Group operations available in \texttt{LieGroups.jl}.}
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Group operation} $\cdot$ & comment/code\\
+        \midrule
+        \verb|AdditionGroupOperation| & falls back to \verb|+|\\
+        \verb|AbelianMultiplicationGroupOperation| & falls back to \verb|*|\\
+        \verb|LeftMatrixMultiplicationGroupOperation| & falls back to \verb|*|\\
+        \verb|PowerGroupOperation{Op}| & \verb|PowerLieGroup|\\
+        \verb|ProductGroupOperation{Ops}| & product Lie groups\\
+        \verb|SemidirectProductGroupOperation| & semidirect products\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:GroupOperations}
+\end{table}
+
+For the group operation \(\cdot\) of a Lie group \(\mathcal{G}\), the abstract supertype \verb|AbstractGroupOperation| is mandatory.
+There are two main group operation types provided in \verb|LieGroups.jl|.
+On the one hand operations that fall back to using $\cdot = +$ or $\cdot = *$, where the latter has two variants, the one where it is Abelian (like for numbers) and where it is not (like for matrices).
+On the other hand, specific meta groups like the (direct) product of two or more groups, the case where the product is taken for just one group, i.e.\ the power group, or semidirect products.
+The available group operation types are summarized in Table~\ref{tab:GroupOperations}.
+More on these meta groups is explained in Subsection~\ref{subsec:MetaLieGroups}.
+
+For all these group operations, the following functions have default implementations. They might not be the most performant ones for every case, but provide working implementations out-of-the-box.
+It is always possible to override these by defining a new group operation type and implementing the following functions for a \verb|LieGroup| with that new group operation type.
+
+To avoid unnecessary allocations, the identity element can be used without allocations, calling \verb|e = IdentityElement(G)| or \verb|IdentityElement(op)| where \verb|op| is the group operation of \verb|G|.
+If the actual value is needed, one can call
+
+\verb|identity_element(G)| or \verb|identity_element(G, typeof(g))|, where \verb|g| is an element of the group.
+The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|.\footnote{The in-place variant \texttt{identity\_element!(G, e)} infers the representation from the type of \texttt{e}.}
+
+The two central functions are \verb|compose(G, g, h)|\footnote{in-place variant \texttt{compose!(G, k, g, h)}.} to compute the group operation \(g \cdot h\) for two points \(g, h \in \mathcal{G}\) and the inverse \verb|inv(G, g)|\footnote{in-place variant \texttt{inv!(G, h, g)}.} which computes \(g^{-1}\) for a point \(g \in \mathcal{G}\).
+
+Another function is $c_g(h) = g \cdot h \cdot g^{-1}$ called \term{conjugation}, which is available as \verb|conjugate(G, g, h)|\footnote{in-place variant \texttt{conjugate!(G, k, g, h)}.}.
+
+For these three functions, also the differentials are available: adding a \verb|diff_| prefix to the function name and a final argument for the Lie algebra tangent vector, for compose additionally the argument with which to  differentiate to, i.e.\ \verb|diff_left_compose(G, g, h, X)| and \verb|diff_right_compose(G, g, h, X)|, respectively.\footnote{in-place variants have a second argument \texttt{Y} to compute the result in.}
+
+Finally, there is a specific function for the differential of the conjugate at the identity $h=e$ called \verb|adjoint(G, g, X)|\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.} as well as the combinations \verb|inv_left_compose(G, g, h)|\footnote{in-place variant \texttt{inv\_left\_compose!(G, k, g, h)}} for computing $g^{-1}\cdot h$ and \verb|inv_right_compose(G, g, h)|\footnote{in-place variant \texttt{inv\_right\_compose!(G, k, g, h)}} for computing $g \cdot h^{-1}$.
+All three fall back to the previusly defined functions, but provide an interface to possibly implement more efficient variants in case such exist.
+
+\subsubsection*{exponential and logarithm.}
+An important tool to “move around” on the Lie group is the exponential map.\\
+The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfills the differential equation
+\begin{equation*}
+\gamma'(t) = D\lambda_{\gamma(t)}(e)[X]
+\quad\text{ with }\quad\gamma(0) = e \text{ and }\gamma'(0) = X
+\end{equation*}
+at time $t=1$, i.e.\ \(\exp(X) = \gamma(1)\).
+
+This can be interpreted as “starting at the identity element \(e \in \mathcal{G}\) and following the curve whose velocity at each point is given by the left-invariant vector field induced by \(X \in \mathfrak{g}\) for one time unit”.
+
+In our examples, for \((\mathbb{R}^n, +)\) the exponential map is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
+Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential map is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).
+
+In \verb|LieGroups.jl|, the exponential function is given by \verb|exp(G,X)|\footnote{in-place variant \texttt{exp!(G, g, X)}.}.
+If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\frak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
+In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}.
+As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map.
+The Riemannian exponential map \verb|exp(M, p, X)|\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.} can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
+Here, again, to access the Riemannian exponential, one can use the base manifold of the Lie group, i.e.\ \verb|exp(base_manifold(G), g, X)| and have to make sure that \verb|X| is from the tangent space at \verb|g|.
+
+Locally around the identity element $e\in\mathcal G$, the exponential map is a diffeomorphism, i.e.\ there exists an inverse map on some neighborhood of $e$.
+The logarithmic function \(\log\colon \mathcal{G} \to \mathfrak{g}\) and logarithmic map \(\log_g\colon \mathcal{G} \to \mathfrak{g}\)have the function signatures \verb|log(G, g)|\footnote{in-place variant \texttt{log!(G, X, g)}.} and \verb|log(G, h, g)|\footnote{in-place variant \texttt{log!(G, X, h, g)}.}, respectively, with the same caveat to the Riemannian logarithmic map as for the exponential.
+
+When the exponential and logarithmic map are not known in closed form, it might be beneficial to use retractions and inverse retractions instead, respectively.
+These are first or second order approximations of the exponential and logarithmic map, respectively and their interface is already provided in \verb|ManifoldsBase.jl|\footnote{see \href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/retractions/}{juliamanifolds.github.io/ManifoldsBase.jl/stable/retractions/}.}.
+In \verb|LieGroups.jl|, one can either implement new variants based on a subtype of \verb|AbstractRetractionMethod| and \verb|AbstractInverseRetractionMethod|, resp., or use the wrappers \verb|BaseManifoldRetraction| and \verb|BaseManifoldInverseRetraction|, resp., to directly use the retraction and inverse retraction from the underlying manifold.
 
 \subsection{Lie algebras}
+
+Similar to elements of the Lie group, elements of the Lie algebra can have arbitrary \verb|struct|s to be represented in.
+The optional abstract supertype \verb|AbstractLieAlgebraTangentVector <: AbstractTangentVector| is provided to also distinguish different representations here.
+
+\verb|get_vector|, \verb|get_coordinates|,
+\verb|is_vector|, \verb|inner|, \verb|norm|,
+
+jacobians of exp and log
+
+\verb|pull_back_tangent|, \verb|push_forward_tangent|
+
 \subsection{Group actions}
 
+\begin{table}
+    \centering
+    \caption{Group actions types available in \texttt{LieGroups.jl}.}
+    \begin{tabular}{@{}ll@{}}
+        \toprule
+        \textbf{Action type} & comment/code\\
+        \midrule
+        \verb|AdditionGroupAction| & falls back to \verb|+|\\
+        \verb|ColumnwiseGroupAction{A}| & of a group action type \verb|A|\\
+        \verb|(Inverse)LeftGroupOperationAction| & requires $\mathcal M = \mathcal G$\\
+        \verb|LeftMultiplicationAction| & falls back to \verb|*|\\
+        \verb|(Inverse)RightGroupOperationAction| & requires $\mathcal M = \mathcal G$\\
+        \verb|RotationAroundAxisAction| & \\
+        \verb|RowwiseGroupAction{A}| & of a group action type \verb|A|\\
+        \bottomrule
+    \end{tabular}
+    \label{tab:GroupActionTypes}
+\end{table}
+
+Table~\ref{tab:GroupActionTypes} summarizes the currently implemented group action types in \verb|LieGroups.jl|.
+
+\verb|apply|,
+\verb|diff_apply|
+\verb|diff_group_apply|
+\verb|inv(A::GroupAction)|
+
 \section{Implemented Lie groups}\label{sec:LieGroups}
+
 \begin{table}[tbp]
     \centering
-    \caption{Implemented Lie groups in \texttt{LieGroups.jl} as of version 0.1.6}
+    \caption{Implemented Lie groups in \texttt{LieGroups.jl} version~\LieGroupsVersion{}.}
     \begin{tabular}{@{}lll@{}}
         \toprule
-        \textbf{Group} & \textbf{Symbol} & comment/code\\
+        \textbf{Lie Group $\mathcal G$} & \textbf{Symbol} & comment/code\\
         \midrule
@@
         \verb|UnitaryGroup(n)| & \(\mathrm{U}(n)\) & \\
+        \verb|ValidationLieGroup(G)| & \multicolumn{2}{l}{wraps \texttt{G} for numerical verification}\\
         \bottomrule
     \end{tabular}
+    \label{tab:available_lie_groups}
 \end{table}
+
+Table~\ref{tab:available_lie_groups} summarizes the currently implemented Lie groups in \verb|LieGroups.jl| \LieGroupsVersion.
+
+\subsection{Meta Lie groups}\label{subsec:MetaLieGroups}
+
+There are three Lie groups that are built upon other Lie groups. We mention them here breifly and point out specific functions and features that are additionally available for these.
+
+\subsubsection{Product Lie groups}
+\subsubsection{Power Lie groups}
+\subsubsection{Semidirect product Lie groups}
+
+\subsection{A Lie group for Numerical verification}
+
+The \verb|ValidationLieGroup| is a special Lie group that is intended for numerical verification and debugging of code using Lie groups.
+
+The \verb|ValidationLieGroup| is implemented as a wrapper around any existing Lie group, \verb|G2 = ValidationLieGroup(G)|.
+It provides additional functionality to check the correctness of computations involving the Lie group.
+Most prominently,
+all input and output of a function is checked for validity, e.g.\ by calling \verb|is_point| and \verb|is_vector| on all points and tangent vectors, respectively.
+Similarly the inner manifold is wrapped into the similar concept from \verb|ManifoldsBase.jl|, the \verb|ValidationManifold|.
+While by default these checks do result in errors, this can be changed setting the \verb|error=| keyword to either \verb|:warn| or \verb|:info|.
+
+\subsection{Concrete Lie groups}\label{subsec:ConcreteLieGroups}
 

### 2025-10-28 paper/paper.tex
Finish most of Lie algebra.

@@
 The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$, since we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
-Then, using the differential (or pushforward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
+Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
 \footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields.
@@
 We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
-Here, a tangent vector $X$ induces the constrant vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
+Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
 As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
@@
 
-In general, the interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
+The interface follows the philosophy of \verb|ManifoldsBase.jl|, that the Lie group or algebra is the first argument and even a mutated argument comes second.
 Similarly, if a function computes something like a new point on the Lie group or a tangent vector, there also exists a variant, that computes this in-place.
-For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default represention, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.
+For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.
 
@@
 From a topological viewpoint, one has to first distinguish between points on the Lie group and tangent vectors on the Lie algebra.
-\verb|LieGroups.jl| intrduces a \verb|AbstractLieGroupPoint <: AbstractManifoldPoint|, that is used to represent points on the Lie group.
-This abstract supertype is not necessary, points can be represented by arrays or other types as well, but the type allows to intrduce special types when one Lie groups has \emph{different representations} of points, that have to be distinguished.
-When there are different representations, it is recommended to introdcude a point type for each representation and make the default one fall back to (just) using arrays.
+\verb|LieGroups.jl| introduces a \verb|AbstractLieGroupPoint <: AbstractManifoldPoint|, that is used to represent points on the Lie group.
+This abstract supertype is not necessary, points can be represented by arrays or other types as well, but the type allows to introduce special types when one Lie groups has \emph{different representations} of points, that have to be distinguished.
+When there are different representations, it is recommended to introduce a point type for each representation and make the default one fall back to (just) using arrays.
 \\
@@
 \verb|is_approx(G, g, h)| to check for (approximate) equality of two points,
-\verb|is_point(G, g)| to check if a point is a valid point on the manifold,
+\verb|is_point(G, g; error=:none)| to check if a point is a valid point on the manifold, where the keyword can be used to throw an \verb|:error|, a \verb|:warn| or an \verb|:info|.
+Similarly
 \verb|manifold_dimension(G)| to get the dimension of the manifold,
-\verb|project(G, q)|\footnote{in-place variant \texttt{project!(G, p, q)}.}
-to project a point onto the manifold,
+\verb|project(G, q)|\footnote{in-place variant \texttt{project!(G, p, q)}.} to project a point onto the manifold, as well as
 \verb|rand(G)| and \verb|rand(rng, G)|\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
-to sample a random point from the manifold using a random number generator \verb|rng|.
+to sample a random point from the manifold using a random number generator \verb|rng| also pass directly on to the manifold.
+To access the underlying manifold of the Lie group one can use \verb|base_manifold(G)|.
 
@@
 Finally, there is a specific function for the differential of the conjugate at the identity $h=e$ called \verb|adjoint(G, g, X)|\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.} as well as the combinations \verb|inv_left_compose(G, g, h)|\footnote{in-place variant \texttt{inv\_left\_compose!(G, k, g, h)}} for computing $g^{-1}\cdot h$ and \verb|inv_right_compose(G, g, h)|\footnote{in-place variant \texttt{inv\_right\_compose!(G, k, g, h)}} for computing $g \cdot h^{-1}$.
-All three fall back to the previusly defined functions, but provide an interface to possibly implement more efficient variants in case such exist.
+All three fall back to the previously defined functions, but provide an interface to possibly implement more efficient variants in case such exist.
 
@@
 An important tool to “move around” on the Lie group is the exponential map.\\
-The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfills the differential equation
+The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
 \begin{equation*}
@@
 
-Locally around the identity element $e\in\mathcal G$, the exponential map is a diffeomorphism, i.e.\ there exists an inverse map on some neighborhood of $e$.
+Locally around the identity element $e\in\mathcal G$, the exponential map is a diffeomorphism, i.e.\ there exists an inverse map on some neighbourhood of $e$.
 The logarithmic function \(\log\colon \mathcal{G} \to \mathfrak{g}\) and logarithmic map \(\log_g\colon \mathcal{G} \to \mathfrak{g}\)have the function signatures \verb|log(G, g)|\footnote{in-place variant \texttt{log!(G, X, g)}.} and \verb|log(G, h, g)|\footnote{in-place variant \texttt{log!(G, X, h, g)}.}, respectively, with the same caveat to the Riemannian logarithmic map as for the exponential.
@@
 
-Similar to elements of the Lie group, elements of the Lie algebra can have arbitrary \verb|struct|s to be represented in.
-The optional abstract supertype \verb|AbstractLieAlgebraTangentVector <: AbstractTangentVector| is provided to also distinguish different representations here.
+Similar to points on the Lie group, when representing elements $X \in \frak g$ from a Lie algebra, we do not type the general functions of the interface.
+This allows to use either plain arrays or own structures to represent these in code.
+The optional abstract supertype \verb|AbstractLieAlgebraTangentVector|, which is a \verb|AbstractTangentVector|, is provided to also distinguish different representations here.
+Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.
 
-\verb|get_vector|, \verb|get_coordinates|,
-\verb|is_vector|, \verb|inner|, \verb|norm|,
+A major difference to the usual representation of tangent vectors on the underlying manifold is, that here the usual representation is done in the Lie algebra.
 
-jacobians of exp and log
+\subsubsection*{Topological basics.}
+Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\frak g$ \verb|= LieAlgebra(G)|.
+To access the Lie group again, use \verb|base_lie_group(|$\frak g$\verb|)|. Similarly \verb|base_manifold(|$\frak g$\verb|)| returns the underlying manifold of the Lie group.\\
+As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
+A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|.\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} contains the type automatically.}
 
-\verb|pull_back_tangent|, \verb|push_forward_tangent|
+The main topological function is to test the validity of a tangent vector \verb|is_vector(|$\frak g$\verb|, X; error=:none)| using the same \verb|error=| keyword as \verb|is_point| on the Lie group.
+
+\subsubsection*{Vector space related functions}
+For the following functions related to vector space features, using the Lie algebra $\frak g$ as first argument is equivalent to specifying the Lie Group $G$ and an arbitrary point $g \in \mathcal G$.
+This yields that the Lie group complies with the general interface for manifolds.
+
+The inner product and norm on the Lie algebra are available as \verb|inner(|$\frak g$\verb|, X, Y)| and \verb|norm(|$\frak g$\verb|, X)|, respectively.
+furthermore, there are two functions to convert between a coordinate-free representation of $X$ as a tangent vector and its representation in coordinates of a basis.
+
+Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, B)| where $B$ is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
+When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, B, T)| specifying the tangent vector type.\footnote{in-place variant \texttt{get\_vector!($\frak g$, X::T, c, B)} contains the type information directly.}
+Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, B)}}, where the Basis is again optional.
+
+For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\frak g$\verb|, c)|\footnote{in-place variant \texttt{hat!($\frak g$, X::T, c)}} (again with an optional vector type \verb|T|) and \verb|vee(|$\frak g$\verb|, X)|\footnote{in-place variant \texttt{vee!($\frak g$, c, X)}} are implemented
+
+\subsubsection*{Push forward and pull back of tangent vectors}
+To the representation of a tangent vector $X \in \frak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$.
+This is implemented as \verb|push_forward_tangent(G, g, X)|\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
+Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\frak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$.
+This is implemented as \verb|pull_back_tangent(G, g, Y)|\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
+
+\subsubsection*{Jacobians.}
+On Euclidean space, the terms differential and Jacobian are often used interchangeably.
+On manifolds, the differential of a smooth map between two manifolds is a map between the corresponding tangent spaces.
+To represent this map in coordinates, one has to choose bases for the tangent spaces, resulting in a Jacobian matrix.
 
@@
 
-There are three Lie groups that are built upon other Lie groups. We mention them here breifly and point out specific functions and features that are additionally available for these.
+There are three Lie groups that are built upon other Lie groups. We mention them here briefly and point out specific functions and features that are additionally available for these.
 

### 2025-10-28 paper/paper.tex
check sections 1 and 2

@@
     \href{https://juliamanifolds.github.io/LieGroups.jl/stable/}{\texttt{LieGroups.jl}} is a
-    Julia package that provides on the one hand an interface to define Lie groups as well as the corresponding Lie algebra and Lie group actions.
-On the other hand, it offers a well-documented, performant, and well-tested library of existing Lie groups, their Algebra and corresponding group actions.
+    Julia package that provides an interface to define Lie groups as well as the corresponding Lie algebra and Lie group actions.
+The package also offers a well-documented, performant, and well-tested library of existing Lie groups, their algebra and corresponding group actions.
 
@@
 In many situations, one encounters data that does not reside in a vector space.
-We can hence not used standard linear algebra tools to work with such data.
+We can hence not use standard linear algebra tools to work with such data.
 For example in robotics, the configuration space of a rigid body in three-dimensional
@@
 These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
-They have applications e.g.\ in physics, robotics, stochastic processes, information geometry, see e.g.~\cite{Chirikjian:2009,Chirikjian:2012},
+They have applications in physics, robotics, stochastic processes, information geometry, and many other areas see~\cite{Chirikjian:2009,Chirikjian:2012},
 but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
-and their numerical aspects, e.g.\ when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998}.
+and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998}.
 
@@
     \includegraphics[width=0.17\textwidth]{logo.png}
-    \caption*{\ \\[-.5\baselineskip]Logo of \texttt{LieGroups.jl}.}%
+    \vspace{.5\baselineskip}
+    \caption*{Logo of \texttt{LieGroups.jl}.}%
     \label{fig:liegroups_logo}
@@
 \end{equation*}
+% MB: maybe it would be better to replace it with an example that can have a group structure? For example the 3-dimensional sphere?
 
@@
 \\
-As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
+As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
 Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
-As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), i.e.\ for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
-For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$, or in other words the circle. % chktex 9ok
+As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
+For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{The identification is not continuous.}, or, continuously, with the circle.
+% chktex 9ok
 The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose (or an angle $-\alpha$).
+% -(-\pi) would be out of the range
 
 \subsection{Tangent spaces and the Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
-The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$, since we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
-Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
+The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
+Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$\footnote{\color{red} MB: TODO: does the reader know that a differential is?}, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
 \footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields.
@@
 As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
-The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
+The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
 For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$.
@@
 In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \frak{g}$.
+
+An important tool to “move around” on the Lie group is the exponential map.
+The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
+\begin{equation*}
+\gamma'(t) = D\lambda_{\gamma(t)}(e)[X]
+\quad\text{ with }\quad\gamma(0) = e \text{ and }\gamma'(0) = X
+\end{equation*}
+at time $t=1$, that is \(\exp(X) = \gamma(1)\).
+This can be interpreted as starting at the identity element \(e \in \mathcal{G}\) and following the curve whose velocity at each point is given by the left-invariant vector field induced by \(X \in \mathfrak{g}\) for one time unit.
+
+For \((\mathbb{R}^n, +)\) the exponential map is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
+Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential map is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).
 
@@
 
-\subsection{Group Actions} A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).
+Note that a Riemannian metric can also be used to define an exponential map using the Levi-Civita affine connection.
+This exponential map often differs from the Lie group exponential, in particular many Lie groups, such as the special Euclidean group in two or more dimensions, can not have a metric compatible with the Lie group exponential.
+
+\subsection{Group Actions} A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).\footnote{This is the convention for left actions.
+Alternatively, right actions fulfil \(\sigma(g, \sigma(h, p)) = \sigma(h \cdot g, p)\).}
 Informally a group action describes how elements of the Lie group \(\mathcal{G}\) “act on” points on the manifold \(\mathcal{M}\).
@@
 \subsubsection*{exponential and logarithm.}
-An important tool to “move around” on the Lie group is the exponential map.\\
-The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
-\begin{equation*}
-\gamma'(t) = D\lambda_{\gamma(t)}(e)[X]
-\quad\text{ with }\quad\gamma(0) = e \text{ and }\gamma'(0) = X
-\end{equation*}
-at time $t=1$, i.e.\ \(\exp(X) = \gamma(1)\).
-
-This can be interpreted as “starting at the identity element \(e \in \mathcal{G}\) and following the curve whose velocity at each point is given by the left-invariant vector field induced by \(X \in \mathfrak{g}\) for one time unit”.
-
-In our examples, for \((\mathbb{R}^n, +)\) the exponential map is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
-Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential map is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).
 

### 2025-10-28 paper/paper.tex
Further work on section 4.

--- 2025-10-28 paper/paper.tex
@@
     \centering
-    \includegraphics[width=0.17\textwidth]{logo.png}
+    \includegraphics[width=0.15\textwidth]{logo.png}
     \caption*{\ \\[-.5\baselineskip]Logo of \texttt{LieGroups.jl}.}%
@@
 \begin{equation*}
-\mathbb{S}^2 = \{x \in \mathbb{R}^3 \mid \lVert x\rVert = 1\},
+\mathbb{S}^2 = \bigl\{ p \in \mathbb{R}^3 \bigm| \lVert p\rVert = 1\bigr\},
 \end{equation*}
@@
 
-\subsubsection*{exponential and logarithm.}
+\subsubsection*{Exponential and logarithm.}
 An important tool to “move around” on the Lie group is the exponential map.\\
@@
 The inner product and norm on the Lie algebra are available as \verb|inner(|$\frak g$\verb|, X, Y)| and \verb|norm(|$\frak g$\verb|, X)|, respectively.
-furthermore, there are two functions to convert between a coordinate-free representation of $X$ as a tangent vector and its representation in coordinates of a basis.
+Furthermore, there are two functions to convert between a coordinate-free representation of $X$ as a tangent vector and its representation in coordinates of a basis.
 
-Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, B)| where $B$ is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
-When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, B, T)| specifying the tangent vector type.\footnote{in-place variant \texttt{get\_vector!($\frak g$, X::T, c, B)} contains the type information directly.}
-Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, B)}}, where the Basis is again optional.
+Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
+When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type.\footnote{in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} contains the type information directly.}
+Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, b)}}, where the Basis is again optional.
 
@@
 
-\subsubsection*{Jacobians.}
+\subsubsection*{Jacobians}
 On Euclidean space, the terms differential and Jacobian are often used interchangeably.
-On manifolds, the differential of a smooth map between two manifolds is a map between the corresponding tangent spaces.
-To represent this map in coordinates, one has to choose bases for the tangent spaces, resulting in a Jacobian matrix.
+Within differential geometry, the differential (or push forward) of a smooth map $f\colon \mathcal M \to \mathcal N$ between two manifolds is a map between the corresponding tangent spaces.
+We denote it by $Df(p)\colon T_p\mathcal M \to T_{f(p)}\mathcal N$, where $p \in \mathcal M$.
+The differential is a linear map between the tangent spaces.
+However, in the coordinate-free representation of tangent vectors, there is no matrix representation of this linear map.
+But as soon as we choose bases for the tangent spaces and we represent tangent vectors in coordinates (cf.~\verb|get_coordinates| previously), we can represent the differential as a matrix, called the Jacobian matrix.
+
+Currently for both \verb|exp| as well as \verb|conjugate|, the Jacobian is implemented.
+Note that for \verb|exp|, the domain is the Lie algebra.
+Since that is a vector space, its tangent space(s) can be identified with the Lie algebra again.
+Hence both Jacobians are maps from the Lie algebra to itself.
+The signatures are
+\verb|jacobian_exp(G, p, X, b)|\footnote{in-place variant \texttt{jacobian\_exp!($\frak g$, J, p, X, b)}} and \verb|jacobian_conjugate(G, g, X, b)|\footnote{in-place variant \texttt{jacobian\_conjugate!($\frak g$, J, g, X, b)}}, where \verb|b| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.
 
@@
         \toprule
-        \textbf{Action type} & comment/code\\
+        \textbf{Group action type} & comment/code\\
         \midrule
@@
 
-Table~\ref{tab:GroupActionTypes} summarizes the currently implemented group action types in \verb|LieGroups.jl|.
+Group actions as defined in Subsection~\ref{subsec:GroupActions} are implemented as a struct \verb|GroupAction{G, M, A} <: AbstractGroupAction| containing the Lie group \verb|G <: AbstractLieGroup|, the manifold \verb|M|, and a group action tpe \verb|A| in order to distinguish possible different ways a group could act on a manifold.
+Table~\ref{tab:GroupActionTypes} summarizes the currently available group action types in \verb|LieGroups.jl|.
 
-\verb|apply|,
-\verb|diff_apply|
-\verb|diff_group_apply|
-\verb|inv(A::GroupAction)|
+the main functions then are \verb|apply(a::GroupAction, g, p)|%
+\footnote{in-place variant \texttt{apply!(a::GroupAction, q, g, p)}.}
+to compute \(\sigma(g, p)\) for a group action \verb|a|, where $g\in \mathcal G$ and $p \in \mathcal M$, as well as its differential \verb|diff_apply(a::GroupAction, g, p, X)|%
+\footnote{in-place variant \texttt{diff\_apply!(a::GroupAction, Y, g, p, X)}.}
+to compute \(D\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|%
+\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
+to compute \(D\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \frak g\).
+
+Due to the property that \(\sigma(g^{-1}, \sigma(g, p)) = p\) for all \(g \in \mathcal{G}\) and \(p \in \mathcal{M}\), there is also an inverse group action available $\sigma^{-1}(g, p) = \sigma(g^{-1}, p)$, which can be obtained by calling \verb|inv(a)|.
+
+{\color{red} TODO: Discuss left and right group actions here?}
 
@@
 
+For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([0, 2\pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
+These are constructed by specifying the representation of points, i.e.
+\verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively.
+Note that each of these representations uses a different underlying manifold, namely \verb|Circle(|$\mathbb R$\verb|)|, \verb|Circle(|$\mathbb C$\verb|)|, and \verb|Circle(|$\mathbb R^2$\verb|)|, respectively.
+This is a case, where it is not only the representation of points and tangent vectors that differ, but indeed even the underlying manifold.
+
 \section{An example how to use {\texttt{LieGroups.jl}}}\label{sec:Example}

### 2025-10-28 paper/paper.tex

--- 2025-10-28 paper/paper.tex
@@
     \centering
-    \includegraphics[width=0.17\textwidth]{logo.png}
+    \includegraphics[width=0.12\textwidth]{logo.png}
     \vspace{.5\baselineskip}
@@
 
-\subsubsection*{exponential and logarithm.}
+\subsubsection*{Exponential and logarithm.}
 

### 2025-10-28 paper/paper.tex
a few small improvements

@@
 
-The main type for Lie Groups is the \verb|LieGroup{|$\mathbb {F}$\verb|, O, M} <: AbstractLieGroup|, which contains a manifold \verb|M <: AbstractManifold{|$\mathbb F$\verb|}| as well as the group operation \verb| O <: AbstractGroupOperation|, where $\mathbb {F}$ is the field of scalars used in the representation of the manifold, usually \(\mathbb {F} = \mathbb{R}\) and in some cases \(\mathbb {F} = \mathbb{C}\) or \(\mathbb F = \mathbb H\).
+The main type for Lie Groups is the \verb|LieGroup{|$\mathbb {F}$\verb|, O, M} <: AbstractLieGroup|, which contains a manifold \verb|M <: AbstractManifold{|$\mathbb F$\verb|}| as well as the group operation \verb| O <: AbstractGroupOperation|, where $\mathbb {F}$ is the set of scalars used in the representation of the manifold, usually \(\mathbb {F} = \mathbb{R}\) and in some cases \(\mathbb {F}\) is the field of complex numbers \(\mathbb{C}\) or the ring of quaternions \(\mathbb H\).
 
 \subsubsection*{Topological functions}
-From a topological viewpoint, one has to first distinguish between points on the Lie group and tangent vectors on the Lie algebra.
-\verb|LieGroups.jl| introduces a \verb|AbstractLieGroupPoint <: AbstractManifoldPoint|, that is used to represent points on the Lie group.
-This abstract supertype is not necessary, points can be represented by arrays or other types as well, but the type allows to introduce special types when one Lie groups has \emph{different representations} of points, that have to be distinguished.
+Points and tangent vectors can be represented by the same Julia objects, although they are distinct from a topological viewpoint.
+\footnote{\texttt{LieGroups.jl} introduces a \texttt{AbstractLieGroupPoint <: AbstractManifoldPoint}, that is used to represent points on the Lie group.
+This abstract supertype is not necessary, points can be represented by arrays or other types as well.}
+Separate types are necessary when one Lie group has \emph{different representations} of points, that have to be distinguished.
 When there are different representations, it is recommended to introduce a point type for each representation and make the default one fall back to (just) using arrays.
@@
 This allows to use either plain arrays or own structures to represent these in code.
-The optional abstract supertype \verb|AbstractLieAlgebraTangentVector|, which is a \verb|AbstractTangentVector|, is provided to also distinguish different representations here.
+\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}
 Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.

### 2025-10-28 paper/paper.tex
Started with meta Lie groups.

--- 2025-10-28 paper/paper.tex
@@
 
-\subsubsection{Product Lie groups}
+\subsubsection{Product Lie group}
+
+Given two Lie groups \(\mathcal{G} = (\mathcal M, \ast) \) and \(\mathcal{H} = (\mathcal N, \diamond)\), their (direct) product \(\mathcal{G} \times \mathcal{H}\) is again a Lie group \((\mathcal{M} \times \mathcal{N}, \cdot)\) with group operation defined component-wise, i.e.
+\begin{equation*}
+    (g_1, h_1) \cdot (g_2, h_2) = (g_1 \ast g_2, h_1 \diamond h_2)
+\end{equation*}
+for \(g_1, g_2 \in \mathcal{G}\) and \(h_1, h_2 \in \mathcal{H}\).
+Since $\times$ is a binary operator in Julia, given two Lie Groups \verb|G,H|, their product Lie group can be constructed via \verb|G|\,$\times$\,\verb|H|.
+The same applies for more than two groups, i.e. you can construct arbitrary long product Lie groups \verb|G1|\,$\times$\,\verb|G2|\,$\times$\,\verb|G3|\,$\times$\,\ldots.
+
+As for the representation of points and tangent vectors, the default representation requires to load \verb|RecursiveArrayTools.jl|\footnote{see \href{https://docs.sciml.ai/RecursiveArrayTools/stable/}{docs.sciml.ai/RecursiveArrayTools/stable/}}.
+
+If you want to construct nested Products of Lie groups, e.g.\ \((\mathcal G_1 \times \mathcal G_2) \times \mathcal G_3\), the binary operator would not work, since it automatically flattens the input; use \verb|ProductLieGroup(ProductLieGroup(G1, G2), G3)| can be used instead.
+
 \subsubsection{Power Lie groups}
+Though mathematically the case of a product Lie Group, where $G=H$ is also (just) a product, computationally we can benefit from the special structure.
+Therefore the \verb|Powermanifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(G\) with itself.
+As a shortcut, you can use \verb|G^n| to construct the \(n\)-fold Cartesian product of the Lie group \(G\) with itself.
+
+Note that inheriting from the power manifold, there are two different data types to represent data, per default a single large array, or a
+
 \subsubsection{Semidirect product Lie groups}

### 2025-10-28 paper/paper.tex
Finish section 4, first complete draft.

@@
 
-\subsubsection{Product Lie group}
+\subsubsection*{Product Lie group}
 
@@
 
-\subsubsection{Power Lie groups}
+\subsubsection*{Power Lie groups}
 Though mathematically the case of a product Lie Group, where $G=H$ is also (just) a product, computationally we can benefit from the special structure.
@@
 
-Note that inheriting from the power manifold, there are two different data types to represent data, per default a single large array, or a
+Note that inheriting from the power manifold, there are two different data types to represent data, per default a single large array.
+Alternatively, one can use the constructor \verb|PowerLieGroup(G, NestedPowerRepresentation(), n)| to represent points and tangent vectors as a vector of points.
 
-\subsubsection{Semidirect product Lie groups}
+To abstract the access of elements here, similar to the \verb|PowerManifold| in \verb|ManifoldsBase.jl|, the access functions \verb|g[G, i]| are available.
+
+\subsubsection*{Semidirect product Lie groups}
+
+Given two Lie groups \(\mathcal{G} = (\mathcal M, \ast)\) and \(\mathcal{H} = (\mathcal N, \diamond)\) as well as a group action \(\sigma\colon \mathcal{G} \times \mathcal{H} \to \mathcal{H}\),
+the \term{(left) semidirect product Lie Group} $\mathcal{G} \ltimes \mathcal{H} = (\mathcal{M} \times \mathcal{N}, \cdot)$ is defined via the group operation given by
+\begin{equation*}
+    (g_1, h_1) \cdot (g_2, h_2) = (g_1 \ast g_2, h_1 \diamond \sigma(g_1, h_2))
+\end{equation*}
+This semidirect product Lie group is constructed calling \verb|LeftSemidirectProductLieGroup(G, H, a)|, where the \verb|GroupAction a| is optional, since one can define a default action by setting \verb|default_left_action(G, H)|.
+With this default, also the shortcut \verb|G|\,$\ltimes$\,\verb|H| is available.
+%
+Similarly, the \term{right semidirect product Lie Group} $\mathcal{H} \rtimes \mathcal{G}$ is defined analogously with the positions of the groups swapped in the group operation and the default for the group action is given by \verb|default_right_action(G, H)|.
+This can be constructed via \verb|H|\,$\rtimes$\,\verb|G| then as well.
 
@@
 
+From the list of available Lie groups in Table~\ref{tab:available_lie_groups}, we want to highlight some specific features of a few of them.
+
 For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([0, 2\pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
@@
 This is a case, where it is not only the representation of points and tangent vectors that differ, but indeed even the underlying manifold.
+
+The special Euclidean group \(\mathcal G = \mathrm{SE}(n)\), there are three different representations avalable:
+\\
+The first one is the matrix representation, where points are represented as \((n+1) \times (n+1)\) matrices combining rotation and translation, i.e.\ as
+\begin{equation*}
+    g = \begin{pmatrix} R & t \\
+    0 & 1 \end{pmatrix},
+\end{equation*}
+where \(R \in \mathrm{SO}(n)\) and \(t \in \mathbb{R}^n\).
+
+The second representation is the tuple representation, where points are represented as tuples \(g' = (R, t)\) and the third is the tuple representation with the translation in the first component, i.e.\ \(g'' = (t, R)\).
+These two are available when \verb|RecursiveArrayTools.jl| is loaded and specifying the \verb|variant=:left| (default) or \verb|variant=:right|, respectively.
+The first, matrix representation work on either \verb|variant|.
+
+For all three representations, one can access the component semantically by writing \verb|g[G, :Rotation]| and \verb|g[G, :Translation]|, which works independent of the representation used.
+
+{\color{red}TODO: Check whether there are details for other Lie groups that should be mentioned here.}
 

### 2025-10-28 paper/paper.tex
Extend a footnote.

@@
 As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
-For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{The identification is not continuous.}, or, continuously, with the circle.
+For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{Note that the identification is not continuous.}, or, continuously, with the circle.
 % chktex 9ok

### 2025-10-28 paper/paper.tex
Start reworking the footnotes.

@@
 The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
-Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$\footnote{\color{red} MB: TODO: does the reader know that a differential is?}, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
+Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$\footnote{\color{red} MB: TODO: does the reader know that a differential is?
+(RB) Sure I can expand on that, I wrote a sentence more further below to get from differential to Jacobian, so I can also add a sentence there as well.}, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
 \footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields.
@@
 
+In general, when a function within the interfaces of \verb|JuliaManifolds| returns a point, a tangent vector or any other mutable object, there always exist two variants of the function, one that allocates a new object and returns it, and one that takes an additional argument to write the result into.
+For example for a function \verb|f(G, g)| that returns a point on the Lie group \verb|G|, there also exists a variant \verb|f!(G, h, g)| that computes the result in-place of \verb|h|, possibly using the memory also for interims results.
+By default the allocating variant is implemented by allocating memory accordingly and calling the in-place variant.
+This default can be overriden, for example for performance reasons.
+
 \subsubsection*{Topological functions}
-Points and tangent vectors can be represented by the same Julia objects, although they are distinct from a topological viewpoint.
+Points and tangent vectors can be represented by the same Julia objects, although they are distinct from a topological viewpoint.%
 \footnote{\texttt{LieGroups.jl} introduces a \texttt{AbstractLieGroupPoint <: AbstractManifoldPoint}, that is used to represent points on the Lie group.
@@
 When there are different representations, it is recommended to introduce a point type for each representation and make the default one fall back to (just) using arrays.
-\\
+Several functions in the interface return points or tangent vectors, where the type might not be known from the input argument.
+For such a function, for example again a function \verb|f(G)| a positional argument \verb|T| can be used to specify the desired return type, i.e.\ \verb|f(G, T)| returns a point or tangent vector of type \verb|T|.
+The in-place variant does not require this, since in its signatgure of the form \verb|f!(G, g::T)| the type \verb|T| is known from the argument \verb|g|.
+
 Alternatively to points, also different manifolds can be used.
@@
 
-Beyond that, the following functions are available and pass directly to the underlying manifold interface from \verb|ManifoldsBase.jl|:
+The following functions are available and pass directly to the underlying manifold interface from \verb|ManifoldsBase.jl|:
 \verb|is_approx(G, g, h)| to check for (approximate) equality of two points,
@@
 \verb|manifold_dimension(G)| to get the dimension of the manifold,
-\verb|project(G, q)|\footnote{in-place variant \texttt{project!(G, p, q)}.} to project a point onto the manifold, as well as
-\verb|rand(G)| and \verb|rand(rng, G)|\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
+\verb|project(G, q)|%
+%\footnote{in-place variant \texttt{project!(G, p, q)}.} to project a point onto the manifold, as well as
+\verb|rand(G)| and \verb|rand(rng, G)|%
+%\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
 to sample a random point from the manifold using a random number generator \verb|rng| also pass directly on to the manifold.
@@
 \verb|identity_element(G)| or \verb|identity_element(G, typeof(g))|, where \verb|g| is an element of the group.
-The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|.\footnote{The in-place variant \texttt{identity\_element!(G, e)} infers the representation from the type of \texttt{e}.}
+The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|.%
+\footnote{The in-place variant \texttt{identity\_element!(G, e::T)} infers the representation from the type \texttt{T} of \texttt{e}.}
 
-The two central functions are \verb|compose(G, g, h)|\footnote{in-place variant \texttt{compose!(G, k, g, h)}.} to compute the group operation \(g \cdot h\) for two points \(g, h \in \mathcal{G}\) and the inverse \verb|inv(G, g)|\footnote{in-place variant \texttt{inv!(G, h, g)}.} which computes \(g^{-1}\) for a point \(g \in \mathcal{G}\).
+The two central functions are \verb|compose(G, g, h)|
+%\footnote{in-place variant \texttt{compose!(G, k, g, h)}.}
+to compute the group operation \(g \cdot h\) for two points \(g, h \in \mathcal{G}\) and the inverse \verb|inv(G, g)|
+%\footnote{in-place variant \texttt{inv!(G, h, g)}.}
+which computes \(g^{-1}\) for a point \(g \in \mathcal{G}\).
 
-Another function is $c_g(h) = g \cdot h \cdot g^{-1}$ called \term{conjugation}, which is available as \verb|conjugate(G, g, h)|\footnote{in-place variant \texttt{conjugate!(G, k, g, h)}.}.
+Another function is $c_g(h) = g \cdot h \cdot g^{-1}$ called \term{conjugation}, which is available as \verb|conjugate(G, g, h)|%
+%\footnote{in-place variant \texttt{conjugate!(G, k, g, h)}.}
+.
 
-For these three functions, also the differentials are available: adding a \verb|diff_| prefix to the function name and a final argument for the Lie algebra tangent vector, for compose additionally the argument with which to  differentiate to, i.e.\ \verb|diff_left_compose(G, g, h, X)| and \verb|diff_right_compose(G, g, h, X)|, respectively.\footnote{in-place variants have a second argument \texttt{Y} to compute the result in.}
+For these three functions, also the differentials are available: adding a \verb|diff_| prefix to the function name and a final argument for the Lie algebra tangent vector, for compose additionally the argument with which to  differentiate to, i.e.\ \verb|diff_left_compose(G, g, h, X)| and \verb|diff_right_compose(G, g, h, X)|, respectively%
+%\footnote{in-place variants have a second argument \texttt{Y} to compute the result in.}
+.
 
-Finally, there is a specific function for the differential of the conjugate at the identity $h=e$ called \verb|adjoint(G, g, X)|\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.} as well as the combinations \verb|inv_left_compose(G, g, h)|\footnote{in-place variant \texttt{inv\_left\_compose!(G, k, g, h)}} for computing $g^{-1}\cdot h$ and \verb|inv_right_compose(G, g, h)|\footnote{in-place variant \texttt{inv\_right\_compose!(G, k, g, h)}} for computing $g \cdot h^{-1}$.
-All three fall back to the previously defined functions, but provide an interface to possibly implement more efficient variants in case such exist.
+Finally, there is a specific function for the differential of the conjugate at the identity $h=e$ called \verb|adjoint(G, g, X)|
+%\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.}
+as well as the combinations \verb|inv_left_compose(G, g, h)|
+%\footnote{in-place variant \texttt{inv\_left\_compose!(G, k, g, h)}}
+for computing $g^{-1}\cdot h$ and \verb|inv_right_compose(G, g, h)|
+%\footnote{in-place variant \texttt{inv\_right\_compose!(G, k, g, h)}
+for computing $g \cdot h^{-1}$. All three fall back to the previously defined functions, but provide an interface to possibly implement more efficient variants in case such exist.
 
@@
 
-In \verb|LieGroups.jl|, the exponential function is given by \verb|exp(G,X)|\footnote{in-place variant \texttt{exp!(G, g, X)}.}.
+In \verb|LieGroups.jl|, the exponential function is given by \verb|exp(G,X)|%
+%\footnote{in-place variant \texttt{exp!(G, g, X)}.}
+.
 If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\frak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
-In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}.
-As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map.
-The Riemannian exponential map \verb|exp(M, p, X)|\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.} can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
+In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
+%\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}
+.
+As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map. The Riemannian exponential map \verb|exp(M, p, X)|
+%\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.}
+can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
 Here, again, to access the Riemannian exponential, one can use the base manifold of the Lie group, i.e.\ \verb|exp(base_manifold(G), g, X)| and have to make sure that \verb|X| is from the tangent space at \verb|g|.
@@
 Locally around the identity element $e\in\mathcal G$, the exponential map is a diffeomorphism, i.e.\ there exists an inverse map on some neighbourhood of $e$.
-The logarithmic function \(\log\colon \mathcal{G} \to \mathfrak{g}\) and logarithmic map \(\log_g\colon \mathcal{G} \to \mathfrak{g}\)have the function signatures \verb|log(G, g)|\footnote{in-place variant \texttt{log!(G, X, g)}.} and \verb|log(G, h, g)|\footnote{in-place variant \texttt{log!(G, X, h, g)}.}, respectively, with the same caveat to the Riemannian logarithmic map as for the exponential.
+The logarithmic function \(\log\colon \mathcal{G} \to \mathfrak{g}\) and logarithmic map \(\log_g\colon \mathcal{G} \to \mathfrak{g}\)have the function signatures \verb|log(G, g)|%
+%\footnote{in-place variant \texttt{log!(G, X, g)}.}
+and \verb|log(G, h, g)|%
+%\footnote{in-place variant \texttt{log!(G, X, h, g)}.}
+, respectively, with the same caveat to the Riemannian logarithmic map as for the exponential.
 
@@
 As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
-A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|.\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} contains the type automatically.}
+A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|.\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
 
@@
 Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
-When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type.\footnote{in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} contains the type information directly.}
-Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, b)}}, where the Basis is again optional.
+When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type
+\footnote{the in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
+Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|%
+%\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, b)}}
+, where the Basis is again optional.
 

### 2025-10-28 paper/paper.tex
Finish refactor footnotes- and adding a note of left/right actions.

@@
 \verb|identity_element(G)| or \verb|identity_element(G, typeof(g))|, where \verb|g| is an element of the group.
-The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|.%
-\footnote{The in-place variant \texttt{identity\_element!(G, e::T)} infers the representation from the type \texttt{T} of \texttt{e}.}
+The first variant will always generate a point in the default representation of the group, while the latter will generate a point in the same representation as \verb|g|%
+%\footnote{The in-place variant \texttt{identity\_element!(G, e::T)} infers the representation from the type \texttt{T} of \texttt{e}.}
+.
 
@@
 As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
-A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|.\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
+A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|%
+%\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
+.
 
@@
 Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
-When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type
-\footnote{the in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
+When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type%
+%\footnote{the in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
+.
 Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|%
@@
 
-For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\frak g$\verb|, c)|\footnote{in-place variant \texttt{hat!($\frak g$, X::T, c)}} (again with an optional vector type \verb|T|) and \verb|vee(|$\frak g$\verb|, X)|\footnote{in-place variant \texttt{vee!($\frak g$, c, X)}} are implemented
+For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\frak g$\verb|, c)|
+%\footnote{in-place variant \texttt{hat!($\frak g$, X::T, c)}}
+(again with an optional vector type \verb|T|) and \verb|vee(|$\frak g$\verb|, X)|
+%\footnote{in-place variant \texttt{vee!($\frak g$, c, X)}}
+are implemented
 
@@
 To the representation of a tangent vector $X \in \frak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$.
-This is implemented as \verb|push_forward_tangent(G, g, X)|\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
+This is implemented as \verb|push_forward_tangent(G, g, X)|%
+%\footnote{in-place variant \texttt{push\_forward\_tangent!(G, Y, g, X)}.}
+.
 Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\frak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$.
-This is implemented as \verb|pull_back_tangent(G, g, Y)|\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
+This is implemented as \verb|pull_back_tangent(G, g, Y)|%
+%\footnote{in-place variant \texttt{pull\_back\_tangent!(G, X, g, Y)}.}
+.
 
@@
 The signatures are
-\verb|jacobian_exp(G, p, X, b)|\footnote{in-place variant \texttt{jacobian\_exp!($\frak g$, J, p, X, b)}} and \verb|jacobian_conjugate(G, g, X, b)|\footnote{in-place variant \texttt{jacobian\_conjugate!($\frak g$, J, g, X, b)}}, where \verb|b| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.
+\verb|jacobian_exp(G, p, X, b)|
+%\footnote{in-place variant \texttt{jacobian\_exp!($\frak g$, J, p, X, b)}}
+and \verb|jacobian_conjugate(G, g, X, b)|%
+%\footnote{in-place variant \texttt{jacobian\_conjugate!($\frak g$, J, g, X, b)}}
+, where \verb|b| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.
 
@@
 
-the main functions then are \verb|apply(a::GroupAction, g, p)|%
-\footnote{in-place variant \texttt{apply!(a::GroupAction, q, g, p)}.}
-to compute \(\sigma(g, p)\) for a group action \verb|a|, where $g\in \mathcal G$ and $p \in \mathcal M$, as well as its differential \verb|diff_apply(a::GroupAction, g, p, X)|%
-\footnote{in-place variant \texttt{diff\_apply!(a::GroupAction, Y, g, p, X)}.}
-to compute \(D\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|%
-\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
+the main functions then are \verb|apply(a::GroupAction, g, p)|
+%\footnote{in-place variant \texttt{apply!(a::GroupAction, q, g, p)}.}
+to compute \(\sigma(g, p)\) for a group action \verb|a|, where $g\in \mathcal G$ and $p \in \mathcal M$, as well as its differential \verb|diff_apply(a::GroupAction, g, p, X)|
+%\footnote{in-place variant \texttt{diff\_apply!(a::GroupAction, Y, g, p, X)}.}
+to compute \(D\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|
+%\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
 to compute \(D\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \frak g\).
@@
 
-{\color{red} TODO: Discuss left and right group actions here?}
+Note that while the group action $\sigma(g, \sigmaæ(h, p)) = \sigma(g \cdot h, p)$ “appends” a new action (w.r.t.~$g$) on the left, for the inverse we have
+\begin{align*}
+        \sigma^{-1}(g, \sigma^{-1}(h, p))
+        &= \sigma(g^{-1}\cdot h^{-1}, p)
+        = \sigma((h \cdot g)^{-1}, p)
+        \\
+        &= \sigma^{-1}(h \cdot g, p)
+\end{align*}
+for all \(g, h \in \mathcal{G}\). This is a so-called right action.
+Similarly for a right group action the inverse is a left group action.
+This is tanken into account when using \verb|apply|.
 
@@
 \section{An example how to use {\texttt{LieGroups.jl}}}\label{sec:Example}
+
+{\color{red}TODO: Brainstorm over one or two examples with code.}
 \input{bib.tex}

### 2025-10-28 paper/paper.tex
Add a todo to discuss.
remove unnecessary character

@@
 
-Note that while the group action $\sigma(g, \sigmaæ(h, p)) = \sigma(g \cdot h, p)$ “appends” a new action (w.r.t.~$g$) on the left, for the inverse we have
+Note that while the group action $\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)$ “appends” a new action (w.r.t.~$g$) on the left, for the inverse we have
 \begin{align*}
@@
 
-\section{An example how to use {\texttt{LieGroups.jl}}}\label{sec:Example}
+\section{An example how to use {\texttt{LieGroups.jl}}}%
+\label{sec:Example}
 
 {\color{red}TODO: Brainstorm over one or two examples with code.}
+
 \input{bib.tex}

### 2025-10-28 paper/paper.tex
start an example

--- 2025-10-28 paper/paper.tex
@@
 
-\section{An example how to use {\texttt{LieGroups.jl}}}\label{sec:Example}
+\section{An example how to use {\texttt{LieGroups.jl}}}%
+\label{sec:Example}
+
+In this section, we provide examples of how to use the {\texttt{LieGroups.jl}} package in Julia.
+
+\subsection{A Beziér curve on a Lie group}
+
+In the first example we illustrate how to construct and plot a Beziér curve on a Lie group, here the special Euclidean group \(\mathrm{SE}(2)\).
+
+We first can generalize the de-Casteljau algorithm for Beziér curves to Lie groups by replacing the linear interpolation step with using the exponential and logarithmic map.
+
+\begin{lstlisting}[
+    language = Julia,
+    numbers=left,
+    label={lst:deCasteljauLieGroup},
+    caption={de-Casteljau algorithm on a Lie group.}
+]
+using LieGroups, ManifoldsBase, RecursiveArrayTools
+using CairoMakie
+"""
+    deCasteljauLieGroup(G, cp, t)
+
+Evaluate the Beziér curve at parameter `t` given
+a vector of control points `cp` on the Lie group `G`.
+"""
+function deCasteljauLieGroup(
+        G::AbstractLieGroup, cp::Vector{P}, t::Real
+    ) where {P}
+    n = length(cp)
+    points = [copy(G, g) for g in cp]
+    for r in 1:n-1
+        for i in 1:n-r
+            X = log(G, points[i], points[i+1])
+            exp!(G, points[i], points[i], t * X)
+        end
+    end
+    return points[1]
+end
+\end{lstlisting}
+
+In Algorithm~\ref{lst:deCasteljauLieGroup}, define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
+
+We can now compare the algorithm on the Lie group \(\mathrm{SE}(2)\) with the corresponding Bézier curve on the (direct) product Lie group $\mathrm{SO}(2) \times \mathbb{R}^2$.
+Denote by $R_\alpha$ the rotation matrix in \(\mathrm{SO}(2)\) for an angle \(\alpha\in[0,2\pi)\).
+Then we consider the control points
+\begin{align*}
+    g_1 &= \begin{pmatrix} R_0 & \begin{pmatrix} 0 \\ 0 \end{pmatrix} \\
+    0 & 1 \end{pmatrix}, \quad
+g_2 = \begin{pmatrix} R_{\pi/2} & \begin{pmatrix} 1 \\ 0 \end{pmatrix} \\
+    0 & 1 \end{pmatrix}, \\
+g_3 &= \begin{pmatrix} R_{\pi} & \begin{pmatrix}
+    1 \\ 1 \end{pmatrix} \\
+0 & 1 \end{pmatrix}, \quad
+\end{align*}
+
+As data we take the points
+
+\subsection{A simple forward Euler integration on a Lie group.}
+
+For more details on solving differential equations on Lie groups in Julia, see
+\verb|ManifoldDiffEq.jl|.
+
+{\color{red}TODO: Provide code for de-Casteljau algorithm on Lie groups.}
 
 {\color{red}TODO: Brainstorm over one or two examples with code.}
+
 \input{bib.tex}

### 2025-10-28 src/LieGroups.jl
Fix #78

@@ -93,6 +93,7 @@ include("groups/special_galilean_group.jl")
 export AbstractLieGroup
 export LieGroup, LieAlgebra
 export PowerLieGroup, ProductLieGroup
+export NestedPowerRepresentation, NestedReplacingPowerRepresentation
 export LeftSemidirectProductLieGroup, RightSemidirectProductLieGroup
 export ValidationLieGroup
 export DefaultLieAlgebraOrthogonalBasis

### 2025-10-28 paper/paper.tex
a few small improvements

@@
 For example for a function \verb|f(G, g)| that returns a point on the Lie group \verb|G|, there also exists a variant \verb|f!(G, h, g)| that computes the result in-place of \verb|h|, possibly using the memory also for interims results.
+All methods are written such that the result is correct even if arguments \verb|h| and \verb|g| use the same memory.
 By default the allocating variant is implemented by allocating memory accordingly and calling the in-place variant.
@@
 Points and tangent vectors can be represented by the same Julia objects, although they are distinct from a topological viewpoint.%
-\footnote{\texttt{LieGroups.jl} introduces a \texttt{AbstractLieGroupPoint <: AbstractManifoldPoint}, that is used to represent points on the Lie group.
+\footnote{\texttt{LieGroups.jl} introduces an abstract type \texttt{AbstractLieGroupPoint <: AbstractManifoldPoint}, that is used to represent points on the Lie group.
 This abstract supertype is not necessary, points can be represented by arrays or other types as well.}
@@
 \verb|manifold_dimension(G)| to get the dimension of the manifold,
-\verb|project(G, q)|%
+\verb|project(G, q)|, %
 %\footnote{in-place variant \texttt{project!(G, p, q)}.} to project a point onto the manifold, as well as
-\verb|rand(G)| and \verb|rand(rng, G)|%
+\verb|rand(G)| and \verb|rand(rng, G)| %
 %\footnote{in-place variant \texttt{rand!(G, p)} and \texttt{rand!(rng, G, p)}, respectively.}
@@
         \verb|AbelianMultiplicationGroupOperation| & falls back to \verb|*|\\
-        \verb|LeftMatrixMultiplicationGroupOperation| & falls back to \verb|*|\\
+        \verb|MatrixMultiplicationGroupOperation| & falls back to \verb|*|\\
         \verb|PowerGroupOperation{Op}| & \verb|PowerLieGroup|\\
@@
 
-To avoid unnecessary allocations, the identity element can be used without allocations, calling \verb|e = IdentityElement(G)| or \verb|IdentityElement(op)| where \verb|op| is the group operation of \verb|G|.
+The identity element can be represented as a value of a special type, \verb|e = Identity(G)| or \verb|Identity(op)| where \verb|op| is the group operation of \verb|G|.
+This allows for dispatching on the type when defining functions and avoiding unnecessary allocations.
 If the actual value is needed, one can call
@@
 
-\subsubsection*{Exponential and logarithm.}
+\subsubsection*{Exponential and logarithm}
 
-In \verb|LieGroups.jl|, the exponential function is given by \verb|exp(G,X)|%
+In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G,X)|%
 %\footnote{in-place variant \texttt{exp!(G, g, X)}.}

### 2025-10-28 paper/ForwardEulerSE2.jl
Finish a first complete sketch of the SE2 example and create a file for the ForwardEuler.

@@ -1,0 +1,3 @@
+using Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+using LieGroups, ManifoldsBase, RecursiveArrayTools
+using CairoMakie

### 2025-10-28 paper/bezier_example.jl

@@ -1,0 +1,94 @@
+using Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+using LieGroups, ManifoldsBase, RecursiveArrayTools
+using CairoMakie
+"""
+    deCasteljauLieGroup(G, cp, t)
+
+Evaluate the Beziér curve at parameter `t` given
+a vector of control points `cp` on the Lie group `G`.
+"""
+function deCasteljauLieGroup(
+        G::AbstractLieGroup, cp::Vector{P}, t::Real
+    ) where {P}
+    n = length(cp)
+    points = [copy(G, g) for g in cp]
+    for r in 1:n-1
+        for i in 1:n-r
+            X = log(G, points[i], points[i+1])
+            exp!(G, points[i], points[i], t * X)
+        end
+    end
+    return points[1]
+end
+
+R(α) = [cos(α) -sin(α); sin(α) cos(α)]
+α(R) = atan(R[2, 1], R[1, 1])
+SO2 = SpecialOrthogonalGroup(2)
+R2 = TranslationGroup(2)
+SE2 = SpecialEuclideanGroup(2)
+SO2xR2 = SO2 × R2
+
+cp = [
+    ArrayPartition(R(π/2), [0.0, 0.0]),
+    ArrayPartition(R(π/4), [0.33, 0.5]),
+    ArrayPartition(R(3π/4), [0.66, -0.5]),
+    ArrayPartition(R(π), [1.0, 0.0]),
+]
+
+# Start a figure and draw the control points
+fig = Figure(size = (600, 600))
+ax = Axis(fig[1, 1])
+positions = [p.x[2] for p in cp]
+angles = [α(p.x[1]) for p in cp]
+directions = [[cos(a), sin(a)] for a in angles]
+scatter!(ax, [p[1] for p in positions], [p[2] for p in positions], color=:blue, markersize=5)
+arrows2d!(ax,
+    [p[1] for p in positions],
+    [p[2] for p in positions],
+    [d[1] for d in directions],
+    [d[2] for d in directions];
+    lengthscale=0.15, color=:blue
+)
+
+# Evaluate the Beziér curve at multiple t values
+ts1 = range(0, 1, length=35)
+ts1_dense = range(0, 1, length=500)
+bezier1_points = [deCasteljauLieGroup(SE2, cp, t) for t in ts1]
+bezier1_dense_points = [deCasteljauLieGroup(SE2, cp, t) for t in ts1_dense]
[…]

### 2025-10-28 paper/paper.tex
Fix one small error on manifolds for circle.

@@
 \verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively.
-Note that each of these representations uses a different underlying manifold, namely \verb|Circle(|$\mathbb R$\verb|)|, \verb|Circle(|$\mathbb C$\verb|)|, and \verb|Circle(|$\mathbb R^2$\verb|)|, respectively.
+Note that each of these representations uses a different underlying manifold, namely \verb|Circle(|$\mathbb R$\verb|)|, \verb|Circle(|$\mathbb C$\verb|)|, and \verb|Sphere(1)|, respectively.
 This is a case, where it is not only the representation of points and tangent vectors that differ, but indeed even the underlying manifold.
@@
 
-{\color{red}TODO: Brainstorm over one or two examples with code.}
+{\color{red}TODO: Do a Forward Euler simulation on a nice physics problem}
 
@@
 \end{document}
-%
-% ------------------------------------------------------------------------------------------
-%
-
-\subsection{Writing Julia code}
-
-A special environment is already defined for Julia code,
-built on top of \textit{listings} and \textit{jlcode}.
-
-\begin{verbatim}
-\begin{lstlisting}[
-    language = Julia,
-    numbers=left,
-    label={lst:exmplg},
-    caption={Example Code Block.}
-]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-\end{verbatim}
-\begin{lstlisting}[
-    language = Julia,
-    numbers=left,
-    label={lst:exmplg},
-    caption={Example Code Block.}
-]
-using Plots
-
-x = -3.0:0.01:3.0
-y = rand(length(x))
-plot(x, y)
-\end{lstlisting}
-
-\subsection{Balancing column at last page}
-\label{subsub:Balance}
-For balancing the both column length at last page use :
-\begin{verbatim}
-\vadjust{\vfill\pagebreak}
-\end{verbatim}
-
-%\vadjust{\vfill\pagebreak}
-
-at appropriate place in your \TeX{} file or in bibliography file.

### 2025-10-28 ext/LieGroupsRecursiveArrayToolsExt/special_euclidean_group_RAT_ext.jl
Fix #78 (#79)
* Fix #78

* bump version

* getindex ambiguities are fine

* added zenodo fix to news

@@ -168,6 +168,35 @@ function ManifoldsBase.exp!(
     return g
 end
 
+function Base.getindex(
+        g::ArrayPartition,
+        G::SpecialEuclideanGroup,
+        s::Union{Symbol, Int},
+    )
+    return submanifold_component(G, g, s)
+end
+function Base.getindex(
+        g::ArrayPartition,
+        G::SpecialEuclideanGroup,
+        ::Colon,
+    )
+    return submanifold_components(G, g)
+end
+function Base.getindex(
+        g::ArrayPartition,
+        𝔤::LieAlgebra{ℝ, <:LieGroups.SpecialEuclideanGroupOperation, <:SpecialEuclideanGroup},
+        s::Union{Symbol, Int},
+    )
+    return submanifold_component(𝔤, g, s)
+end
+function Base.getindex(
+        g::ArrayPartition,
+        𝔤::LieAlgebra{ℝ, <:LieGroups.SpecialEuclideanGroupOperation, <:SpecialEuclideanGroup},
+        ::Colon,
+    )
+    return submanifold_components(𝔤, g)
+end
+
 function LieGroups.identity_element(
         G::SpecialEuclideanGroup, ::Type{<:SpecialEuclideanProductPoint{A}}
     ) where {A <: ArrayPartition}

### 2025-10-28 src/groups/power_group.jl

@@ -216,6 +216,12 @@ function ManifoldsBase.exp!(
     return h
 end
 
+Base.@propagate_inbounds function Base.getindex(
+        p::AbstractArray, PoG::LieGroup{𝔽, Op, M}, I::Union{Integer, Colon, AbstractVector}...,
+    ) where {𝔽, Op <: PowerGroupOperation, M <: ManifoldsBase.AbstractPowerManifold}
+    return get_component(base_manifold(PoG), p, I...)
+end
+
 function ManifoldsBase.hat!(
         Po𝔤::LieAlgebra{𝔽, Op, LieGroup{𝔽, Op, M}}, X, c
     ) where {𝔽, Op <: PowerGroupOperation, M <: ManifoldsBase.AbstractPowerManifold}

### 2025-10-28 src/groups/special_euclidean_group.jl

@@ -460,7 +460,7 @@ Use `:` to access all submanifold components as a unified tuple.
 
 @doc "$(_doc_getindex_SE)"
 function Base.getindex(
-        g::Union{SpecialEuclideanMatrixPoint, SpecialEuclideanProductPoint, Identity},
+        g::Union{SpecialEuclideanMatrixPoint, SpecialEuclideanProductPoint, Identity, AbstractMatrix},
         G::SpecialEuclideanGroup,
         s::Union{Symbol, Int},
     )
@@ -468,7 +468,7 @@ function Base.getindex(
 end
 
 function Base.getindex(
-        g::Union{SpecialEuclideanMatrixPoint, SpecialEuclideanProductPoint, Identity},
+        g::Union{SpecialEuclideanMatrixPoint, SpecialEuclideanProductPoint, Identity, AbstractMatrix},
         G::SpecialEuclideanGroup,
         ::Colon,
     )
@@ -477,7 +477,7 @@ end
 
 @doc "$(_doc_getindex_SE)"
 function Base.getindex(
-        g::Union{SpecialEuclideanMatrixTangentVector, SpecialEuclideanProductTangentVector},
+        g::Union{SpecialEuclideanMatrixTangentVector, SpecialEuclideanProductTangentVector, AbstractMatrix},
         𝔤::LieAlgebra{ℝ, <:SpecialEuclideanGroupOperation, <:SpecialEuclideanGroup},
         s::Union{Symbol, Int},
     )
@@ -486,7 +486,7 @@ end
 
 @doc "$(_doc_getindex_SE)"
 function Base.getindex(
-        g::Union{SpecialEuclideanMatrixTangentVector, SpecialEuclideanProductTangentVector},
+        g::Union{SpecialEuclideanMatrixTangentVector, SpecialEuclideanProductTangentVector, AbstractMatrix},
         𝔤::LieAlgebra{ℝ, <:SpecialEuclideanGroupOperation, <:SpecialEuclideanGroup},
         ::Colon,
     )

### 2025-10-28 test/groups/test_special_euclidean_group.jl

@@ -72,6 +72,17 @@ using StaticArrays
                 SpecialEuclideanMatrixTangentVector
             @test ManifoldsBase.tangent_vector_type(G, SpecialEuclideanProductPoint) ==
                 SpecialEuclideanProductTangentVector
+
+            # convenience access
+            @test (pts[1])[G, :Translation] isa AbstractVector
+            @test (pts[1])[G, :] isa Tuple
+            @test (pts[1])[G, :Rotation] isa AbstractMatrix
+
+            𝔤 = LieAlgebra(G)
+
+            @test (vec[1])[𝔤, :Translation] isa AbstractVector
+            @test (vec[1])[𝔤, :] isa Tuple
+            @test (vec[1])[𝔤, :Rotation] isa AbstractMatrix
         end
         #
         # Right variant – exchange product cases

### 2025-10-28 test/test_aqua.jl

@@ -6,12 +6,7 @@ using Aqua, LieGroups, Test, LinearAlgebra
         ambiguities = (
             broken = false,
             exclude = [
-                Base.:+, #temporary ambiguities between Manifolds.Identity and LieGroups.Identity
-                Base.:-, #temporary ambiguities between Manifolds.Identity and LieGroups.Identity
-                Base.:*, #temporary ambiguities between Manifolds.Identity and LieGroups.Identity
-                Base.:/, #temporary ambiguities between Manifolds.Identity and LieGroups.Identity
-                Base.:\, #temporary ambiguities between Manifolds.Identity and LieGroups.Identity
-                LinearAlgebra.mul!, #temporary ambiguities between Manifolds.Identity and LieGroups.Identity
+                getindex, # ambiguities from convenience access methods; they were manually verified as safe
             ],
         ),
     )

### 2025-10-28 paper/paper.tex
first csv export.
Finish a first draft of the Bézier example.

@@
 \usepackage{amsmath,booktabs,mathtools,amssymb}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+\pgfplotsset{compat=1.18}
+
 \newcommand{\term}[1]{\emph{#1}}
@@
 \newcommand{\LieGroupsVersion}{0.1.6}
+%
+% Define muted colors.
+\definecolor{TolMutedBlue}{HTML}{332288}
+\definecolor{TolMutedCyan}{HTML}{88CCEE}
+\definecolor{TolMutedTeal}{HTML}{44AA99}
+\definecolor{TolMutedGreen}{HTML}{117733}
+\definecolor{TolMutedOlive}{HTML}{999933}
+\definecolor{TolMutedSand}{HTML}{DDCC77}
+\definecolor{TolMutedRose}{HTML}{CC6677}
+\definecolor{TolMutedWine}{HTML}{882255}
+\definecolor{TolMutedPurple}{HTML}{AA4499}
 %
@@
 
-In Algorithm~\ref{lst:deCasteljauLieGroup}, define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
+In Code~\ref{lst:deCasteljauLieGroup}, we define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
 
-We can now compare the algorithm on the Lie group \(\mathrm{SE}(2)\) with the corresponding Bézier curve on the (direct) product Lie group $\mathrm{SO}(2) \times \mathbb{R}^2$.
+We can now compare the algorithm on the Lie group \(\mathrm{SE}(2)\) with the corresponding Bézier curve on the (direct) product Lie group $\mathrm{SO}(2) \times \mathcal T(2)$, where \(\mathcal T(2) = (\mathbb{R}^2, +)\) is the \verb|TranslationGroup(2)|.
 Denote by $R_\alpha$ the rotation matrix in \(\mathrm{SO}(2)\) for an angle \(\alpha\in[0,2\pi)\).
-Then we consider the control points
+We consider the control points
 \begin{align*}
-    g_1 &= \begin{pmatrix} R_0 & \begin{pmatrix} 0 \\ 0 \end{pmatrix} \\
-    0 & 1 \end{pmatrix}, \quad
-g_2 = \begin{pmatrix} R_{\pi/2} & \begin{pmatrix} 1 \\ 0 \end{pmatrix} \\
-    0 & 1 \end{pmatrix}, \\
-g_3 &= \begin{pmatrix} R_{\pi} & \begin{pmatrix}
-    1 \\ 1 \end{pmatrix} \\
-0 & 1 \end{pmatrix}, \quad
+    g_1 &= \Bigl(R_{\frac{\pi}{2}}, \begin{pmatrix} 0 \\ 0 \end{pmatrix}\Bigr),
+    &\quad
+    g_2 &= \Bigl(R_{\frac{\pi}{4}}, \begin{pmatrix} 1/3 \\ 1/2 \end{pmatrix}\Bigr),
+    \\%\quad
+    g_3 &= \Bigl(R_{\frac{3\pi}{4}}, \begin{pmatrix} 2/3 \\ -1/2 \end{pmatrix}\Bigr),
+    &\quad
+    g_4 &= \Bigl(R_{\pi}, \begin{pmatrix} 1 \\ 0 \end{pmatrix}\Bigr).
 \end{align*}
 
-As data we take the points
+We evaluate the corresponding Bézier curves on both Lie groups using the code in~\ref{lst:bezierLieGroupExample}.
+\begin{lstlisting}[
+    language = Julia,
+    numbers=left,
+    label={lst:bezierLieGroupExample},
+    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).}
+]
+ts = range(0, 1, length=35)
+q1 = [
+    deCasteljauLieGroup(SO2xT2, cp, t) for t in ts
+]
+q2 = [deCasteljauLieGroup(SE2, cp, t) for t in ts]
+\end{lstlisting}
+
+\begin{figure}
+    % Arrow scale factor
+    \def\arrowscale{0.15}
+
+    \begin{tikzpicture}
+        \begin{axis}[
+            grid=major,
+            width=.5\textwidth,
+            legend style={at={(0.95,0.9)}},
+            xmin=-0.05, xmax=1.05,
+            ymin=-0.525, ymax=0.625,
+            unit vector ratio=1 1,
+            title style={yshift=-.5\baselineskip,},
+            title={Two Bézier curves on different Lie groups},
+        ]
+            % Control points
+            % Quiver plot from CSV file
+            \addplot[
+                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
+                -stealth, very thick, TolMutedBlue,
+            ] table[col sep=comma] {data/bezier_example_control_points.csv};
+            \addlegendentry{Control points}
+            % Plot the base points from CSV file
+            \addplot[only marks, mark=*, mark size=2pt, TolMutedBlue, forget plot,
+                ] table[col sep=comma] {data/bezier_example_control_points.csv};
+            % SO2xR2 Curve and quiver
+            \addplot[solid, thick, TolMutedSand, forget plot] table[col sep=comma, x=x, y=y] {data/bezier_example_so2r2_dense_curve.csv};
+            % Quiver plot from CSV file
+            \addplot[
+                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
+                -stealth, TolMutedSand,
+            ] table[col sep=comma] {data/bezier_example_so2r2.csv};
+            \addlegendentry{$\mathrm{SO}(2)\times \mathcal T(2)$}
+            % Plot the base points from CSV file
+            \addplot[only marks, mark=*, mark size=1.5pt, TolMutedSand, forget plot,
+                ] table[col sep=comma] {data/bezier_example_so2r2.csv};
+            % SE2 Curve and quiver
+            \addplot[solid, thick, TolMutedTeal, forget plot] table[col sep=comma, x=x, y=y] {data/bezier_example_se2_dense_curve.csv};
+            % Quiver plot from CSV file
+            \addplot[
+                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
+                -stealth, TolMutedTeal,
+            ] table[col sep=comma] {data/bezier_example_se2.csv};
+            \addlegendentry{$\mathrm{SE}(2)$}
+            % Plot the base points from CSV file
+            \addplot[only marks, mark=*, mark size=1.5pt, TolMutedTeal, forget plot,
+                ] table[col sep=comma] {data/bezier_example_se2.csv};
+        \end{axis}
+    \end{tikzpicture}
+    \caption{Bézier curves on the Lie groups \(\mathrm{SE}(2)\) (teal) and \(\mathrm{SO}(2) \times \mathbb{R}^2\) (sand) based on the same control points
+    $g_1,\ldots,g_4$ (blue).}
+\end{figure}
 

### 2025-10-28 paper/bezier_example.jl

@@ -1,6 +1,15 @@
 using Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+cd(@__DIR__)
 using LieGroups, ManifoldsBase, RecursiveArrayTools
 using CairoMakie
+
+n = 35 # number of points to plot arrows at
+N = 500 # number of points to plot the curve
+export_csv = true
+show_figure = true
+export_folder = "data/"
+experiment_name = "bezier_example"
+
 """
     deCasteljauLieGroup(G, cp, t)
 
@@ -89,6 +98,45 @@ arrows2d!(ax,
     [d[2] for d in bezier2_directions];
     lengthscale=0.15, color=:green, shaftwidth=2.0
 )
-fig
+show_figure && display(fig)
 
+if export_csv
+    using CSV, DataFrames
+
+    df1 = DataFrame(
+        x = [p[1] for p in positions],
+        y = [p[2] for p in positions],
+        u = [d[1] for d in directions],
+        v = [d[2] for d in directions],
+    )
+    CSV.write(export_folder * experiment_name * "_control_points.csv", df1)
+
+    df1 = DataFrame(
+        x = [p[1] for p in bezier1_positions],
+        y = [p[2] for p in bezier1_positions],
+        u = [d[1] for d in bezier1_directions],
+        v = [d[2] for d in bezier1_directions],
+    )
+    CSV.write(export_folder * experiment_name * "_se2.csv", df1)
+
+    df1_dense = DataFrame(
+        x = [p[1] for p in bezier1_dense_positions],
+        y = [p[2] for p in bezier1_dense_positions],
+    )
+    CSV.write(export_folder * experiment_name * "_se2_dense_curve.csv", df1_dense)
+
+    df2 = DataFrame(
+        x = [p[1] for p in bezier2_positions],
+        y = [p[2] for p in bezier2_positions],
+        u = [d[1] for d in bezier2_directions],
+        v = [d[2] for d in bezier2_directions],
+    )
+    CSV.write(export_folder * experiment_name * "_so2r2.csv", df2)
+
+    df2_dense = DataFrame(
+        x = [p[1] for p in bezier2_dense_positions],
+        y = [p[2] for p in bezier2_dense_positions],
+    )
+    CSV.write(export_folder * experiment_name * "_so2r2_dense_curve.csv", df2_dense)
+end
 # export CSV, x,y, u ,v and try to quiver these in TikZ

### 2025-10-28 paper/paper.tex
A few final touhes/formulations for the Bezier example.

@@
 
-\section{An example how to use {\texttt{LieGroups.jl}}}%
+\section{Examples how to use {\texttt{LieGroups.jl}}}%
 \label{sec:Example}
@@
 
-We evaluate the corresponding Bézier curves on both Lie groups using the code in~\ref{lst:bezierLieGroupExample}.
+We evaluate the corresponding Bézier curves on both Lie groups using the Code~\ref{lst:bezierLieGroupExample}.
 \begin{lstlisting}[
@@
     label={lst:bezierLieGroupExample},
-    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).}
+    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).},
+    mathescape
 ]
+SO2 = SpecialOrthogonalGroup(2)
+R2 = TranslationGroup(2)
+SO2xR2 = SO2 $\times$ R2
 ts = range(0, 1, length=35)
@@
 ]
+SE2 = SpecialEuclideanGroup(2)
 q2 = [deCasteljauLieGroup(SE2, cp, t) for t in ts]
@@
     $g_1,\ldots,g_4$ (blue).}
+    \label{fig:Bezier}
 \end{figure}
+
+The resulting Bézier curves on both Lie groups are shown in Figure~\ref{fig:Bezier}.
+One can see that the product Lie group works the same as if one would have done that
+separately component-wise, while the Bézier curve on \(\mathrm{SE}(2)\) behaves differently.
 
@@
 
-{\color{red}TODO: Provide code for de-Casteljau algorithm on Lie groups.}
-
 {\color{red}TODO: Do a Forward Euler simulation on a nice physics problem}

### 2025-10-29 paper/header.tex
Fix 2022 -> 2025.

@@
 pdftitle = {Groups and smooth geometry using LieGroups.jl},
-pdfsubject = {JuliaCon 2022 Proceedings},
+pdfsubject = {JuliaCon 2025 Proceedings},
 pdfauthor = {Ronny Bergmann},

### 2025-10-30 paper/rod_example.jl
rod example work

@@ -4,97 +4,97 @@ using LieGroups, Manifolds, RecursiveArrayTools
 
 # ---------- Physical parameters ----------
 
-ϵ0 = 8.854e-12       # vacuum permittivity (F/m)
-L   = 1.0             # rod length (m)
-M   = 1.0             # rod mass (kg)
-λ   = 1e-6            # linear charge density (C/m)
-Icm = M * L^2 / 12    # moment of inertia about center
-Nq  = 20              # number of sample points along the rod for integration
 
-# Point charges (Coulombs)
+SE2 = SpecialEuclideanGroup(2)
+se2 = LieAlgebra(SE2)
 
-q1, q2 = +1e-6, -1e-6
-R1 = [-1.0, 0.0]      # position of +q1
-R2 = [ 1.0, 0.0]      # position of -q2
+Base.@kwdef struct SystemParameters
+    L = 1.0             # rod length (m)
+    M = 1.0             # rod mass (kg)
+    λ = 1e-6            # linear charge density (C/m)
+    Icm = M * L^2 / 12  # moment of inertia about center
+    qs = [+1e-6, -1e-6] # point charges (Coulombs)
+    Rs = [[-1.0, 0.0], [1.0, 0.0]] # positions of point charges
+    initial_rod_angle = 0.0
+    initial_rod_position = [0.0, -1.0]
+    initial_rod_angular_velocity = 0.1
+    initial_rod_velocity = [0.0, 0.0]
+    tmax::Real = 30.0
+end
+
+ϵ0 = 8.854e-12  # vacuum permittivity (F/m)
+Nq = 20         # number of sample points along the rod for integration
+
 
 # ---------- Electric potential and field ----------
 
-function E_field(pos)
-    r1 = pos .- R1
-    r2 = pos .- R2
-    E = (q1 .* r1 ./ norm(r1)^3 .+ q2 .* r2 ./ norm(r2)^3) ./ (4π*ϵ0)
+function E_field(sp::SystemParameters, pos)
+    E = zeros(2)
+    for i in eachindex(sp.qs)
+        ri = pos .- sp.Rs[i]
+        E .+= (sp.qs[i] .* ri ./ norm(ri)^3) ./ (4π*ϵ0)
+    end
     return E
 end
 
-# ---------- Force and torque on the rod ----------
+# ---------- Equations of motion ----------
+# first block for the paper
 
[…]

### 2025-10-30 src/groups/special_euclidean_group.jl

@@ -323,7 +323,7 @@ function diff_left_compose!(G::SpecialEuclideanGroup, Y, ::Identity, h, X)
     R = submanifold_component(G, h, :Rotation)
     t = submanifold_component(G, h, :Translation)
     Yr .= Xr
-    Yt .= R * Xt .- Xr * t
+    Yt .= R * Xt
     return Y
 end
 

### 2025-10-30 paper/paper.tex
bump version.
Add a final short argument on the code.

@@
 \newcommand{\im}{\ensuremath{\mathrm{i}}}
-\newcommand{\LieGroupsVersion}{0.1.6}
+\newcommand{\LieGroupsVersion}{0.1.7}
 %
@@
 A major difference to the usual representation of tangent vectors on the underlying manifold is, that here the usual representation is done in the Lie algebra.
+
+A central function on the Lie algebra is the Lie bracket $[\cdot, \cdot]\colon \frak g \times \frak g \to \frak g$, which is available as \verb|lie_bracket(|$\frak g$\verb|, X, Y)|.
+% \footnote{in-place variant \texttt{lie\_bracket!($\frak g$, Z, X, Y)}.}
+For more details on the Lie bracket see~\cite[Ch.~5]{HilgertNeeb:2012}.
 
@@
 separately component-wise, while the Bézier curve on \(\mathrm{SE}(2)\) behaves differently.
+A main feature here is, that due to the generic implementation on arbitrary Lie groups,
+we can easily switch between different Lie groups, where the data is still valid.
 

### 2025-10-30 paper/rod_example.jl
rod CSVs

@@ -2,6 +2,8 @@ using LinearAlgebra, Plots
 
 using LieGroups, Manifolds, RecursiveArrayTools
 
+using DataFrames, CSV
+
 # ---------- Physical parameters ----------
 
 
@@ -104,9 +106,6 @@ end
 
 # ---------- Simulation ----------
 function sim()
-    # initial_angles = [0.1, 0.3, 0.6]
-    # initial_positions = [[0.0, -1.5], [0.0, -1.0], [-0.5, -1.5]]
-
     plt = plot(title="Charged rod trajectories near ±q charges",
                xlabel="x (m)", ylabel="y (m)", aspect_ratio=1, legend=false)
     
@@ -118,7 +117,8 @@ function sim()
            SystemParameters(; initial_rod_angle = 0.3, initial_rod_position=[-0.5, +1.0], tmax=100.0),
            SystemParameters(; initial_rod_angle = 0.6, initial_rod_position=[-0.5, -0.5], tmax=40.0, initial_rod_velocity = [-0.01, 0.15])]
 
-    for sp in sps
+    export_folder = "paper/data/"
+    for (isp, sp) in enumerate(sps)
         ang = sp.initial_rod_angle
         pos = sp.initial_rod_position
         centers, Fθs = simulate_trajectory(SOxT, sp, ang, pos, sp.initial_rod_angular_velocity, sp.initial_rod_velocity)
@@ -137,6 +137,16 @@ function sim()
             # @show u
             plot!(plt, [p1[1], p2[1]], [p1[2], p2[2]], color=:black, lw=1, label=false)
         end
+
+        df_dense = DataFrame(x = x[begin:10:end], y = y[begin:10:end])
+        df_rod = DataFrame(
+            up = [centers[i][1] + 0.5sp.L * Fθs[i][1, 1] for i in idxs],
+            vp = [centers[i][1] + 0.5sp.L * Fθs[i][1, 2] for i in idxs],
+            um = [centers[i][2] - 0.5sp.L * Fθs[i][1, 1] for i in idxs],
+            vm = [centers[i][2] - 0.5sp.L * Fθs[i][1, 2] for i in idxs],
+        )
+        CSV.write(export_folder * "ex2_rod_$(isp)_dense.csv", df_dense)
+        CSV.write(export_folder * "ex2_rod_$(isp)_sparse.csv", df_rod)
     end
 
[…]

### 2025-10-30 paper/rod_example.jl
fix sparse data

@@ -140,10 +140,10 @@ function sim()
 
         df_dense = DataFrame(x = x[begin:10:end], y = y[begin:10:end])
         df_rod = DataFrame(
-            up = [centers[i][1] + 0.5sp.L * Fθs[i][1, 1] for i in idxs],
-            vp = [centers[i][1] + 0.5sp.L * Fθs[i][1, 2] for i in idxs],
-            um = [centers[i][2] - 0.5sp.L * Fθs[i][1, 1] for i in idxs],
-            vm = [centers[i][2] - 0.5sp.L * Fθs[i][1, 2] for i in idxs],
+            up = [x[i] + 0.5sp.L * Fθs[i][1, 1] for i in idxs],
+            vp = [y[i] + 0.5sp.L * Fθs[i][1, 2] for i in idxs],
+            um = [x[i] - 0.5sp.L * Fθs[i][1, 1] for i in idxs],
+            vm = [y[i] - 0.5sp.L * Fθs[i][1, 2] for i in idxs],
         )
         CSV.write(export_folder * "ex2_rod_$(isp)_dense.csv", df_dense)
         CSV.write(export_folder * "ex2_rod_$(isp)_sparse.csv", df_rod)

### 2025-10-30 paper/rod_example.jl
4 -> 6; Plots -> Makie

@@ -1,6 +1,8 @@
-using LinearAlgebra, Plots
+using LinearAlgebra
 
 using LieGroups, Manifolds, RecursiveArrayTools
+
+using CairoMakie
 
 using DataFrames, CSV
 
@@ -106,9 +108,13 @@ end
 
 # ---------- Simulation ----------
 function sim()
-    plt = plot(title="Charged rod trajectories near ±q charges",
-               xlabel="x (m)", ylabel="y (m)", aspect_ratio=1, legend=false)
-    
+    fig = Figure(size = (800, 800))
+    ax = Axis(fig[1, 1],
+        title = "Charged rod trajectories near ±q charges",
+        xlabel = "x (m)", ylabel = "y (m)",
+        aspect = DataAspect(), 
+    )
+
     SE2 = SpecialEuclideanGroup(2)
     SOxT = ProductLieGroup(SpecialOrthogonalGroup(2), TranslationGroup(2))
     sp = SystemParameters()
@@ -118,39 +124,47 @@ function sim()
            SystemParameters(; initial_rod_angle = 0.6, initial_rod_position=[-0.5, -0.5], tmax=40.0, initial_rod_velocity = [-0.01, 0.15])]
 
     export_folder = "paper/data/"
+    colors = [:red, :blue, :green]
+
     for (isp, sp) in enumerate(sps)
         ang = sp.initial_rod_angle
         pos = sp.initial_rod_position
         centers, Fθs = simulate_trajectory(SOxT, sp, ang, pos, sp.initial_rod_angular_velocity, sp.initial_rod_velocity)
         x = [c[1] for c in centers]
         y = [c[2] for c in centers]
-        plot!(plt, x, y, label="θ₀=$(ang), pos=$(pos)")
+        lines!(ax, x, y, color = colors[isp])
 
         # Draw rods at 10 roughly equidistant times
         nsteps = length(centers)
         idxs = round.(Int, range(1, nsteps, length=10))
         for i in idxs
             c = centers[i]
-            u = Fθs[i][1, :]           # rod's local x-axis
+            u = Fθs[i][1, :]
             p1 = c .- 0.5sp.L .* u
             p2 = c .+ 0.5sp.L .* u
-            # @show u
[…]

### 2025-10-30 paper/rod_example.jl
fewer points

@@ -145,7 +145,7 @@ function sim()
             lines!(ax, [p1[1], p2[1]], [p1[2], p2[2]], color=:black, linewidth=1)
         end
 
-        df_dense = DataFrame(x = x[begin:10:end], y = y[begin:10:end])
+        df_dense = DataFrame(x = x[begin:500:end], y = y[begin:500:end])
         df_rod = DataFrame(
             x = [x[i] + 0.5sp.L * Fθs[i][1, 1] for i in idxs],
             y = [y[i] + 0.5sp.L * Fθs[i][1, 2] for i in idxs],

### 2025-10-30 paper/rod_example.jl
fix sparse csv

@@ -147,8 +147,8 @@ function sim()
 
         df_dense = DataFrame(x = x[begin:500:end], y = y[begin:500:end])
         df_rod = DataFrame(
-            x = [x[i] + 0.5sp.L * Fθs[i][1, 1] for i in idxs],
-            y = [y[i] + 0.5sp.L * Fθs[i][1, 2] for i in idxs],
+            x = x[idxs],
+            y = y[idxs],
             up = [0.5sp.L * Fθs[i][1, 1] for i in idxs],
             vp = [0.5sp.L * Fθs[i][1, 2] for i in idxs],
             um = [-0.5sp.L * Fθs[i][1, 1] for i in idxs],

### 2025-10-30 paper/rod_example.jl
make sure endpoints are included in dense subsampling

@@ -145,7 +145,7 @@ function sim()
             lines!(ax, [p1[1], p2[1]], [p1[2], p2[2]], color=:black, linewidth=1)
         end
 
-        df_dense = DataFrame(x = x[begin:500:end], y = y[begin:500:end])
+        df_dense = DataFrame(x = vcat(x[begin:500:end], x[end]), y = vcat(y[begin:500:end], y[end]))
         df_rod = DataFrame(
             x = x[idxs],
             y = y[idxs],

### 2025-10-30 paper/paper.tex
Finish the plot.

@@
 
+{\color{red}TODO: Describe the ODE and have 2-3 code blocks}
+
+\begin{figure}
+    % Arrow scale factor
+    \def\arrowscale{1.0}
+
+    \begin{tikzpicture}
+        \begin{axis}[
+            grid=major,
+            width=.5\textwidth,
+            legend style={at={(0.95,0.5)}, anchor=east},
+            unit vector ratio=1 1,
+            xmin = -1, xmax = 8,
+            ymin = -4.25, ymax = 2.25,
+            title style={yshift=-.5\baselineskip,},
+            title={Two Bézier curves on different Lie groups},
+        ] %
+            \addplot[only marks, mark=*, mark size=2pt, color=TolMutedSand] coordinates {(-0.5, 0.0)};
+            \addlegendentry{$q_1$};
+            \addplot[only marks, mark=*, mark size=2pt, color=TolMutedOlive] coordinates {(0.5, 0.0)};
+            \addlegendentry{$q_2$};
+            % Curve and quiver+-
+            \addplot[solid, thick, TolMutedBlue] table[col sep=comma, x=x, y=y] {data/ex2_rod_1_dense.csv};
+            \addlegendentry{$\gamma_1$}
+            % Quiver plot from CSV file
+            \addplot[
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedBlue, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
+            \addplot[
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedBlue, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
+            % Plot the base points from CSV file
+            \addplot[
+                only marks, mark=*, mark size=.5pt, TolMutedBlue, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
+            %
+            %
+            % Curve and quiver+-
+            \addplot[solid, thick, TolMutedGreen] table[col sep=comma, x=x, y=y] {data/ex2_rod_2_dense.csv};
+            \addlegendentry{$\gamma_2$}
+            % Quiver plot from CSV file
+            \addplot[
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedGreen, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
+            \addplot[
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedGreen, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
+            % Plot the base points from CSV file
+            \addplot[
+                only marks, mark=*, mark size=.5pt, TolMutedGreen, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
+            %
+            %
+            % Curve and quiver+-
+            \addplot[solid, thick, TolMutedWine] table[col sep=comma, x=x, y=y] {data/ex2_rod_3_dense.csv};
+            \addlegendentry{$\gamma_3$}
+            % Quiver plot from CSV file
+            \addplot[
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedWine, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
+            \addplot[
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedWine, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
+            % Plot the base points from CSV file
+            \addplot[
+                only marks, mark=*, mark size=.5pt, TolMutedWine, forget plot,
+            ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
+        \end{axis}
+    \end{tikzpicture}
+    \caption{Three solution curves \(\gamma_1\) (blue), \(\gamma_2\) (green), and \(\gamma_3\) (wine) of a Lie group Heun method on the product of Lie groups \(\mathrm{SO}(2) \times \mathbb R^2\)
+    modelling the movement within an electric field with two point charges located at
+    $q_1$ (sand) and $q_2$ (olive). }
+    \label{fig:Heun}
+\end{figure}
+
 For more details on solving differential equations on Lie groups in Julia, see

### 2025-10-30 paper/ForwardEulerSE2.jl

@@ -1,3 +1,0 @@
-using Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
-using LieGroups, ManifoldsBase, RecursiveArrayTools
-using CairoMakie

### 2025-10-30 paper/paper.tex
denser sparse files

@@
 
-\subsection{A simple forward Euler integration on a Lie group.}
+\subsection{Simulating a charged rod in a static electric field}
+
+Physical systems can often be described by ordinary differential equations on Lie groups.
+For example, the state of a charged metal rod moving on a plane can be specified by its position, orientation, linear velocity and angular velocity.
+Such state naturally belongs to a Lie group: a direct product of \(G_1 = \mathrm{SO}(2) \times \mathbb{R}^2\) and its Lie algebra, \(\mathfrak{g}_1\), denoted \(G = G_1 \times \mathfrak{g}_1\).
 
@@
             % Curve and quiver+-
-            \addplot[solid, thick, TolMutedWine] table[col sep=comma, x=x, y=y] {data/ex2_rod_3_dense.csv};
+            \addplot[solid, thick, TolMutedPurple] table[col sep=comma, x=x, y=y] {data/ex2_rod_3_dense.csv};
             \addlegendentry{$\gamma_3$}
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedWine, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedWine, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
@@
             \addplot[
-                only marks, mark=*, mark size=.5pt, TolMutedWine, forget plot,
+                only marks, mark=*, mark size=.5pt, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
@@
     \end{tikzpicture}
-    \caption{Three solution curves \(\gamma_1\) (blue), \(\gamma_2\) (green), and \(\gamma_3\) (wine) of a Lie group Heun method on the product of Lie groups \(\mathrm{SO}(2) \times \mathbb R^2\)
+    \caption{Three solution curves \(\gamma_1\) (blue), \(\gamma_2\) (green), and \(\gamma_3\) (purple) of a Lie group Heun method on the product of Lie groups \(\mathrm{SO}(2) \times \mathbb R^2\)
     modelling the movement within an electric field with two point charges located at

### 2025-10-30 paper/rod_example.jl

@@ -136,7 +136,7 @@ function sim()
 
         # Draw rods at 10 roughly equidistant times
         nsteps = length(centers)
-        idxs = round.(Int, range(1, nsteps, length=10))
+        idxs = round.(Int, range(1, nsteps, length=20))
         for i in idxs
             c = centers[i]
             u = Fθs[i][1, :]

### 2025-10-30 paper/paper.tex
Use slightly asymmetric visualization of the rod to better illustrate its orientation.

@@
 \usepackage{tikz}
+\usetikzlibrary{arrows.meta}
 \usepackage{pgfplots}
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedBlue, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=4pt]}, TolMutedBlue, forget plot,
             ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedBlue, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=2pt]}, TolMutedBlue, forget plot,
             ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedGreen, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=4pt]}, TolMutedGreen, forget plot,
             ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedGreen, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=2pt]}, TolMutedGreen, forget plot,
             ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, TolMutedPurple, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=4pt]}, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, TolMutedPurple, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=2pt]}, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};

### 2025-10-30 paper/paper.tex
Fix a wrong title.

@@
             title style={yshift=-.5\baselineskip,},
-            title={Two Bézier curves on different Lie groups},
+            title={Trajectories of a charged rod between two point charges},
         ] %

### 2025-10-30 paper/paper.tex
example

--- 2025-10-30 paper/paper.tex
@@
 Such state naturally belongs to a Lie group: a direct product of \(G_1 = \mathrm{SO}(2) \times \mathbb{R}^2\) and its Lie algebra, \(\mathfrak{g}_1\), denoted \(G = G_1 \times \mathfrak{g}_1\).
+The ordinary differential equation governing the dynamics of the system can be derived from the electrostatic approximation, in particular the Coulomb's law.
+Given \(N\) electric charges \(q_i, i \in {1, \dots, N}\) at positions \(R_i\) the force and the moment of force affecting the rod can be expressed as
+\begin{align}
+    \ddot{f}
+\end{align}
 

### 2025-10-30 paper/rod_example.jl

@@ -13,7 +13,7 @@ SE2 = SpecialEuclideanGroup(2)
 se2 = LieAlgebra(SE2)
 
 Base.@kwdef struct SystemParameters
-    L = 1.0             # rod length (m)
+    L = 0.67             # rod length (m)
     M = 1.0             # rod mass (kg)
     λ = 1e-6            # linear charge density (C/m)
     Icm = M * L^2 / 12  # moment of inertia about center
@@ -119,9 +119,9 @@ function sim()
     SOxT = ProductLieGroup(SpecialOrthogonalGroup(2), TranslationGroup(2))
     sp = SystemParameters()
 
-    sps = [SystemParameters(; initial_rod_angle = 0.1, initial_rod_position=[0.0, -1.5], tmax=150.0),
-           SystemParameters(; initial_rod_angle = 0.3, initial_rod_position=[-0.5, +1.0], tmax=100.0),
-           SystemParameters(; initial_rod_angle = 0.6, initial_rod_position=[-0.5, -0.5], tmax=40.0, initial_rod_velocity = [-0.01, 0.15])]
+    sps = [SystemParameters(; initial_rod_angle = 0.1, initial_rod_position=[0.0, -1.5], tmax=135.0),
+           SystemParameters(; initial_rod_angle = 0.2, initial_rod_position=[-0.2, +1.0], tmax=30.0, initial_rod_velocity = [-0.12, -0.15]),
+           SystemParameters(; initial_rod_angle = 0.6, initial_rod_position=[-0.5, -0.5], tmax=30.0, initial_rod_velocity = [-0.04, 0.12])]
 
     export_folder = "paper/data/"
     colors = [:red, :blue, :green]

### 2025-10-30 paper/paper.tex
second example

--- 2025-10-30 paper/paper.tex
@@
 \pgfplotsset{compat=1.18}
+
+\usepackage{siunitx}
 
@@
 Such state naturally belongs to a Lie group: a direct product of \(G_1 = \mathrm{SO}(2) \times \mathbb{R}^2\) and its Lie algebra, \(\mathfrak{g}_1\), denoted \(G = G_1 \times \mathfrak{g}_1\).
+The ordinary differential equation governing the dynamics of the system can be derived from the electrostatic approximation, in particular the Coulomb's law.
+Electric point charges \(q_i, i \in {1, \dots, N}\) at positions \(R_i\) generate a static electric field \(E(p)\) at point \(p\).
+The force and the moment of force affecting the rod of length \(L\), mass \(M\), moment of inertia \(I_M\) linear charge density \(\lambda\), with the center at position \(p\), orientation \(\theta\) and where \(u\) is the direction along the rod can be expressed as
+\begin{align}
+    M\ddot{p} &= \lambda \int_{-L/2}^{L/2} E(p + su) ds,  \\
+    I_M \ddot{\theta} &= \int_{-L/2}^{L/2} \lambda s u_{\perp} \cdot E(p + su) ds,
+\end{align}
+where \(u_{\perp}\) is the vector perpendicular to \(u\).
+This can be understood as an ODE on \(G\) and solved using Lie groups methods, such as the RKMK variant of the Heun method~\cite{IserlesMunthe-KaasNoersettZanna:2000}.
 
-{\color{red}TODO: Describe the ODE and have 2-3 code blocks}
+Listing~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
+The function \lstinline{force_and_torque} calculates the force and torque affecting the rod, while \lstinline{dynamics} converts them to the Lie algebra element can be used by the solver.
+\begin{lstlisting}[
+    language = Julia,
+    numbers=left,
+    label={lst:rod-dynamics},
+    caption={Dynamics of a rod affected by point charges.}
+]
+function force_and_torque(sp::SystemParameters, Fθ, r)
+    u, u⊥ = eachrow(Fθ)
+    F, τ = zeros(2), 0.0
+    for s in range(-sp.L/2, sp.L/2; length=Nq)
+        E = E_field(sp, r .+ s .* u)
+        F .+= sp.λ * E * (sp.L/(Nq-1))
+        τ += sp.λ * s * dot(u⊥, E) * (sp.L/(Nq-1))
+    end
+    return F, τ
+end
+
+function dynamics(state, sp::SystemParameters, t)
+    p, dp = state.x
+    F, tau = force_and_torque(sp, p.x...)
+    a = F ./ sp.M
+    coords = [tau / sp.Icm, a...]
+    ddx = hat(se2, coords, ArrayPartition)
+    return ArrayPartition(dp, ddx)
+end
+\end{lstlisting}
+
+The RKMK Heun method is a second order method that can solve ordinary differential equations defined on Lie groups.
+A single step of an RKMK-type Heun solver is implemented in~\ref{lst:heun}.
+The main advantage of using a Lie group solver for this problem is that we do not have to consider parametrization of the Lie group.
+The presented approach can be easily generalized to three spatial dimensions, where parametrization of rotations would be even more challenging numerically.
+
+\begin{lstlisting}[
+    language = Julia,
+    numbers=left,
+    label={lst:heun},
+    caption={Step calculation for the RKMK Heun method.}
+]
+function integrator_step!(A::GroupAction,
+    sp::SystemParameters, y, f, t, dt)
+    Ie = Identity(A.group)
+    F1 = (dt / 2) * f(y, sp, t)
+    tmp = zero_vector(A.manifold, y)
+    diff_group_apply!(A, tmp, Ie, y, F1)
+    y2 = exp(A.manifold, y, tmp)
+    F2 = dt * f(y2, sp, t + dt/2)
+    diff_group_apply!(A, tmp, Ie, y, F2)
+    exp!(base_manifold(A), y, y, tmp)
+end
+\end{lstlisting}
+
+The presented method was applied to a case with two static charges: \SI{1e-6}{C} at position (\SI{-1}{m}, \SI{0}{m}) and \SI{-1e-6}{C} at position (\SI{1}{m}, \SI{0}{m}).
+A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial positions.
+The trajectories obtained in a simulation with time step \SI{0.001}{s} are presented in Figure~\ref{fig:Heun}.
+Simulation times between \SI{30}{s} and \SI{135}{s} were chosen to have the rod remain relatively close to point charges.
+Position and orientation of the rod is indicated at 20 equally spaced times, along with a densely sampled trajectory of the center of the rod.
 
@@
             width=.5\textwidth,
-            legend style={at={(0.95,0.5)}, anchor=east},
+            legend style={at={(0.95,0.25)}, anchor=east},
             unit vector ratio=1 1,
-            xmin = -1, xmax = 8,
-            ymin = -4.25, ymax = 2.25,
+            xmin = -2, xmax = 5,
+            ymin = -3.5, ymax = 3.5,
             title style={yshift=-.5\baselineskip,},
@@
         ] %
-            \addplot[only marks, mark=*, mark size=2pt, color=TolMutedSand] coordinates {(-0.5, 0.0)};
+            \addplot[only marks, mark=*, mark size=2pt, color=TolMutedSand] coordinates {(-1, 0.0)};
             \addlegendentry{$q_1$};
-            \addplot[only marks, mark=*, mark size=2pt, color=TolMutedOlive] coordinates {(0.5, 0.0)};
+            \addplot[only marks, mark=*, mark size=2pt, color=TolMutedOlive] coordinates {(1, 0.0)};
             \addlegendentry{$q_2$};
@@
 
-For more details on solving differential equations on Lie groups in Julia, see
+For more advanced methods of solving differential equations on Lie groups in Julia, see
 \verb|ManifoldDiffEq.jl|.
-
-{\color{red}TODO: Do a Forward Euler simulation on a nice physics problem}
 

### 2025-10-30 paper/paper.tex
Maybe fix UTF8

@@
     label={lst:rod-dynamics},
-    caption={Dynamics of a rod affected by point charges.}
+    caption={Dynamics of a rod affected by point charges.},
+    mathescape
 ]
 function force_and_torque(sp::SystemParameters, Fθ, r)
-    u, u⊥ = eachrow(Fθ)
-    F, τ = zeros(2), 0.0
+    u, u$\bot$ = eachrow(Fθ)
+    F, $\tau$ = zeros(2), 0.0
     for s in range(-sp.L/2, sp.L/2; length=Nq)
         E = E_field(sp, r .+ s .* u)
-        F .+= sp.λ * E * (sp.L/(Nq-1))
-        τ += sp.λ * s * dot(u⊥, E) * (sp.L/(Nq-1))
+        F .+= sp.$\lambda$ * E * (sp.L/(Nq-1))
+        $\tau$ += sp.$\lambda$ * s * dot(u$\bot$, E) * (sp.L/(Nq-1))
     end
-    return F, τ
+    return F, $\tau$
 end

### 2025-10-30 paper/paper.tex
Rephrase a few sentences in 5.2

@@
 Physical systems can often be described by ordinary differential equations on Lie groups.
-For example, the state of a charged metal rod moving on a plane can be specified by its position, orientation, linear velocity and angular velocity.
+For example, the state of a charged metal rod moving on a plane can be specified by its position, orientation, linear velocity, and angular velocity.
 Such state naturally belongs to a Lie group: a direct product of \(G_1 = \mathrm{SO}(2) \times \mathbb{R}^2\) and its Lie algebra, \(\mathfrak{g}_1\), denoted \(G = G_1 \times \mathfrak{g}_1\).
 The ordinary differential equation governing the dynamics of the system can be derived from the electrostatic approximation, in particular the Coulomb's law.
-Electric point charges \(q_i, i \in {1, \dots, N}\) at positions \(R_i\) generate a static electric field \(E(p)\) at point \(p\).
-The force and the moment of force affecting the rod of length \(L\), mass \(M\), moment of inertia \(I_M\) linear charge density \(\lambda\), with the center at position \(p\), orientation \(\theta\) and where \(u\) is the direction along the rod can be expressed as
-\begin{align}
+Electric point charges \(q_i, i \in {1, \dots, N}\) at positions \(R_i\) generate a static electric field \(E(p)\) at every point \(p\) in the plane.
+The force and the moment of force affecting the rod of length \(L\), mass \(M\), moment of inertia \(I_M\) linear charge density \(\lambda\), with the center at position \(p\), orientation \(\theta\), and where \(u\) is the direction along the rod can be expressed as the following two second order ODEs
+\begin{align*}
     M\ddot{p} &= \lambda \int_{-L/2}^{L/2} E(p + su) ds,  \\
     I_M \ddot{\theta} &= \int_{-L/2}^{L/2} \lambda s u_{\perp} \cdot E(p + su) ds,
-\end{align}
-where \(u_{\perp}\) is the vector perpendicular to \(u\).
-This can be understood as an ODE on \(G\) and solved using Lie groups methods, such as the RKMK variant of the Heun method~\cite{IserlesMunthe-KaasNoersettZanna:2000}.
+\end{align*}
+where \(u_{\perp}\) is the vector perpendicular to \(u\). The pair \((p, \theta)\) represents a point on \(G_1\), while \((\dot{p}, \dot{\theta})\) is an element of the Lie algebra \(\mathfrak{g}_1\).
+Hence the second order ODE on \(G_1\) can be converted to a first order ODE on \(G = G_1 \times \mathfrak{g}_1\) by introducing \((\dot{p}, \dot{\theta})\) as additional state variables.
 
-Listing~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
+This can be solved using Lie groups methods, such as the RKMK variant of the Heun method~\cite{IserlesMunthe-KaasNoersettZanna:2000}.
+
+Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
 The function \lstinline{force_and_torque} calculates the force and torque affecting the rod, while \lstinline{dynamics} converts them to the Lie algebra element can be used by the solver.
@@
 The RKMK Heun method is a second order method that can solve ordinary differential equations defined on Lie groups.
-A single step of an RKMK-type Heun solver is implemented in~\ref{lst:heun}.
+A single step of an RKMK-type Heun solver is implemented in Code~\ref{lst:heun}.
 The main advantage of using a Lie group solver for this problem is that we do not have to consider parametrization of the Lie group.

### 2025-10-30 paper/paper.tex
a few smaller fixes.

@@
 The main advantage of using a Lie group solver for this problem is that we do not have to consider parametrization of the Lie group.
-The presented approach can be easily generalized to three spatial dimensions, where parametrization of rotations would be even more challenging numerically.
+The presented approach can be easily generalized to three spatial dimensions, where parametrization of rotations on $\mathrm{SO}(3)$ would be even more challenging numerically.
 
@@
 
-For more advanced methods of solving differential equations on Lie groups in Julia, see
-\verb|ManifoldDiffEq.jl|.
+For more advanced methods of solving differential equations on Manifolds and Lie groups in Julia, see
+\verb|ManifoldDiffEq.jl|.\footnote{see \href{https://juliamanifolds.github.io/ManifoldDiffEq.jl/stable/}{juliamanifolds.github.io/ManifoldDiffEq.jl/stable/}.}
 

### 2025-10-30 paper/paper.tex
make a reference more precise.
A few reformulations, a few questions on the code.

@@
 For example, the state of a charged metal rod moving on a plane can be specified by its position, orientation, linear velocity, and angular velocity.
-Such state naturally belongs to a Lie group: a direct product of \(G_1 = \mathrm{SO}(2) \times \mathbb{R}^2\) and its Lie algebra, \(\mathfrak{g}_1\), denoted \(G = G_1 \times \mathfrak{g}_1\).
+Such state naturally belongs to a Lie group: a direct product of \(\mathcal G_1 = \mathrm{SO}(2) \times \mathbb{R}^2\) and its Lie algebra, \(\mathfrak{g}_1\), denoted \(\mathcal G = \mathcal G_1 \times \mathfrak{g}_1\).
 The ordinary differential equation governing the dynamics of the system can be derived from the electrostatic approximation, in particular the Coulomb's law.
@@
 \end{align*}
-where \(u_{\perp}\) is the vector perpendicular to \(u\). The pair \((p, \theta)\) represents a point on \(G_1\), while \((\dot{p}, \dot{\theta})\) is an element of the Lie algebra \(\mathfrak{g}_1\).
-Hence the second order ODE on \(G_1\) can be converted to a first order ODE on \(G = G_1 \times \mathfrak{g}_1\) by introducing \((\dot{p}, \dot{\theta})\) as additional state variables.
+where \(u_{\perp}\) is the vector perpendicular to \(u\).
+The pair \(g=(p, \theta)\) represents a point on \(\mathcal G_1\), while \(\dot g = (\dot{p}, \dot{\theta})\) is an element of the Lie algebra \(\mathfrak{g}_1\).
+This second order ODE on \(\mathcal G_1\) can be converted to a first order ODE on \(\mathcal G = \mathcal G_1 \times \mathfrak{g}_1\) by introducing \((\dot{p}, \dot{\theta})\) as additional state variables.
 
-This can be solved using Lie groups methods, such as the RKMK variant of the Heun method~\cite{IserlesMunthe-KaasNoersettZanna:2000}.
+The overall ODE has the form
+\begin{equation*}
+    \frac{d}{dt}\begin{pmatrix} g \\ \dot{g} \end{pmatrix}
+    = \begin{pmatrix} \dot{g} \\ f(t, g, \dot{g}) \end{pmatrix},
+\end{equation*}
+where \(g \in \mathcal G_1\) and \(f(t, g, \dot{g})\) is derived from the above equations. We further require initial conditions for both \(g\) and \(\dot{g}\) at time \(t_0=0\).
+Then this can be solved using Lie group methods, such as the Runge-Kutta-Munthe-Kaas (RKMK) variant \cite{Munthe-Kaas:1998} of Heun's method~\cite[App.~A.1]{IserlesMunthe-KaasNoersettZanna:2000}.
 
-Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
+Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE, where \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
 The function \lstinline{force_and_torque} calculates the force and torque affecting the rod, while \lstinline{dynamics} converts them to the Lie algebra element can be used by the solver.
@@
 ]
-function force_and_torque(sp::SystemParameters, Fθ, r)
+function force_and_torque(
+        sp::SystemParameters, Fθ, r
+    )
     u, u$\bot$ = eachrow(Fθ)
@@
 end
-
+# f((p,θ), (dp,dθ), t) including system parameters
 function dynamics(state, sp::SystemParameters, t)
@@
     coords = [tau / sp.Icm, a...]
-    ddx = hat(se2, coords, ArrayPartition)
+    ddx = hat(se2, coords, ArrayPartition) #TODO se2 from global scope?
     return ArrayPartition(dp, ddx)
@@
 A single step of an RKMK-type Heun solver is implemented in Code~\ref{lst:heun}.
+For our special case the \verb|manifold| is equal to the \verb|group|, namely $\mathcal G$, and the group action is the left group operation.
+
 The main advantage of using a Lie group solver for this problem is that we do not have to consider parametrization of the Lie group.
@@
 ]
-function integrator_step!(A::GroupAction,
-    sp::SystemParameters, y, f, t, dt)
+function integrator_step!(
+        A::GroupAction, sp::SystemParameters,
+        y, f, t, dt
+) # TODO how are f and dynamics are they the same related?
     Ie = Identity(A.group)
@@
             unit vector ratio=1 1,
-            xmin = -2, xmax = 5,
-            ymin = -3.5, ymax = 3.5,
+            xmin = -1.5, xmax = 4.5,
+            ymin = -3.25, ymax = 2.75,
             title style={yshift=-.5\baselineskip,},

### 2025-10-30 paper/paper.tex
add labels fig 2

@@
             title={Trajectories of a charged rod between two point charges},
+            xlabel={x [m]},
+            ylabel={y [m]},
         ] %

### 2025-10-31 paper/rod_example.jl
Balance the last page.

@@ -1,9 +1,6 @@
 using LinearAlgebra
-
 using LieGroups, Manifolds, RecursiveArrayTools
-
 using CairoMakie
-
 using DataFrames, CSV
 
 # ---------- Physical parameters ----------
@@ -112,7 +109,7 @@ function sim()
     ax = Axis(fig[1, 1],
         title = "Charged rod trajectories near ±q charges",
         xlabel = "x (m)", ylabel = "y (m)",
-        aspect = DataAspect(), 
+        aspect = DataAspect(),
     )
 
     SE2 = SpecialEuclideanGroup(2)

### 2025-10-31 paper/paper.tex
Balance the last page.
improve second example

@@
     )
-    u, u$\bot$ = eachrow(Fθ)
+    u, u${}_\perp$ = eachrow(Fθ)
     F, $\tau$ = zeros(2), 0.0
@@
         F .+= sp.$\lambda$ * E * (sp.L/(Nq-1))
-        $\tau$ += sp.$\lambda$ * s * dot(u$\bot$, E) * (sp.L/(Nq-1))
+        $\tau$ += sp.$\lambda$ * s * dot(u${}_\perp$, E) * (sp.L/(Nq-1))
     end
@@
     coords = [tau / sp.Icm, a...]
-    ddx = hat(se2, coords, ArrayPartition) #TODO se2 from global scope?
+    ddx = hat(sp.se2, coords, ArrayPartition)
     return ArrayPartition(dp, ddx)
@@
         y, f, t, dt
-) # TODO how are f and dynamics are they the same related?
+) # The dynamics function is passed as f
     Ie = Identity(A.group)
@@
 The presented method was applied to a case with two static charges: \SI{1e-6}{C} at position (\SI{-1}{m}, \SI{0}{m}) and \SI{-1e-6}{C} at position (\SI{1}{m}, \SI{0}{m}).
-A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial positions.
+A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial states:
+\begin{enumerate}
+    \item position (\SI{0}{m}, \SI{-1.5}{m}), angle \SI{0.1}{rad}, velocity (\SI{0}{m/s}, \SI{0}{m/s}), angular velocity \SI{0}{rad/s},
+    \item position (\SI{-0.2}{m}, \SI{1}{m}), angle \SI{0.2}{rad}, velocity (\SI{-0.12}{m/s}, \SI{-0.15}{m/s}), angular velocity \SI{0}{rad/s},
+    \item position (\SI{-0.5}{m}, \SI{-0.5}{m}), angle \SI{0.6}{rad}, velocity (\SI{-0.04}{m/s}, \SI{0.12}{m/s}), angular velocity \SI{0}{rad/s}.
+\end{enumerate}
 The trajectories obtained in a simulation with time step \SI{0.001}{s} are presented in Figure~\ref{fig:Heun}.

### 2025-10-31 paper/rod_example.jl

@@ -8,11 +8,9 @@ using DataFrames, CSV
 
 # ---------- Physical parameters ----------
 
-
-SE2 = SpecialEuclideanGroup(2)
-se2 = LieAlgebra(SE2)
-
 Base.@kwdef struct SystemParameters
+    SE2 = SpecialEuclideanGroup(2)
+    se2 = LieAlgebra(SE2)
     L = 0.67             # rod length (m)
     M = 1.0             # rod mass (kg)
     λ = 1e-6            # linear charge density (C/m)
@@ -59,7 +57,7 @@ function dynamics(state, sp::SystemParameters, t)
     p, dp = state.x
     F, τ = force_and_torque(sp, p.x...)
     a = F ./ sp.M
-    ddx = hat(se2, [τ / sp.Icm, a...], ArrayPartition)
+    ddx = hat(sp.se2, [τ / sp.Icm, a...], ArrayPartition)
     return ArrayPartition(dp, ddx)
 end
 

### 2025-10-31 test/groups/test_metric_group.jl
Start testing on metric lie

@@ -8,4 +8,36 @@ using LieGroupsTestSuite
 
 @testset "MetricLieGroup: A metric decorator for LieGroups" begin
 
+    @testset "Pass through with a dummy metric" begin
+        SO3 = SpecialOrthogonalGroup(3)
+        G = MetricLieGroup(SO3, DummyMetric())
+        g1 = 1 / sqrt(2) * [1.0 1.0; -1.0 1.0]
+        g2 = [0.0 -1.0; 1.0 0.0]
+        g3 = [1.0 0.0; 0.0 1.0]
+        X1, X2, X3 = [0.0 0.1; -0.1 0.0], [0.0 -0.2; 2.0 0.0], [0.0 0.0; 0.0 0.0]
+        properties = Dict(
+            :Name => "The orthogonal group O(2)",
+            :Points => [g1, g2, g3],
+            :Vectors => [X1, X2, X3],
+            :Rng => Random.MersenneTwister(),
+            :Functions => [
+                adjoint,
+                compose,
+                conjugate,
+                diff_inv,
+                diff_left_compose,
+                diff_right_compose,
+                identity_element,
+                inv,
+                inv_left_compose,
+                inv_right_compose,
+                is_identity,
+                lie_bracket,
+                rand,
+                show,
+            ],
+        )
+        expectations = Dict(:repr => "OrthogonalGroup(2)")
+        test_lie_group(G, properties, expectations)
+
 end

### 2025-10-31 paper/paper.tex
minor improvements

@@
 \usepackage{pgfplotstable}
+\usepackage{balance}
 \pgfplotsset{compat=1.18}
@@
     \frac{d}{dt}\begin{pmatrix} g \\ \dot{g} \end{pmatrix}
-    = \begin{pmatrix} \dot{g} \\ f(t, g, \dot{g}) \end{pmatrix},
+    = \begin{pmatrix} \dot{g} \\ f(g, \dot{g}, t) \end{pmatrix},
 \end{equation*}
-where \(g \in \mathcal G_1\) and \(f(t, g, \dot{g})\) is derived from the above equations. We further require initial conditions for both \(g\) and \(\dot{g}\) at time \(t_0=0\).
+where \(g \in \mathcal G_1\) and \(f(g, \dot{g}, t)\) is derived from the above equations. We further require initial conditions for both \(g\) and \(\dot{g}\) at time \(t_0=0\).
 Then this can be solved using Lie group methods, such as the Runge-Kutta-Munthe-Kaas (RKMK) variant \cite{Munthe-Kaas:1998} of Heun's method~\cite[App.~A.1]{IserlesMunthe-KaasNoersettZanna:2000}.
@@
 end
-# f((p,θ), (dp,dθ), t) including system parameters
+# f((p,θ), (dp,dθ), t) where (p,θ) and (dp,dθ) are
+# passed as the first argument, system parameters
+# as the second one and time as the last one
 function dynamics(state, sp::SystemParameters, t)
@@
     a = F ./ sp.M
-    coords = [tau / sp.Icm, a...]
-    ddx = hat(sp.se2, coords, ArrayPartition)
+    c = [tau / sp.Icm, a...]
+    ddx = hat(sp.lie_algebra, c, ArrayPartition)
     return ArrayPartition(dp, ddx)
@@
 
+ \balance
 \input{bib.tex}

### 2025-10-31 paper/rod_example.jl

@@ -6,8 +6,8 @@ using DataFrames, CSV
 # ---------- Physical parameters ----------
 
 Base.@kwdef struct SystemParameters
-    SE2 = SpecialEuclideanGroup(2)
-    se2 = LieAlgebra(SE2)
+    group = ProductLieGroup(SpecialOrthogonalGroup(2), TranslationGroup(2))
+    lie_algebra = LieAlgebra(group)
     L = 0.67             # rod length (m)
     M = 1.0             # rod mass (kg)
     λ = 1e-6            # linear charge density (C/m)
@@ -54,7 +54,7 @@ function dynamics(state, sp::SystemParameters, t)
     p, dp = state.x
     F, τ = force_and_torque(sp, p.x...)
     a = F ./ sp.M
-    ddx = hat(sp.se2, [τ / sp.Icm, a...], ArrayPartition)
+    ddx = hat(sp.lie_algebra, [τ / sp.Icm, a...], ArrayPartition)
     return ArrayPartition(dp, ddx)
 end
 

### 2025-10-31 src/LieGroups.jl
testing all passthrough functions.

@@ -31,7 +31,7 @@ using Manifolds:
     UnitaryMatrices
 using Manifolds: DeterminantOneMatrixType
 using Manifolds: base_manifold
-
+using Manifolds: DefaultMetric
 import LinearAlgebra: adjoint, adjoint!
 using ManifoldsBase:
     AbstractBasis, AbstractMetric, AbstractNumbers,

### 2025-10-31 src/groups/metric_group.jl

@@ -49,32 +49,35 @@ end
 
 Identity(G::MetricLieGroup) = Identity(G.lie_group)
 
-adjoint(G::MetricLieGroup, g, X) = adjoind(G.lie_group, g, X)
+adjoint(G::MetricLieGroup, g, X) = adjoint(G.lie_group, g, X)
 adjoint!(G::MetricLieGroup, Y, g, X) = adjoint!(G.lie_group, Y, g, X)
 
 base_lie_group(G::MetricLieGroup) = G.lie_group
 ManifoldsBase.base_manifold(G::MetricLieGroup) = base_manifold(base_lie_group(G))
 
-_compose(G::MetricLieGroup, k, g, h) = compose(base_lie_group(G), k, g, h)
-_compose!(G::MetricLieGroup, g, h) = compose!(base_lie_group(G), g, h)
+_compose(G::MetricLieGroup, g, h) = compose(base_lie_group(G), g, h)
+_compose!(G::MetricLieGroup, k, g, h) = compose!(base_lie_group(G), k, g, h)
 
 conjugate(G::MetricLieGroup, g, h) = conjugate(base_lie_group(G), g, h)
 conjugate!(G::MetricLieGroup, k, g, h) = conjugate!(base_lie_group(G), k, g, h)
 
-diff_conjugate(G::MetricLieGroup, g, X) = diff_conjugate(base_lie_group(G), g, X)
-diff_conjugate!(G::MetricLieGroup, Y, g, X) = diff_conjugate!(base_lie_group(G), Y, g, X)
+diff_conjugate(G::MetricLieGroup, g, h, X) = diff_conjugate(base_lie_group(G), g, h, X)
+diff_conjugate!(G::MetricLieGroup, Y, g, h, X) = diff_conjugate!(base_lie_group(G), Y, g, h, X)
 
 diff_inv(G::MetricLieGroup, g, X) = diff_inv(base_lie_group(G), g, X)
 diff_inv!(G::MetricLieGroup, Y, g, X) = diff_inv!(base_lie_group(G), Y, g, X)
 
-diff_left_compose(G::MetricLieGroup, g, X) = diff_left_compose(base_lie_group(G), g, X)
-diff_left_compose!(G::MetricLieGroup, Y, g, X) = diff_left_compose!(base_lie_group(G), Y, g, X)
+diff_left_compose(G::MetricLieGroup, g, h, X) = diff_left_compose(base_lie_group(G), g, h, X)
[…]

### 2025-10-31 test/LieGroupsTestSuite.jl/LieGroupsTestSuite.jl

@@ -16,6 +16,7 @@ The following functions are expected to be available, since their defaults just 
 module LieGroupsTestSuite
 using LieGroups
 using ManifoldsBase
+using Manifolds: AbstractMetric
 using Test, Random
 using LinearAlgebra: I
 
@@ -30,6 +31,7 @@ struct DummyManifold <: LieGroups.AbstractManifold{LieGroups.ℝ} end
 struct DummyActionType <: AbstractGroupActionType end
 struct DummyLeftActionType <: AbstractLeftGroupActionType end
 struct DummyRightActionType <: AbstractRightGroupActionType end
+struct DummyMetric <: AbstractMetric end
 const DummyLieGroup = LieGroup{LieGroups.ℝ, DummyOperation, DummyManifold}
 DummyLieGroup() = LieGroup(DummyManifold(), DummyOperation())
 LieGroups.switch(a::DummyActionType) = a

### 2025-10-31 test/groups/test_metric_group.jl

@@ -9,8 +9,10 @@ using LieGroupsTestSuite
 @testset "MetricLieGroup: A metric decorator for LieGroups" begin
 
     @testset "Pass through with a dummy metric" begin
-        SO3 = SpecialOrthogonalGroup(3)
-        G = MetricLieGroup(SO3, DummyMetric())
+        SO2 = SpecialOrthogonalGroup(2)
+        G = MetricLieGroup(SO2, LieGroupsTestSuite.DummyMetric())
+        @test base_lie_group(G) === SO2
+
         g1 = 1 / sqrt(2) * [1.0 1.0; -1.0 1.0]
         g2 = [0.0 -1.0; 1.0 0.0]
         g3 = [1.0 0.0; 0.0 1.0]
@@ -32,12 +34,11 @@ using LieGroupsTestSuite
                 inv_left_compose,
                 inv_right_compose,
                 is_identity,
-                lie_bracket,
                 rand,
                 show,
             ],
         )
-        expectations = Dict(:repr => "OrthogonalGroup(2)")
+        expectations = Dict(:repr => "MetricLieGroup(SpecialOrthogonalGroup(2), LieGroupsTestSuite.DummyMetric())")
         test_lie_group(G, properties, expectations)
-
+    end
 end

### 2025-10-31 src/groups/metric_group.jl
Passthrough for getindex.

@@ -33,15 +33,8 @@ struct MetricLieGroup{
     metric::G
 end
 
-function ManifoldsBase.submanifold_component(G::MetricLieGroup, g, i::Val)
-    return submanifold_component(base_lie_group(G), g, i)
-end
-function ManifoldsBase.submanifold_component(G::MetricLieGroup, g, i::Integer)
-    return submanifold_component(base_lie_group(G), g, i)
-end
-function ManifoldsBase.submanifold_components(G::MetricLieGroup, g)
-    return submanifold_components(G.lie_group, g)
-end
+Base.getindex(g, G::MetricLieGroup, i) = getindex(g, base_lie_group(G), i)
+Base.getindex(g::AbstractArray, G::MetricLieGroup, i) = getindex(g, base_lie_group(G), i)
 
 #
 #
@@ -87,20 +80,8 @@ end
 _inv(G::MetricLieGroup, g) = inv(base_lie_group(G), g)
 _inv!(G::MetricLieGroup, k, g) = inv!(base_lie_group(G), k, g)
 
-_inv_left_compose(G::MetricLieGroup, h, g) = _inv_left_compose(base_lie_group(G), h, g)
-_inv_left_compose!(G::MetricLieGroup, k, h, g) = _inv_left_compose!(base_lie_group(G), k, h, g)
-
-_inv_right_compose(G::MetricLieGroup, h, g) = _inv_right_compose(base_lie_group(G), h, g)
-_inv_right_compose!(G::MetricLieGroup, k, h, g) = _inv_right_compose!(base_lie_group(G), k, h, g)
-
 ManifoldsBase.is_point(G::MetricLieGroup, g; kwargs...) = is_point(base_lie_group(G), g; kwargs...)
 ManifoldsBase.is_point(G::MetricLieGroup, e::Identity; kwargs...) = is_point(base_lie_group(G), e; kwargs...)
-
-function ManifoldsBase.is_vector(
-        G::MetricLieGroup{𝔽, O}, e::Identity{O}, X; kwargs...
-    ) where {𝔽, O <: AbstractGroupOperation}
-    return is_vector(base_lie_group(G), e, X; kwargs...)
-end
 
 function Base.show(io::IO, G::MetricLieGroup)
     return print(io, "MetricLieGroup($(G.lie_group), $(G.metric))")

### 2025-10-31 test/groups/test_metric_group.jl

@@ -41,4 +41,14 @@ using LieGroupsTestSuite
         expectations = Dict(:repr => "MetricLieGroup(SpecialOrthogonalGroup(2), LieGroupsTestSuite.DummyMetric())")
         test_lie_group(G, properties, expectations)
     end
+    @testset "Passthrough for index access" begin
+        SE2 = SpecialEuclideanGroup(2)
+        G = MetricLieGroup(SE2, LieGroupsTestSuite.DummyMetric())
+        g = [1.0 0.0 2.0; 0.0 1.0 3.0; 0.0 0.0 1.0]
+        X = [0.0 -0.1 0.5; 0.1 0.0 1.0; 0.0 0.0 0.0]
+        @test g[SE2, :Rotation] === g[G, :Rotation]
+        @test g[SE2, :Translation] === g[G, :Translation]
+        @test X[SE2, :Rotation] === X[G, :Rotation]
+        @test X[SE2, :Translation] === X[G, :Translation]
+    end
 end

### 2025-10-31 paper/paper.tex
N -> n
Move the figure so it appears a page earlier.
Mark start of trajectories with a dot.

@@
 
-\subsection{A Lie group for Numerical verification}
+\subsection{A Lie group for numerical verification}
 
@@
 
-The presented method was applied to a case with two static charges: \SI{1e-6}{C} at position (\SI{-1}{m}, \SI{0}{m}) and \SI{-1e-6}{C} at position (\SI{1}{m}, \SI{0}{m}).
-A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial states:
-\begin{enumerate}
-    \item position (\SI{0}{m}, \SI{-1.5}{m}), angle \SI{0.1}{rad}, velocity (\SI{0}{m/s}, \SI{0}{m/s}), angular velocity \SI{0}{rad/s},
-    \item position (\SI{-0.2}{m}, \SI{1}{m}), angle \SI{0.2}{rad}, velocity (\SI{-0.12}{m/s}, \SI{-0.15}{m/s}), angular velocity \SI{0}{rad/s},
-    \item position (\SI{-0.5}{m}, \SI{-0.5}{m}), angle \SI{0.6}{rad}, velocity (\SI{-0.04}{m/s}, \SI{0.12}{m/s}), angular velocity \SI{0}{rad/s}.
-\end{enumerate}
-The trajectories obtained in a simulation with time step \SI{0.001}{s} are presented in Figure~\ref{fig:Heun}.
-Simulation times between \SI{30}{s} and \SI{135}{s} were chosen to have the rod remain relatively close to point charges.
-Position and orientation of the rod is indicated at 20 equally spaced times, along with a densely sampled trajectory of the center of the rod.
-
 \begin{figure}
@@
             % Curve and quiver+-
-            \addplot[solid, thick, TolMutedBlue] table[col sep=comma, x=x, y=y] {data/ex2_rod_1_dense.csv};
+            \addplot[solid, thick, TolMutedBlue] table[col sep=comma, x=x, y=y] {data/ex2_rod_1_dense.csv} node[circle, black, inner sep=1.25pt, fill=TolMutedBlue, pos=0.0] {};
             \addlegendentry{$\gamma_1$}
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=4pt]}, TolMutedBlue, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=3pt]}, TolMutedBlue, forget plot,
             ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=2pt]}, TolMutedBlue, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=1.75pt]}, TolMutedBlue, forget plot,
             ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
@@
             \addplot[
-                only marks, mark=*, mark size=.5pt, TolMutedBlue, forget plot,
+                only marks, mark=*, mark size=.0pt, TolMutedBlue, forget plot,
             ] table[col sep=comma] {data/ex2_rod_1_sparse.csv};
@@
             % Curve and quiver+-
-            \addplot[solid, thick, TolMutedGreen] table[col sep=comma, x=x, y=y] {data/ex2_rod_2_dense.csv};
+            \addplot[solid, thick, TolMutedGreen] table[col sep=comma, x=x, y=y] {data/ex2_rod_2_dense.csv}node[circle, black, inner sep=1.25pt, fill=TolMutedGreen, pos=0.0] {};
             \addlegendentry{$\gamma_2$}
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=4pt]}, TolMutedGreen, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=3pt]}, TolMutedGreen, forget plot,
             ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=2pt]}, TolMutedGreen, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=1.75pt]}, TolMutedGreen, forget plot,
             ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
@@
             \addplot[
-                only marks, mark=*, mark size=.5pt, TolMutedGreen, forget plot,
+                only marks, mark=*, mark size=.0pt, TolMutedGreen, forget plot,
             ] table[col sep=comma] {data/ex2_rod_2_sparse.csv};
@@
             % Curve and quiver+-
-            \addplot[solid, thick, TolMutedPurple] table[col sep=comma, x=x, y=y] {data/ex2_rod_3_dense.csv};
+            \addplot[solid, thick, TolMutedPurple] table[col sep=comma, x=x, y=y] {data/ex2_rod_3_dense.csv} node[circle, black, inner sep=1.25pt, fill=TolMutedPurple, pos=0.0] {};
             \addlegendentry{$\gamma_3$}
@@
             \addplot[
-                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=4pt]}, TolMutedPurple, forget plot,
+                quiver={u=\thisrow{up}, v=\thisrow{vp}, scale arrows=\arrowscale}, -{Bar[width=3pt]}, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
             \addplot[
-                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=2pt]}, TolMutedPurple, forget plot,
+                quiver={u=\thisrow{um}, v=\thisrow{vm}, scale arrows=\arrowscale}, -{Circle[length=1.75pt]}, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
@@
             \addplot[
-                only marks, mark=*, mark size=.5pt, TolMutedPurple, forget plot,
+                only marks, mark=*, mark size=.0pt, TolMutedPurple, forget plot,
             ] table[col sep=comma] {data/ex2_rod_3_sparse.csv};
@@
     modelling the movement within an electric field with two point charges located at
-    $q_1$ (sand) and $q_2$ (olive). }
+    $q_1$ (sand) and $q_2$ (olive).
+    The initial start points of the rods are indicated by circles, the rods themselves have two
+    different end markers to indicate their orientation.
+    }
     \label{fig:Heun}
 \end{figure}
+
+The presented method was applied to a case with two static charges: \SI{1e-6}{C} at position (\SI{-1}{m}, \SI{0}{m}) and \SI{-1e-6}{C} at position (\SI{1}{m}, \SI{0}{m}).
+A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial states:
+\begin{enumerate}
+    \item position (\SI{0}{m}, \SI{-1.5}{m}), angle \SI{0.1}{rad}, velocity (\SI{0}{m/s}, \SI{0}{m/s}), angular velocity \SI{0}{rad/s},
+    \item position (\SI{-0.2}{m}, \SI{1}{m}), angle \SI{0.2}{rad}, velocity (\SI{-0.12}{m/s}, \SI{-0.15}{m/s}), angular velocity \SI{0}{rad/s},
+    \item position (\SI{-0.5}{m}, \SI{-0.5}{m}), angle \SI{0.6}{rad}, velocity (\SI{-0.04}{m/s}, \SI{0.12}{m/s}), angular velocity \SI{0}{rad/s}.
+\end{enumerate}
+The trajectories obtained in a simulation with time step \SI{0.001}{s} are presented in Figure~\ref{fig:Heun}.
+Simulation times between \SI{30}{s} and \SI{135}{s} were chosen to have the rod remain relatively close to point charges.
+Position and orientation of the rod is indicated at 20 equally spaced times, along with a densely sampled trajectory of the center of the rod.
 

### 2025-10-31 src/groups/metric_group.jl
fix element access.

@@ -35,6 +35,18 @@ end
 
 Base.getindex(g, G::MetricLieGroup, i) = getindex(g, base_lie_group(G), i)
 Base.getindex(g::AbstractArray, G::MetricLieGroup, i) = getindex(g, base_lie_group(G), i)
+function Base.getindex(
+        X, 𝔤::LieAlgebra{𝔽, <:O, <:MetricLieGroup}, i
+    ) where {𝔽, O <: AbstractGroupOperation}
+    # unwrap Algebra and metric decorator
+    return getindex(X, LieAlgebra(base_lie_group(base_lie_group(𝔤))), i)
+end
+function Base.getindex(
+        X::AbstractArray, 𝔤::LieAlgebra{𝔽, <:O, <:MetricLieGroup}, i
+    ) where {𝔽, O <: AbstractGroupOperation}
+    # unwrap Algebra and metric decorator
+    return getindex(X, LieAlgebra(base_lie_group(base_lie_group(𝔤))), i)
+end
 
 #
 #

### 2025-10-31 test/groups/test_metric_group.jl

@@ -26,6 +26,7 @@ using LieGroupsTestSuite
                 adjoint,
                 compose,
                 conjugate,
+                diff_conjugate,
                 diff_inv,
                 diff_left_compose,
                 diff_right_compose,
@@ -43,12 +44,20 @@ using LieGroupsTestSuite
     end
     @testset "Passthrough for index access" begin
         SE2 = SpecialEuclideanGroup(2)
+        se2 = LieAlgebra(SE2)
         G = MetricLieGroup(SE2, LieGroupsTestSuite.DummyMetric())
+        𝔤 = LieAlgebra(G)
         g = [1.0 0.0 2.0; 0.0 1.0 3.0; 0.0 0.0 1.0]
         X = [0.0 -0.1 0.5; 0.1 0.0 1.0; 0.0 0.0 0.0]
         @test g[SE2, :Rotation] === g[G, :Rotation]
         @test g[SE2, :Translation] === g[G, :Translation]
-        @test X[SE2, :Rotation] === X[G, :Rotation]
-        @test X[SE2, :Translation] === X[G, :Translation]
+        @test X[se2, :Rotation] === X[𝔤, :Rotation]
+        @test X[se2, :Translation] === X[𝔤, :Translation]
+        gT = SpecialEuclideanMatrixPoint(g)
+        XT = SpecialEuclideanMatrixTangentVector(X)
+        @test gT[SE2, :Rotation] == g[G, :Rotation]
+        @test gT[SE2, :Translation] == g[G, :Translation]
+        @test XT[se2, :Rotation] == XT[𝔤, :Rotation]
+        @test XT[se2, :Translation] == XT[𝔤, :Translation]
     end
 end

### 2025-10-31 src/groups/semidirect_product_group.jl
Lots of small fixes and diff_left_compose on SE(n)

@@ -397,7 +397,7 @@ function _compose!(
     # this is inplace if both are not aliased and creates a copy G otherwise to avoid overwriting hG
     _semidirect_maybe_inv!(AO, a, G, kG, submanifold_component(SDPG, g, Val(g_ind)))
     # a) group action (first to avoid side effects in g, set kH to σ_{gG}(hH) - since we might have inverted, we have to use kG
-    apply!(a, kH, kG, submanifold_component(PM, h, h_ind)) #accidentially overwriting hH is fine, we do not need it.
+    apply!(a, kH, kG, submanifold_component(PM, h, h_ind)) #accidentally overwriting hH is fine, we do not need it.
     # b) group operation on G
     _compose!(G, submanifold_component(PM, k, g_ind), submanifold_component(PM, g, g_ind), submanifold_component(PM, h, g_ind))
     # c) group operation on H (note that the action on hH is already in kH

### 2025-10-31 src/groups/special_euclidean_group.jl

@@ -315,6 +315,49 @@ function default_right_action(::TranslationGroup, ::SpecialOrthogonalGroup)
     return LeftMultiplicationGroupAction()
 end
 
+@doc raw"""
+    diff_left_compose(G::SpecialEuclideanGroup, ::Identity, h, X)
+
+Compute the differential of left composition by `h` on Lie-algebra elements.
+
+Let
+````
+h=\begin{pmatrix} R & t\\[4pt] 0 & 1\end{pmatrix},\qquad
+X=\begin{pmatrix} X_r & X_t\\[4pt] 0 & 0\end{pmatrix},
+````
+
+where ``R\in SO(n)``, ``t\in\mathbb{R}^n``, ``X_r`` is the skew-symmetric rotation block and ``X_t`` the translation column.
+Then the differential is the adjoint action by ``h^{-1}``:
+````
+\mathrm{D}L_h(X)=h^{-1}Xh=
+\begin{pmatrix}
+R^\top X_r R & R^\top\bigl(X_r\,t + X_t\bigr)\\[4pt]
+0 & 0
+\end{pmatrix}.
+````
+
+Component-wise:
+````
+Y_r = R^\top X_r R,\qquad Y_t = R^\top\bigl(X_r\,t + X_t\bigr).
+````
+"""
+diff_left_compose(G::SpecialEuclideanGroup, ::Identity, h, X)
+
+function diff_left_compose!(G::SpecialEuclideanGroup, Y, ::Identity, h, X)
+    GA = LieAlgebra(G)
+    init_constants!(GA, Y)
+    Xr = submanifold_component(GA, X, Val(:Rotation))
+    Xt = submanifold_component(GA, X, Val(:Translation))
+    Yr = submanifold_component(GA, Y, Val(:Rotation))
+    Yt = submanifold_component(GA, Y, Val(:Translation))
+    R = submanifold_component(G, h, Val(:Rotation))
+    t = submanifold_component(G, h, Val(:Translation))
+    A = R' * Xr
+    mul!(Yr, A, R)
+    Yt .= A * t .+ R' * Xt
+    return Y
+end
+
 _doc_exp_SE2_id = """
     exp(G::SpecialEuclidean, X)
     exp!(G::SpecialEuclidean, g, X)
@@ -538,6 +581,10 @@ function init_constants!(G::SpecialEuclideanGroup, g::AbstractMatrix)
     g[n + 1, n + 1] = 1
     return g
 end
[…]

### 2025-10-31 src/interface.jl

@@ -417,6 +417,10 @@ function diff_left_compose(G::AbstractLieGroup, g, h, X)
     Y = ManifoldsBase.allocate_result(G, diff_left_compose, X, g, h)
     return diff_left_compose!(G, Y, g, h, X)
 end
+function diff_left_compose(G::AbstractLieGroup{𝔽, O}, g::Identity{O}, h, X) where {𝔽, O <: AbstractGroupOperation}
+    Y = ManifoldsBase.allocate_result(G, diff_left_compose, X, h)
+    return diff_left_compose!(G, Y, g, h, X)
+end
 
 function diff_left_compose! end
 @doc "$(_doc_diff_left_compose)"
@@ -1237,6 +1241,10 @@ function _vector_transport_to_basemanifold(
     # now Y is in ThM so we still ave to pull it back to TeM using
     # the left group opp diff.
     return pull_back_tangent(G, g, X)
+end
+
+function ManifoldsBase.zero_vector(G::AbstractLieGroup{𝔽, O}, ::Identity{O}) where {𝔽, O <: AbstractGroupOperation}
+    return zero_vector(LieAlgebra(G))
 end
 
 #

### 2025-10-31 test/groups/test_special_euclidean_group.jl

@@ -134,7 +134,7 @@ using StaticArrays
         Y3 = [0.0 0.1 0.0 1.0; -0.1 0.0 0.0 0.0; 0.0 0.0 0.0 0.0; 0.0 0.0 0.0 0.0]
         YL1 = ArrayPartition([0.0 -0.23 0.0; 0.23 0.0 0.0; 0.0 0.0 0.0], [0.0, 1.0, 0.0])
         YL2 = ArrayPartition([0.0 0.3 0.0; -0.3 0.0 0.0; 0.0 0.0 0.0], [1.0, 1.0, 0.0])
-        YL3 = ArrayPartition([9.0 0.1 0.0; -0.1 0.0 0.0; 0.0 0.0 0.0], [1.0, 0.0, 0.0])
+        YL3 = ArrayPartition([0.0 0.1 0.0; -0.1 0.0 0.0; 0.0 0.0 0.0], [1.0, 0.0, 0.0])
         hA = [h1, h2, h3]
         hM = SpecialEuclideanMatrixPoint.(hA)
         hP = [hL1, hL2, hL3]
@@ -143,6 +143,10 @@ using StaticArrays
         YM = SpecialEuclideanMatrixTangentVector.(YA)
         YP = [YL1, YL2, YL3]
         YQ = SpecialEuclideanProductTangentVector.(YP)
+
+        GL4 = GeneralLinearGroup(4)
+        expected_diff_left_compose = [diff_left_compose(GL4, Identity(GL4), hi, Yi) for (hi, Yi) in zip(hA, YA)]
+
         for G in [G3f, G3p], (pts, vec) in zip([hA, hM, hP, hQ], [YA, YM, YP, YQ])
             properties = Dict(
                 :Name => "The special Euclidean group ($G, $(eltype(pts)))",
@@ -155,6 +159,13 @@ using StaticArrays
                 :repr => "SpecialEuclideanGroup(3)", :atol => 1.0e-14, :is_flat => false
             )
             test_lie_group(G, properties, expectations)
+
+            @testset "diff_left_compose" begin
+                for i in eachindex(vec)
+                    Y = convert(SpecialEuclideanMatrixTangentVector, diff_left_compose(G, Identity(G), pts[i], vec[i]))
+                    @test Y.value ≈ expected_diff_left_compose[i]
+                end
+            end
         end
     end
     #

### 2025-10-31 test/groups/test_metric_group.jl
Improve test coverage

@@ -55,8 +55,8 @@ using LieGroupsTestSuite
         @test X[se2, :Translation] === X[𝔤, :Translation]
         gT = SpecialEuclideanMatrixPoint(g)
         XT = SpecialEuclideanMatrixTangentVector(X)
-        @test gT[SE2, :Rotation] == g[G, :Rotation]
-        @test gT[SE2, :Translation] == g[G, :Translation]
+        @test gT[SE2, :Rotation] == gT[G, :Rotation]
+        @test gT[SE2, :Translation] == gT[G, :Translation]
         @test XT[se2, :Rotation] == XT[𝔤, :Rotation]
         @test XT[se2, :Translation] == XT[𝔤, :Translation]
     end

### 2025-10-31 src/group_operations/addition_operation.jl
more methods and stuff

@@ -86,6 +86,7 @@ Here it simplifies for [`AdditionGroupOperation`](@ref) to ``$(_math(:d))ρ_h(g)
 
 @doc "$(_doc_diff_left_compose_add)"
 diff_left_compose(G::LieGroup{𝔽, AdditionGroupOperation}, g, h, X) where {𝔽} = X
+diff_left_compose(G::LieGroup{𝔽, AdditionGroupOperation}, ::Identity{AdditionGroupOperation}, h, X) where {𝔽} = X
 
 @doc "$(_doc_diff_left_compose_add)"
 function diff_left_compose!(G::LieGroup{𝔽, AdditionGroupOperation}, Y, g, h, X) where {𝔽}

### 2025-10-31 src/groups/circle_group_real.jl

@@ -88,8 +88,6 @@ function diff_inv!(G::_RealCircleGroup, Y, g, X)
     return copyto!(LieAlgebra(G), Y, -X)
 end
 
-diff_left_compose(::_RealCircleGroup, g, h, X::Number) = X
-
 diff_right_compose(::_RealCircleGroup, g, h, X::Number) = X
 
 _doc_exp_real_circ = """

### 2025-10-31 src/groups/special_euclidean_group.jl

@@ -318,32 +318,33 @@ end
 @doc raw"""
     diff_left_compose(G::SpecialEuclideanGroup, ::Identity, h, X)
 
-Compute the differential of left composition by `h` on Lie-algebra elements.
+Compute the differential of left composition by `h` on [`SpecialEuclideanGroup`](@ref)
+for tangent vector `X` at `g`.
 
 Let
-````
+```math
 h=\begin{pmatrix} R & t\\[4pt] 0 & 1\end{pmatrix},\qquad
 X=\begin{pmatrix} X_r & X_t\\[4pt] 0 & 0\end{pmatrix},
-````
+```
 
 where ``R\in SO(n)``, ``t\in\mathbb{R}^n``, ``X_r`` is the skew-symmetric rotation block and ``X_t`` the translation column.
 Then the differential is the adjoint action by ``h^{-1}``:
-````
-\mathrm{D}L_h(X)=h^{-1}Xh=
+```math
+\mathrm{D}λ_h(X)=h^{-1}Xh=
 \begin{pmatrix}
 R^\top X_r R & R^\top\bigl(X_r\,t + X_t\bigr)\\[4pt]
 0 & 0
 \end{pmatrix}.
-````
+```
 
 Component-wise:
-````
+```math
 Y_r = R^\top X_r R,\qquad Y_t = R^\top\bigl(X_r\,t + X_t\bigr).
-````
+```
 """
-diff_left_compose(G::SpecialEuclideanGroup, ::Identity, h, X)
+diff_left_compose(G::SpecialEuclideanGroup, g, h, X)
 
-function diff_left_compose!(G::SpecialEuclideanGroup, Y, ::Identity, h, X)
+function diff_left_compose!(G::SpecialEuclideanGroup, Y, g, h, X)
     GA = LieAlgebra(G)
     init_constants!(GA, Y)
     Xr = submanifold_component(GA, X, Val(:Rotation))
@@ -355,6 +356,20 @@ function diff_left_compose!(G::SpecialEuclideanGroup, Y, ::Identity, h, X)
     A = R' * Xr
     mul!(Yr, A, R)
     Yt .= A * t .+ R' * Xt
+    return Y
+end
+
+
+@doc raw"""
+    diff_right_compose(G::SpecialEuclideanGroup, g, h, X)
+
+Compute the differential of right composition by `h` on [`SpecialEuclideanGroup`](@ref)
[…]

### 2025-10-31 src/groups/validation_group.jl

@@ -301,6 +301,15 @@ function diff_left_compose(G::ValidationLieGroup, g, h, X; kwargs...)
     is_point(LieAlgebra(G), Y; within = diff_left_compose, context = (:Output,), kwargs...)
     return ValidationLieAlgebraTangentVector(Y)
 end
+function diff_left_compose(G::ValidationLieGroup{𝔽, O}, g::Identity{O}, h, X; kwargs...) where {𝔽, O <: AbstractGroupOperation}
+    is_point(G, h; within = diff_left_compose, context = (:Input,), kwargs...)
+    is_point(LieAlgebra(G), X; within = diff_left_compose, context = (:Input,), kwargs...)
+    Y = diff_left_compose(
+        G.lie_group, g, unwrap_validation(h), unwrap_validation(X)
+    )
+    is_point(LieAlgebra(G), Y; within = diff_left_compose, context = (:Output,), kwargs...)
+    return ValidationLieAlgebraTangentVector(Y)
+end
 function diff_left_compose!(G::ValidationLieGroup, Y, g, h, X; kwargs...)
     is_point(G, g; within = diff_left_compose, context = (:Input,), kwargs...)
     is_point(G, h; within = diff_left_compose, context = (:Input,), kwargs...)

### 2025-10-31 src/interface.jl

@@ -459,6 +459,10 @@ function diff_right_compose(G::AbstractLieGroup, g, h, X)
     Y = ManifoldsBase.allocate_result(G, diff_right_compose, X, g, h)
     return diff_right_compose!(G, Y, g, h, X)
 end
+function diff_right_compose(G::AbstractLieGroup, g::Identity, h, X)
+    Y = ManifoldsBase.allocate_result(G, diff_right_compose, X, h)
+    return diff_right_compose!(G, Y, g, h, X)
+end
 
 function diff_right_compose! end
 @doc "$(_doc_diff_right_compose)"

### 2025-10-31 test/groups/test_special_euclidean_group.jl

@@ -145,7 +145,11 @@ using StaticArrays
         YQ = SpecialEuclideanProductTangentVector.(YP)
 
         GL4 = GeneralLinearGroup(4)
-        expected_diff_left_compose = [diff_left_compose(GL4, Identity(GL4), hi, Yi) for (hi, Yi) in zip(hA, YA)]
+        expected_diff_left_compose_id = [diff_left_compose(GL4, Identity(GL4), hi, Yi) for (hi, Yi) in zip(hA, YA)]
+        expected_diff_right_compose_id = [diff_right_compose(GL4, Identity(GL4), hi, Yi) for (hi, Yi) in zip(hA, YA)]
+
+        expected_diff_left_compose_h1 = [diff_left_compose(GL4, h1, hi, Yi) for (hi, Yi) in zip(hA, YA)]
+        expected_diff_right_compose_h1 = [diff_right_compose(GL4, h1, hi, Yi) for (hi, Yi) in zip(hA, YA)]
 
         for G in [G3f, G3p], (pts, vec) in zip([hA, hM, hP, hQ], [YA, YM, YP, YQ])
             properties = Dict(
@@ -163,7 +167,16 @@ using StaticArrays
             @testset "diff_left_compose" begin
                 for i in eachindex(vec)
                     Y = convert(SpecialEuclideanMatrixTangentVector, diff_left_compose(G, Identity(G), pts[i], vec[i]))
-                    @test Y.value ≈ expected_diff_left_compose[i]
+                    @test Y.value ≈ expected_diff_left_compose_id[i]
+
+                    Y = convert(SpecialEuclideanMatrixTangentVector, diff_right_compose(G, Identity(G), pts[i], vec[i]))
+                    @test Y.value ≈ expected_diff_right_compose_id[i]
+
+                    Y = convert(SpecialEuclideanMatrixTangentVector, diff_left_compose(G, pts[1], pts[i], vec[i]))
+                    @test Y.value ≈ expected_diff_left_compose_h1[i]
+
+                    Y = convert(SpecialEuclideanMatrixTangentVector, diff_right_compose(G, pts[1], pts[i], vec[i]))
+                    @test Y.value ≈ expected_diff_right_compose_h1[i]
                 end
             end
         end

### 2025-10-31 src/groups/circle_group_real.jl
disambiguate stuff

@@ -88,8 +88,6 @@ function diff_inv!(G::_RealCircleGroup, Y, g, X)
     return copyto!(LieAlgebra(G), Y, -X)
 end
 
-diff_right_compose(::_RealCircleGroup, g, h, X::Number) = X
-
 _doc_exp_real_circ = """
     exp(::LieGroup{ℝ, AdditionGroupOperation, Circle{ℝ}}, X)
     exp!(::LieGroup{ℝ, AdditionGroupOperation, Circle{ℝ}}, g, X)

### 2025-10-31 src/groups/special_euclidean_group.jl

@@ -316,7 +316,7 @@ function default_right_action(::TranslationGroup, ::SpecialOrthogonalGroup)
 end
 
 @doc raw"""
-    diff_left_compose(G::SpecialEuclideanGroup, ::Identity, h, X)
+    diff_left_compose(G::SpecialEuclideanGroup, g, h, X)
 
 Compute the differential of left composition by `h` on [`SpecialEuclideanGroup`](@ref)
 for tangent vector `X` at `g`.
@@ -324,22 +324,22 @@ for tangent vector `X` at `g`.
 Let
 ```math
 h=\begin{pmatrix} R & t\\[4pt] 0 & 1\end{pmatrix},\qquad
-X=\begin{pmatrix} X_r & X_t\\[4pt] 0 & 0\end{pmatrix},
+X=\begin{pmatrix} X_{\mathrm{R}} & X_{\mathrm{t}}\\[4pt] 0 & 0\end{pmatrix},
 ```
 
-where ``R\in SO(n)``, ``t\in\mathbb{R}^n``, ``X_r`` is the skew-symmetric rotation block and ``X_t`` the translation column.
+where ``R\in SO(n)``, ``t\in\mathbb{R}^n``, ``X_{\mathrm{R}}`` is the skew-symmetric rotation block and ``X_{\mathrm{t}}`` the translation column.
 Then the differential is the adjoint action by ``h^{-1}``:
 ```math
 \mathrm{D}λ_h(X)=h^{-1}Xh=
 \begin{pmatrix}
-R^\top X_r R & R^\top\bigl(X_r\,t + X_t\bigr)\\[4pt]
+R^\top X_{\mathrm{R}} R & R^\top\bigl(X_{\mathrm{R}}\,t + X_{\mathrm{t}}\bigr)\\[4pt]
 0 & 0
 \end{pmatrix}.
 ```
 
 Component-wise:
 ```math
-Y_r = R^\top X_r R,\qquad Y_t = R^\top\bigl(X_r\,t + X_t\bigr).
+Y_{\mathrm{R}} = R^\top X_{\mathrm{R}} R,\qquad Y_{\mathrm{t}} = R^\top\bigl(X_{\mathrm{R}}\,t + X_{\mathrm{t}}\bigr).
 ```
 """
 diff_left_compose(G::SpecialEuclideanGroup, g, h, X)
@@ -347,14 +347,14 @@ diff_left_compose(G::SpecialEuclideanGroup, g, h, X)
[…]

### 2025-10-31 src/groups/validation_group.jl

@@ -325,6 +325,15 @@ function diff_left_compose!(G::ValidationLieGroup, Y, g, h, X; kwargs...)
     return Y
 end
 
+function diff_right_compose(G::ValidationLieGroup{𝔽, O}, g::Identity{O}, h, X; kwargs...) where {𝔽, O <: AbstractGroupOperation}
+    is_point(G, h; within = diff_right_compose, context = (:Input,), kwargs...)
+    is_point(LieAlgebra(G), X; within = diff_right_compose, context = (:Input,), kwargs...)
+    Y = diff_right_compose(
+        G.lie_group, g, unwrap_validation(h), unwrap_validation(X)
+    )
+    is_point(LieAlgebra(G), Y; within = diff_right_compose, context = (:Output,), kwargs...)
+    return ValidationLieAlgebraTangentVector(Y)
+end
 function diff_right_compose(G::ValidationLieGroup, g, h, X; kwargs...)
     is_point(G, g; within = diff_right_compose, context = (:Input,), kwargs...)
     is_point(G, h; within = diff_right_compose, context = (:Input,), kwargs...)

### 2025-10-31 src/interface.jl

@@ -459,7 +459,8 @@ function diff_right_compose(G::AbstractLieGroup, g, h, X)
     Y = ManifoldsBase.allocate_result(G, diff_right_compose, X, g, h)
     return diff_right_compose!(G, Y, g, h, X)
 end
-function diff_right_compose(G::AbstractLieGroup, g::Identity, h, X)
+
+function diff_right_compose(G::AbstractLieGroup{𝔽, O}, g::Identity{O}, h, X) where {𝔽, O <: AbstractGroupOperation}
     Y = ManifoldsBase.allocate_result(G, diff_right_compose, X, h)
     return diff_right_compose!(G, Y, g, h, X)
 end

### 2025-10-31 paper/paper.tex
Add a metric Lie group, fine tuning of code env would be neccessary.

@@
 
-\subsection{A Lie group for numerical verification}
+\subsection{Decorators for Lie groups}
+
+The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing funtionality.
+For Lie groups there are currently two such decorators available
+
+\subsubsection*{Specifying a different inner product}
+
+By implementing \verb|inner(|$\frak g$\verb|, X, Y)| and \verb|norm(|$\frak g$\verb|, X)| for a Lie algebra $\frak g$\verb| = LieAlgebra(G)|,
+the Lie group \verb|G| is equipped with a certain inner product on the Lie algebra and hence on all tangent spaces to the Lie group.
+This is an implicit choice of a default metric on the Lie group, that is considered the default thereafter.
+
+If more than one metric is used in applications, a second metric can be introduced
+by wrapping the Lie group into a \verb|MetricLieGroup|, i.e.\ \verb|G2 = MetricLieGroup(G, m::M)|, where \verb|M| has to be a subtype of
+\verb|AbstractMetric| available from \verb|Manifolds.jl|\footnote{where the analogously a manifold of type~\texttt{MetricManifold} is defined.}.
+All functions unrelated to the metric are forwarded from \verb|G2| to \verb|G|, while all functions
+related to the inner product have to be specified for \verb|G2| anew.
+
+\subsubsection*{Adding numerical validation to a Lie group}
 
@@
 For all three representations, one can access the component semantically by writing \verb|g[G, :Rotation]| and \verb|g[G, :Translation]|, which works independent of the representation used.
-
-{\color{red}TODO: Check whether there are details for other Lie groups that should be mentioned here.}
+This allows to implement algorithgms wotking on \(\mathrm{SE}(n)\) indepentently of the chosen representation.
 
@@
 We first can generalize the de-Casteljau algorithm for Beziér curves to Lie groups by replacing the linear interpolation step with using the exponential and logarithmic map.
+
+In Code~\ref{lst:deCasteljauLieGroup}, we define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
 
@@
 
-In Code~\ref{lst:deCasteljauLieGroup}, we define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
-
 We can now compare the algorithm on the Lie group \(\mathrm{SE}(2)\) with the corresponding Bézier curve on the (direct) product Lie group $\mathrm{SO}(2) \times \mathcal T(2)$, where \(\mathcal T(2) = (\mathbb{R}^2, +)\) is the \verb|TranslationGroup(2)|.
@@
 \end{align*}
-
-We evaluate the corresponding Bézier curves on both Lie groups using the Code~\ref{lst:bezierLieGroupExample}.
-\begin{lstlisting}[
-    language = Julia,
-    numbers=left,
-    label={lst:bezierLieGroupExample},
-    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).},
-    mathescape
-]
-SO2 = SpecialOrthogonalGroup(2)
-R2 = TranslationGroup(2)
-SO2xR2 = SO2 $\times$ R2
-ts = range(0, 1, length=35)
-q1 = [
-    deCasteljauLieGroup(SO2xT2, cp, t) for t in ts
-]
-SE2 = SpecialEuclideanGroup(2)
-q2 = [deCasteljauLieGroup(SE2, cp, t) for t in ts]
-\end{lstlisting}
 
@@
 \end{figure}
+
+We evaluate the corresponding Bézier curves on both Lie groups using the Code~\ref{lst:bezierLieGroupExample}.
+\begin{lstlisting}[
+    language = Julia,
+    numbers=left,
+    label={lst:bezierLieGroupExample},
+    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).},
+    mathescape
+]
+SO2 = SpecialOrthogonalGroup(2)
+R2 = TranslationGroup(2)
+SO2xR2 = SO2 $\times$ R2
+ts = range(0, 1, length=35)
+q1 = [
+    deCasteljauLieGroup(SO2xT2, cp, t) for t in ts
+]
+SE2 = SpecialEuclideanGroup(2)
+q2 = [deCasteljauLieGroup(SE2, cp, t) for t in ts]
+\end{lstlisting}
 

### 2025-10-31 src/groups/metric_group.jl
Update src/groups/metric_group.jl
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@ -36,7 +36,7 @@ end
 Base.getindex(g, G::MetricLieGroup, i) = getindex(g, base_lie_group(G), i)
 Base.getindex(g::AbstractArray, G::MetricLieGroup, i) = getindex(g, base_lie_group(G), i)
 function Base.getindex(
-        X, 𝔤::LieAlgebra{𝔽, <:O, <:MetricLieGroup}, i
+        X, 𝔤::LieAlgebra{𝔽, O, <:MetricLieGroup}, i
     ) where {𝔽, O <: AbstractGroupOperation}
     # unwrap Algebra and metric decorator
     return getindex(X, LieAlgebra(base_lie_group(base_lie_group(𝔤))), i)

### 2025-10-31 src/groups/metric_group.jl
Update src/groups/metric_group.jl
Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@ -42,7 +42,7 @@ function Base.getindex(
     return getindex(X, LieAlgebra(base_lie_group(base_lie_group(𝔤))), i)
 end
 function Base.getindex(
-        X::AbstractArray, 𝔤::LieAlgebra{𝔽, <:O, <:MetricLieGroup}, i
+        X::AbstractArray, 𝔤::LieAlgebra{𝔽, O, <:MetricLieGroup}, i
     ) where {𝔽, O <: AbstractGroupOperation}
     # unwrap Algebra and metric decorator
     return getindex(X, LieAlgebra(base_lie_group(base_lie_group(𝔤))), i)

### 2025-10-31 src/LieGroups.jl
remnarks from code review.

@@ -31,7 +31,7 @@ using Manifolds:
     UnitaryMatrices
 using Manifolds: DeterminantOneMatrixType
 using Manifolds: base_manifold
-using Manifolds: DefaultMetric
+using Manifolds: DefaultMetric, metric
 import LinearAlgebra: adjoint, adjoint!
 using ManifoldsBase:
     AbstractBasis, AbstractMetric, AbstractNumbers,
@@ -177,7 +177,7 @@ export injectivity_radius
 export jacobian_conjugate, jacobian_conjugate!
 export jacobian_exp, jacobian_exp!
 export lie_bracket, lie_bracket!, log, log!
-export manifold_dimension
+export manifold_dimension, metric
 export norm, number_of_coordinates
 export push_forward_tangent, push_forward_tangent!, pull_back_tangent, pull_back_tangent!
 export rand, rand!

### 2025-10-31 src/groups/metric_group.jl

@@ -1,9 +1,12 @@
 """
-    MetricLieGroup{𝔽,M<:AbstractLieGroup{𝔽},G<:AbstractMetric} <: AbstractDecoratorManifold{𝔽}
+    MetricLieGroup{
+        𝔽, O <: AbstractGroupOperation, M <: ManifoldsBase.AbstractManifold{𝔽},
+        L <: LieGroup{𝔽, O, M}, G <: AbstractMetric
+    } <: AbstractLieGroup{𝔽, O, M}}
 
-Equip an [`AbstractLieGroup`](@ref) `L` explicitly with an [`AbstractMetric`](@extref `ManifoldsBase.AbstractMetric`) `G`.
+Equip an [`AbstractLieGroup`](@ref) ``$(_math(:G))`` explicitly with an [`AbstractMetric`](@extref `ManifoldsBase.AbstractMetric`) `⟨⋅,⋅⟩`.
 
-By default every [`AbstractLieGroup`](@ref) `L` is assumed to implicitly
+By default every [`AbstractLieGroup`](@ref) ``$(_math(:G))`` is assumed to implicitly
 implement all functions on their [`LieAlgebra`](@ref) with respect to a metric,
 that corresponds to a certain “default metric”, either because it is a widely recognized metric
 or because the first implementation within `LieGroups.jl` was done with respect to this metric.
@@ -18,20 +21,27 @@ or functions that are already otherwise passed on to the inner manifold like [`m
 
 # Constructor
 
-    MetricManifold(M, G)
+    MetricLieGroup(G::AbstractLieGroup, metric::AbstractMetric)
 
-Generate the [`AbstractLieGroup`](https://juliamanifolds.github.io/Manifolds.jl/latest/interface.html#ManifoldsBase.AbstractLieGroup) `M` as a manifold with the `AbstractMetric` `G`.
+Generate the [`AbstractLieGroup`](@ref) equipped explicitly with the [`AbstractMetric`](@extref `ManifoldsBase.AbstractMetric`) `metric`.
 """
 struct MetricLieGroup{
         𝔽,
         O <: AbstractGroupOperation,
         M <: ManifoldsBase.AbstractManifold{𝔽},
-        L <: LieGroup{𝔽, O, M},
-        G <: AbstractMetric,
+        G <: LieGroup{𝔽, O, M},
+        Metric <: AbstractMetric,
     } <: AbstractLieGroup{𝔽, O, M}
[…]

### 2025-10-31 src/group_operations/addition_operation.jl
coverage improvement stuff

@@ -101,7 +101,7 @@ Compute the differential of the group operation ``g$(_math(:∘))h``, on an [`Ab
 with respect to its second (right) argument `h`.
 
 Another interpretation is to consider a function where we do a fixed multiplication from the left with `g`.
-i..e. the left group multiplication function ``λ_g(h) = g$(_math(:∘))h`` (where the _left_ refers to the fixed argument ``g``.).
+i.e. the left group multiplication function ``λ_g(h) = g$(_math(:∘))h`` (where the _left_ refers to the fixed argument ``g``.).
 
 For the [`AdditionGroupOperation`](@ref) it reads ``$(_math(:d))λ_g(h)[X] = X``.
 """

### 2025-10-31 src/groups/power_group.jl

@@ -164,6 +164,23 @@ function diff_left_compose!(
     end
     return Y
 end
+function diff_left_compose!(
+        PoG::LieGroup{𝔽, Op, M}, Y, ::Identity{Op}, h, X
+    ) where {𝔽, Op <: PowerGroupOperation, M <: ManifoldsBase.AbstractPowerManifold}
+    PM = PoG.manifold
+    rep_size = representation_size(PM)
+    G = LieGroup(PM.manifold, PoG.op.op)
+    for i in ManifoldsBase.get_iterator(PM)
+        diff_left_compose!(
+            G,
+            ManifoldsBase._write(PM, rep_size, Y, i),
+            PoG.op.op,
+            ManifoldsBase._read(PM, rep_size, h, i),
+            ManifoldsBase._read(PM, rep_size, X, i),
+        )
+    end
+    return Y
+end
 
 function diff_right_compose!(
         PoG::LieGroup{𝔽, Op, M}, Y, g, h, X
@@ -176,6 +193,23 @@ function diff_right_compose!(
             G,
             ManifoldsBase._write(PM, rep_size, Y, i),
             ManifoldsBase._read(PM, rep_size, g, i),
+            ManifoldsBase._read(PM, rep_size, h, i),
+            ManifoldsBase._read(PM, rep_size, X, i),
+        )
+    end
+    return Y
+end
+function diff_right_compose!(
+        PoG::LieGroup{𝔽, Op, M}, Y, ::Identity{Op}, h, X
+    ) where {𝔽, Op <: PowerGroupOperation, M <: ManifoldsBase.AbstractPowerManifold}
+    PM = PoG.manifold
+    rep_size = representation_size(PM)
+    G = LieGroup(PM.manifold, PoG.op.op)
+    for i in ManifoldsBase.get_iterator(PM)
+        diff_right_compose!(
+            G,
+            ManifoldsBase._write(PM, rep_size, Y, i),
+            PoG.op.op,
             ManifoldsBase._read(PM, rep_size, h, i),
             ManifoldsBase._read(PM, rep_size, X, i),
         )

### 2025-10-31 test/groups/test_special_euclidean_group.jl

@@ -84,6 +84,12 @@ using StaticArrays
             @test (vec[1])[𝔤, :] isa Tuple
             @test (vec[1])[𝔤, :Rotation] isa AbstractMatrix
         end
+
+        @testset "init_constants" begin
+            g = SpecialEuclideanMatrixPoint(NaN .* g1)
+            LieGroups.init_constants!(G2f, g)
+            @test g.value[3, :] == [0, 0, 1]
+        end
         #
         # Right variant – exchange product cases
         G2r = TranslationGroup(2) ⋊ SpecialOrthogonalGroup(2)

### 2025-10-31 src/LieGroups.jl
Fix lie bracket pass through.

@@ -10,7 +10,7 @@ The package is named after the Norwegian mathematician [Marius Sophus Lie](https
 """
 module LieGroups
 
-using LinearAlgebra, ManifoldsBase, Quaternions, StaticArrays, Random
+using LinearAlgebra, Manifolds, ManifoldsBase, Quaternions, StaticArrays, Random
 
 #
 #
@@ -101,6 +101,7 @@ export LeftSemidirectProductLieGroup, RightSemidirectProductLieGroup
 export ValidationLieGroup
 export DefaultLieAlgebraOrthogonalBasis
 export ×, ^, ⋉, ⋊
+export DefaultMetric
 #
 #
 # Group Operations

### 2025-10-31 src/groups/metric_group.jl

@@ -98,12 +98,24 @@ function identity_element!(G::MetricLieGroup, e)
     return identity_element!(base_lie_group(G), e)
 end
 
-
 _inv(G::MetricLieGroup, g) = inv(base_lie_group(G), g)
 _inv!(G::MetricLieGroup, k, g) = inv!(base_lie_group(G), k, g)
 
 ManifoldsBase.is_point(G::MetricLieGroup, g; kwargs...) = is_point(base_lie_group(G), g; kwargs...)
 ManifoldsBase.is_point(G::MetricLieGroup, e::Identity; kwargs...) = is_point(base_lie_group(G), e; kwargs...)
+
+function lie_bracket(
+        𝔤::LieAlgebra{𝔽, O, <:MetricLieGroup}, X, Y
+    ) where {𝔽, O <: AbstractGroupOperation}
+    G = base_lie_group(base_lie_group(𝔤)) # access Lie group without metric decorator
+    return lie_bracket(LieAlgebra(G), X, Y)
+end
+function lie_bracket!(
+        𝔤::LieAlgebra{𝔽, O, <:MetricLieGroup}, Z, X, Y
+    ) where {𝔽, O <: AbstractGroupOperation}
+    G = base_lie_group(base_lie_group(𝔤)) # access Lie group without metric decorator
+    return lie_bracket!(LieAlgebra(G), Z, X, Y)
+end
 
 function Base.show(io::IO, G::MetricLieGroup)
     return print(io, "MetricLieGroup($(G.lie_group), $(G.metric))")

### 2025-10-31 test/groups/test_metric_group.jl

@@ -10,7 +10,9 @@ using LieGroupsTestSuite
 
     @testset "Pass through with a dummy metric" begin
         SO2 = SpecialOrthogonalGroup(2)
+        @test metric(SO2) === DefaultMetric()
         G = MetricLieGroup(SO2, LieGroupsTestSuite.DummyMetric())
+        @test metric(G) === LieGroupsTestSuite.DummyMetric()
         @test base_lie_group(G) === SO2
 
         g1 = 1 / sqrt(2) * [1.0 1.0; -1.0 1.0]

### 2025-10-31 test/groups/test_metric_group.jl
the test I was thinking about

@@ -1,10 +1,23 @@
 using LieGroups, ManifoldsBase, Random, Test, RecursiveArrayTools
 
 using LieGroups: MetricLieGroup
+using Manifolds
 
 s = joinpath(@__DIR__, "..", "LieGroupsTestSuite.jl")
 !(s in LOAD_PATH) && (push!(LOAD_PATH, s))
 using LieGroupsTestSuite
+
+struct CustomTranslationMetric <: ManifoldsBase.AbstractMetric end
+
+function ManifoldsBase.exp!(
+        ::MetricLieGroup{ℝ, AdditionGroupOperation, <:Euclidean, <:TranslationGroup, CustomTranslationMetric},
+        q,
+        p,
+        X,
+    )
+    return q .= p .+ 2 .* X
+end
+
 
 @testset "MetricLieGroup: A metric decorator for LieGroups" begin
 
@@ -62,5 +75,12 @@ using LieGroupsTestSuite
         @test gT[SE2, :Translation] == gT[G, :Translation]
         @test XT[se2, :Rotation] == XT[𝔤, :Rotation]
         @test XT[se2, :Translation] == XT[𝔤, :Translation]
+    end
+    @testset "Testing that correct defaults for metric functions are used" begin
+        T2 = TranslationGroup(2)
+        G = MetricLieGroup(T2, CustomTranslationMetric())
+        p = [0.0, 2.0]
+        X = [1.0, -2.0]
+        @test exp(G, p, X) == [2.0, -2.0]
     end
 end

### 2025-10-31 paper/paper.tex
Make code blocks floating.

@@
 
-\begin{figure}
+\begin{figure}[tbp]
     \centering
@@
 
-We first can generalize the de-Casteljau algorithm for Beziér curves to Lie groups by replacing the linear interpolation step with using the exponential and logarithmic map.
-
-In Code~\ref{lst:deCasteljauLieGroup}, we define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
-
 \begin{lstlisting}[
+    float=tbp,
     language = Julia,
@@
 
+We first can generalize the de-Casteljau algorithm for Beziér curves to Lie groups by replacing the linear interpolation step with using the exponential and logarithmic map.
+
+In Code~\ref{lst:deCasteljauLieGroup}, we define the general de-Casteljau algorithm on a Lie group \verb|G|, given a vector of control points \verb|cp| and a parameter \verb|t| in \([0,1]\).
+
 We can now compare the algorithm on the Lie group \(\mathrm{SE}(2)\) with the corresponding Bézier curve on the (direct) product Lie group $\mathrm{SO}(2) \times \mathcal T(2)$, where \(\mathcal T(2) = (\mathbb{R}^2, +)\) is the \verb|TranslationGroup(2)|.
@@
 
-\begin{figure}
+We evaluate the corresponding Bézier curves on both Lie groups using the Code~\ref{lst:bezierLieGroupExample}.
+\begin{lstlisting}[
+    float=tbp,
+    language = Julia,
+    numbers=left,
+    label={lst:bezierLieGroupExample},
+    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).},
+    mathescape
+]
+SO2 = SpecialOrthogonalGroup(2)
+R2 = TranslationGroup(2)
+SO2xR2 = SO2 $\times$ R2
+ts = range(0, 1, length=35)
+q1 = [
+    deCasteljauLieGroup(SO2xT2, cp, t) for t in ts
+]
+SE2 = SpecialEuclideanGroup(2)
+q2 = [deCasteljauLieGroup(SE2, cp, t) for t in ts]
+\end{lstlisting}
+
+\begin{figure}[tbp]
     % Arrow scale factor
@@
 
-We evaluate the corresponding Bézier curves on both Lie groups using the Code~\ref{lst:bezierLieGroupExample}.
-\begin{lstlisting}[
-    language = Julia,
-    numbers=left,
-    label={lst:bezierLieGroupExample},
-    caption={Evaluating Bézier curves on \(\mathrm{SE}(2)\) and \(\mathrm{SO}(2) \times \mathcal T(2)\).},
-    mathescape
-]
-SO2 = SpecialOrthogonalGroup(2)
-R2 = TranslationGroup(2)
-SO2xR2 = SO2 $\times$ R2
-ts = range(0, 1, length=35)
-q1 = [
-    deCasteljauLieGroup(SO2xT2, cp, t) for t in ts
-]
-SE2 = SpecialEuclideanGroup(2)
-q2 = [deCasteljauLieGroup(SE2, cp, t) for t in ts]
-\end{lstlisting}
-
 The resulting Bézier curves on both Lie groups are shown in Figure~\ref{fig:Bezier}.
@@
 \begin{lstlisting}[
+    float=tbp,
     language = Julia,
@@
 
-\begin{figure}
+\begin{figure}[tbp]
     % Arrow scale factor

### 2025-10-31 src/groups/special_euclidean_group.jl
Lots of small fixes and diff_left_compose on SE(n) (#81)
* Lots of small fixes and diff_left_compose on SE(n)

* more methods and stuff

* disambiguate stuff

* coverage improvement stuff

@@ -315,6 +315,64 @@ function default_right_action(::TranslationGroup, ::SpecialOrthogonalGroup)
     return LeftMultiplicationGroupAction()
 end
 
+@doc raw"""
+    diff_left_compose(G::SpecialEuclideanGroup, g, h, X)
+
+Compute the differential of left composition by `h` on [`SpecialEuclideanGroup`](@ref)
+for tangent vector `X` at `g`.
+
+Let
+```math
+h=\begin{pmatrix} R & t\\[4pt] 0 & 1\end{pmatrix},\qquad
+X=\begin{pmatrix} X_{\mathrm{R}} & X_{\mathrm{t}}\\[4pt] 0 & 0\end{pmatrix},
+```
+
+where ``R\in SO(n)``, ``t\in\mathbb{R}^n``, ``X_{\mathrm{R}}`` is the skew-symmetric rotation block and ``X_{\mathrm{t}}`` the translation column.
+Then the differential is the adjoint action by ``h^{-1}``:
+```math
+\mathrm{D}λ_h(X)=h^{-1}Xh=
+\begin{pmatrix}
+R^\top X_{\mathrm{R}} R & R^\top\bigl(X_{\mathrm{R}}\,t + X_{\mathrm{t}}\bigr)\\[4pt]
+0 & 0
+\end{pmatrix}.
+```
+
+Component-wise:
+```math
+Y_{\mathrm{R}} = R^\top X_{\mathrm{R}} R,\qquad Y_{\mathrm{t}} = R^\top\bigl(X_{\mathrm{R}}\,t + X_{\mathrm{t}}\bigr).
+```
+"""
+diff_left_compose(G::SpecialEuclideanGroup, g, h, X)
+
+function diff_left_compose!(G::SpecialEuclideanGroup, Y, g, h, X)
+    GA = LieAlgebra(G)
+    init_constants!(GA, Y)
+    XR = submanifold_component(GA, X, Val(:Rotation))
+    Xt = submanifold_component(GA, X, Val(:Translation))
+    YR = submanifold_component(GA, Y, Val(:Rotation))
+    Yt = submanifold_component(GA, Y, Val(:Translation))
+    R = submanifold_component(G, h, Val(:Rotation))
+    t = submanifold_component(G, h, Val(:Translation))
+    A = R' * XR
+    mul!(YR, A, R)
+    Yt .= A * t .+ R' * Xt
+    return Y
+end
+
+
+@doc raw"""
[…]

### 2025-10-31 src/interface.jl

@@ -417,6 +417,10 @@ function diff_left_compose(G::AbstractLieGroup, g, h, X)
     Y = ManifoldsBase.allocate_result(G, diff_left_compose, X, g, h)
     return diff_left_compose!(G, Y, g, h, X)
 end
+function diff_left_compose(G::AbstractLieGroup{𝔽, O}, g::Identity{O}, h, X) where {𝔽, O <: AbstractGroupOperation}
+    Y = ManifoldsBase.allocate_result(G, diff_left_compose, X, h)
+    return diff_left_compose!(G, Y, g, h, X)
+end
 
 function diff_left_compose! end
 @doc "$(_doc_diff_left_compose)"
@@ -453,6 +457,11 @@ $(_math(:d)) λ_g(h)[X] = X ∈ 𝔤.
 @doc "$(_doc_diff_right_compose)"
 function diff_right_compose(G::AbstractLieGroup, g, h, X)
     Y = ManifoldsBase.allocate_result(G, diff_right_compose, X, g, h)
+    return diff_right_compose!(G, Y, g, h, X)
+end
+
+function diff_right_compose(G::AbstractLieGroup{𝔽, O}, g::Identity{O}, h, X) where {𝔽, O <: AbstractGroupOperation}
+    Y = ManifoldsBase.allocate_result(G, diff_right_compose, X, h)
     return diff_right_compose!(G, Y, g, h, X)
 end
 
@@ -1237,6 +1246,10 @@ function _vector_transport_to_basemanifold(
     # now Y is in ThM so we still ave to pull it back to TeM using
     # the left group opp diff.
     return pull_back_tangent(G, g, X)
+end
+
+function ManifoldsBase.zero_vector(G::AbstractLieGroup{𝔽, O}, ::Identity{O}) where {𝔽, O <: AbstractGroupOperation}
+    return zero_vector(LieAlgebra(G))
 end
 
 #

### 2025-10-31 src/groups/metric_group.jl
fix ambiguities

@@ -83,9 +83,15 @@ diff_inv(G::MetricLieGroup, g, X) = diff_inv(base_lie_group(G), g, X)
 diff_inv!(G::MetricLieGroup, Y, g, X) = diff_inv!(base_lie_group(G), Y, g, X)
 
 diff_left_compose(G::MetricLieGroup, g, h, X) = diff_left_compose(base_lie_group(G), g, h, X)
+function diff_left_compose(G::MetricLieGroup{𝔽, O}, g::Identity{O}, h, X) where {𝔽, O <: AbstractGroupOperation}
+    return diff_left_compose(base_lie_group(G), g, h, X)
+end
 diff_left_compose!(G::MetricLieGroup, Y, g, h, X) = diff_left_compose!(base_lie_group(G), Y, g, h, X)
 
 diff_right_compose(G::MetricLieGroup, g, h, X) = diff_right_compose(base_lie_group(G), g, h, X)
+function diff_right_compose(G::MetricLieGroup{𝔽, O}, g::Identity{O}, h, X) where {𝔽, O <: AbstractGroupOperation}
+    return diff_right_compose(base_lie_group(G), g, h, X)
+end
 diff_right_compose!(G::MetricLieGroup, Y, g, h, X) = diff_right_compose!(base_lie_group(G), Y, g, h, X)
 
 function identity_element(G::MetricLieGroup)

### 2025-10-31 test/groups/test_metric_group.jl

@@ -18,7 +18,6 @@ function ManifoldsBase.exp!(
     return q .= p .+ 2 .* X
 end
 
-
 @testset "MetricLieGroup: A metric decorator for LieGroups" begin
 
     @testset "Pass through with a dummy metric" begin

### 2025-10-31 src/LieGroups.jl
merge inv! functions

@@ -32,6 +32,7 @@ using Manifolds:
 using Manifolds: DeterminantOneMatrixType
 using Manifolds: base_manifold
 using Manifolds: DefaultMetric, metric
+import Manifolds: inv!
 import LinearAlgebra: adjoint, adjoint!
 using ManifoldsBase:
     AbstractBasis, AbstractMetric, AbstractNumbers,

### 2025-10-31 paper/paper.tex
A few small adaptions in TeX

@@
 \newcommand{\im}{\ensuremath{\mathrm{i}}}
-\newcommand{\LieGroupsVersion}{0.1.7}
+\newcommand{\LieGroupsVersion}{0.1.8}
 %
@@
 For all three representations, one can access the component semantically by writing \verb|g[G, :Rotation]| and \verb|g[G, :Translation]|, which works independent of the representation used.
-This allows to implement algorithgms wotking on \(\mathrm{SE}(n)\) indepentently of the chosen representation.
 
@@
 \subsection{A Beziér curve on a Lie group}
-
-In the first example we illustrate how to construct and plot a Beziér curve on a Lie group, here the special Euclidean group \(\mathrm{SE}(2)\).
 
@@
 \end{lstlisting}
+
+In the first example we illustrate how to construct and plot a Beziér curve on a Lie group, here the special Euclidean group \(\mathrm{SE}(2)\).
 
@@
 Position and orientation of the rod is indicated at 20 equally spaced times, along with a densely sampled trajectory of the center of the rod.
-
+\\
 For more advanced methods of solving differential equations on Manifolds and Lie groups in Julia, see
@@
 
- \balance
+%\balance
+\newpage
 \input{bib.tex}

### 2025-10-31 src/interface.jl
Introduce a MetricLieGroup (#74)
* Initial setup.
* Fix documentation.
* Start testing on metric lie
* testing all passthrough functions.
* Passthrough for getindex.
* fix element access.
* Improve test coverage
* remnarks from code review.
* Fix lie bracket pass through.
* the test I was thinking about
* Fix docs.
* fix ambiguities
* merge inv! functions
* Bump date.

---------

Co-authored-by: Mateusz Baran <mateuszbaran89@gmail.com>

@@ -629,7 +629,7 @@ end
 
 function inv! end
 @doc "$_doc_inv"
-inv!(G::AbstractLieGroup, h, g) = _inv!(G, h, g)
+Manifolds.inv!(G::AbstractLieGroup, h, g) = _inv!(G, h, g)
 
 function Base.inv(
         ::AbstractLieGroup{𝔽, O}, e::Identity{O}
@@ -637,7 +637,7 @@ function Base.inv(
     return e
 end
 
-function inv!(
+function Manifolds.inv!(
         G::AbstractLieGroup{𝔽, O}, g, ::Identity{O}
     ) where {𝔽, O <: AbstractGroupOperation}
     return identity_element!(G, g)

### 2025-10-31 paper/rod_example.jl
Unify scripts to both use the paper/ environment.

@@ -1,3 +1,5 @@
+using Pkg; Pkg.activate(@__DIR__); Pkg.instantiate()
+cd(@__DIR__)
 using LinearAlgebra
 using LieGroups, Manifolds, RecursiveArrayTools
 using CairoMakie
@@ -118,7 +120,7 @@ function sim()
            SystemParameters(; initial_rod_angle = 0.2, initial_rod_position=[-0.2, +1.0], tmax=30.0, initial_rod_velocity = [-0.12, -0.15]),
            SystemParameters(; initial_rod_angle = 0.6, initial_rod_position=[-0.5, -0.5], tmax=30.0, initial_rod_velocity = [-0.04, 0.12])]
 
-    export_folder = "paper/data/"
+    export_folder = "data/"
     colors = [:red, :blue, :green]
 
     for (isp, sp) in enumerate(sps)
@@ -159,7 +161,7 @@ function sim()
                  color = [:red, :blue], markersize = 12)
 
     display(fig)
-    save("rod_trajectories_makie.png", fig)
+    save("img/rod_trajectories_makie.png", fig)
 end
 
 sim()

### 2025-11-01 paper/header.tex
second example
Fix metadata and giving the paper a careful read.

--- 2025-10-25 paper/header.tex
@@
 \author[1]{Ronny Bergmann}
+\author[2]{Mateusz Baran}
 \affil[1]{Department of Mathematical Sciences, Norwegian University of Science and Technology, Trondheim, Norway}
+\affil[2]{AGH University of Science and Technology, Kraków, Poland}
 
-\keywords{Julia, Riemannian manifolds, Lie groups, Differential geometry}
+\keywords{Riemannian manifolds, Lie groups, differential geometry, numerical analysis, scientific computing}
 
@@
 pdftitle = {Groups and smooth geometry using LieGroups.jl},
-pdfsubject = {JuliaCon 2022 Proceedings},
-pdfauthor = {Ronny Bergmann},
-pdfkeywords = {Julia, Riemannian manifolds, Lie groups, Differential geometry},
+pdfsubject = {JuliaCon 2025 Proceedings},
+pdfauthor = {Ronny Bergmann, Mateusz Baran},
+pdfkeywords = {Riemannian manifolds, Lie groups, differential geometry, numerical analysis, scientific computing},
 }

### 2025-11-01 paper/paper.tex
Fix a typo.
Fix metadata and giving the paper a careful read.

@@
 %
-% Define muted colors.
+% Define muted colours.
 \definecolor{TolMutedBlue}{HTML}{332288}
@@
 but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
-and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998}.
+and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998,IserlesMunthe-KaasNoersettZanna:2000}.
 
@@
 
-This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}.
-After introducing the necessary mathematical background and notation in Section~\ref{sec:Notation},
-we present the interface in Section~\ref{sec:Interface}, split into the Lie group, the Lie algebra, and group actions.
-In Section~\ref{sec:LieGroups}, we present an overview of currently implemented Lie groups.
+This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}, whose logo is shown on the right.
+%
+After introducing some mathematical background in Section~\ref{sec:Notation},
+we present the interface in Section~\ref{sec:Interface} for the Lie group, the Lie algebra, and group actions.
+Section~\ref{sec:LieGroups} provides an overview of all currently implemented Lie groups.
 Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups.jl|.
@@
 \\
-As a group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
+A group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
 Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
-As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
+As an example, consider the special orthogonal group \(\mathrm{SO}(n)\), consisting of all \(n \times n\) orthogonal matrices, with determinant \(1\), that is, for \(p\in \mathrm{SO}(n)\), we have \(p^{\mathrm{T}} p = I\) and \(\det(p) = 1\) together with the group operation \(\cdot\) given by matrix multiplication.
 For $n=2$ these are rotations in the plane, hence each operation can be identified with an angle $\alpha \in [-\pi, \pi)$\footnote{Note that the identification is not continuous.}, or, continuously, with the circle.
 % chktex 9ok
-The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose (or an angle $-\alpha$).
+The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose $p^{\mathrm{T}}$ (or an angle $-\alpha$).
 % -(-\pi) would be out of the range
@@
 The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
-Then, using the differential (or push forward) $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$\footnote{\color{red} MB: TODO: does the reader know that a differential is?
-(RB) Sure I can expand on that, I wrote a sentence more further below to get from differential to Jacobian, so I can also add a sentence there as well.}, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
+For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
+Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
 \footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields.
@@
 Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
-In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \frak{g}$.
+In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \mathfrak{g}$.
 
@@
 
-Concerning a metric on the tangent spaces, manifolds are turned into Riemannian manifolds when they are equipped with a Riemannian metric \(\langle \cdot, \cdot \rangle_p\colon T_p\mathcal{M} \times T_p\mathcal{M} \to \mathbb{R}\) for each point \(p \in \mathcal{M}\) that smoothly varies with \(p\).
+Concerning a metric on the tangent spaces, smooth manifolds are turned into Riemannian manifolds when they are equipped with a Riemannian metric \(\langle \cdot, \cdot \rangle_p\colon T_p\mathcal{M} \times T_p\mathcal{M} \to \mathbb{R}\) for each point \(p \in \mathcal{M}\) that smoothly varies with \(p\).
 For Lie groups, we can use a single inner product \(\langle \cdot, \cdot \rangle\) on the Lie algebra \(\mathfrak{g}\) and use the change in representation as mentioned above to define
@@
 
-\subsection{Group Actions} A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).\footnote{This is the convention for left actions.
+\subsection{Group Actions}\label{subsec:GroupActions}
+A \term{group action} of a Lie group \(\mathcal{G}\) on a smooth manifold \(\mathcal{M}\) is a smooth map \(\sigma\colon \mathcal{G} \times \mathcal{M} \to \mathcal{M}\) such that for all \(g, h \in \mathcal{G}\) and \(p \in \mathcal{M}\) it holds that \(\sigma(e, p) = p\) and \(\sigma(g, \sigma(h, p)) = \sigma(g \cdot h, p)\).\footnote{This is the convention for left actions.
 Alternatively, right actions fulfil \(\sigma(g, \sigma(h, p)) = \sigma(h \cdot g, p)\).}
@@
 
-The same action can also be applied to points from the sphere \(\mathbb{S}^2\), resulting in a similar group action \(\sigma\colon \mathbb{SO}(3) \times \mathbb{S}^2 \to \mathbb{S}^2\), \(\sigma(R, p) \coloneqq Rp\).
+The same action can also be applied to points from the sphere \(\mathbb{S}^2\), resulting in a similar group action \(\sigma\colon \mathrm{SO}(3) \times \mathbb{S}^2 \to \mathbb{S}^2\), \(\sigma(R, p) \coloneqq Rp\).
 
@@
 For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.
+To generate a point in a specific representation of type \verb|T|, one can use \verb|identity_element(G, T)|.
 
@@
 By default the allocating variant is implemented by allocating memory accordingly and calling the in-place variant.
-This default can be overriden, for example for performance reasons.
+This default can be overridden, for example for performance reasons.
 
@@
 For such a function, for example again a function \verb|f(G)| a positional argument \verb|T| can be used to specify the desired return type, i.e.\ \verb|f(G, T)| returns a point or tangent vector of type \verb|T|.
-The in-place variant does not require this, since in its signatgure of the form \verb|f!(G, g::T)| the type \verb|T| is known from the argument \verb|g|.
+The in-place variant does not require this, since in its signature of the form \verb|f!(G, g::T)| the type \verb|T| is known from the argument \verb|g|.
 
@@
 .
-If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\frak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
+If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\mathfrak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
 In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
@@
 
-Similar to points on the Lie group, when representing elements $X \in \frak g$ from a Lie algebra, we do not type the general functions of the interface.
+Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
 This allows to use either plain arrays or own structures to represent these in code.
@@
 
-A central function on the Lie algebra is the Lie bracket $[\cdot, \cdot]\colon \frak g \times \frak g \to \frak g$, which is available as \verb|lie_bracket(|$\frak g$\verb|, X, Y)|.
-% \footnote{in-place variant \texttt{lie\_bracket!($\frak g$, Z, X, Y)}.}
+A central function on the Lie algebra is the Lie bracket $[\cdot, \cdot]\colon \mathfrak g \times \mathfrak g \to \mathfrak g$, which is available as \verb|lie_bracket(|$\mathfrak g$\verb|, X, Y)|.
+% \footnote{in-place variant \texttt{lie\_bracket!($\mathfrak g$, Z, X, Y)}.}
 For more details on the Lie bracket see~\cite[Ch.~5]{HilgertNeeb:2012}.
@@
 \subsubsection*{Topological basics.}
-Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\frak g$ \verb|= LieAlgebra(G)|.
-To access the Lie group again, use \verb|base_lie_group(|$\frak g$\verb|)|. Similarly \verb|base_manifold(|$\frak g$\verb|)| returns the underlying manifold of the Lie group.\\
+Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\mathfrak g$ \verb|= LieAlgebra(G)|.
+To access the Lie group again, use \verb|base_lie_group(|$\mathfrak g$\verb|)|. Similarly \verb|base_manifold(|$\mathfrak g$\verb|)| returns the underlying manifold of the Lie group.\\
 As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
-A zero vector is generated via \verb|zero_vector(|$\frak g$\verb|)| for the default representation and \verb|zero_vector(|$\frak g$\verb|, T)| for a specific representation type \verb|T|%
-%\footnote{the in-place variant \texttt{zero\_vector!($\frak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
+A zero vector is generated via \verb|zero_vector(|$\mathfrak g$\verb|)| for the default representation and \verb|zero_vector(|$\mathfrak g$\verb|, T)| for a specific representation type \verb|T|%
+%\footnote{the in-place variant \texttt{zero\_vector!($\mathfrak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
 .
 
-The main topological function is to test the validity of a tangent vector \verb|is_vector(|$\frak g$\verb|, X; error=:none)| using the same \verb|error=| keyword as \verb|is_point| on the Lie group.
+The main topological function is to test the validity of a tangent vector \verb|is_vector(|$\mathfrak g$\verb|, X; error=:none)| using the same \verb|error=| keyword as \verb|is_point| on the Lie group.
 
 \subsubsection*{Vector space related functions}
-For the following functions related to vector space features, using the Lie algebra $\frak g$ as first argument is equivalent to specifying the Lie Group $G$ and an arbitrary point $g \in \mathcal G$.
+For the following functions related to vector space features, using the Lie algebra $\mathfrak g$ as first argument is equivalent to specifying the Lie Group $G$ and an arbitrary point $g \in \mathcal G$.
 This yields that the Lie group complies with the general interface for manifolds.
 
-The inner product and norm on the Lie algebra are available as \verb|inner(|$\frak g$\verb|, X, Y)| and \verb|norm(|$\frak g$\verb|, X)|, respectively.
+The inner product and norm on the Lie algebra are available as \verb|inner(|$\mathfrak g$\verb|, X, Y)| and \verb|norm(|$\mathfrak g$\verb|, X)|, respectively.
 Furthermore, there are two functions to convert between a coordinate-free representation of $X$ as a tangent vector and its representation in coordinates of a basis.
 
-Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\frak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
-When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\frak g$\verb|, c, b, T)| specifying the tangent vector type%
-%\footnote{the in-place variant \texttt{get\_vector!($\frak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
+Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
+When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b, T)| specifying the tangent vector type%
+%\footnote{the in-place variant \texttt{get\_vector!($\mathfrak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
 .
-Given a tangent vector $X \in \frak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\frak g$\verb|, X, B)|%
-%\footnote{in-place variant \texttt{get\_coordinates!($\frak g$, c, X, b)}}
+Given a tangent vector $X \in \mathfrak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\mathfrak g$\verb|, X, B)|%
+%\footnote{in-place variant \texttt{get\_coordinates!($\mathfrak g$, c, X, b)}}
 , where the Basis is again optional.
 
-For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\frak g$\verb|, c)|
-%\footnote{in-place variant \texttt{hat!($\frak g$, X::T, c)}}
-(again with an optional vector type \verb|T|) and \verb|vee(|$\frak g$\verb|, X)|
-%\footnote{in-place variant \texttt{vee!($\frak g$, c, X)}}
+For the case of the default, the \verb|DefaultLieAlgebraOrthogonalBasis()| the more commonly used names \verb|hat(|$\mathfrak g$\verb|, c)|
+%\footnote{in-place variant \texttt{hat!($\mathfrak g$, X::T, c)}}
+(again with an optional vector type \verb|T|) and \verb|vee(|$\mathfrak g$\verb|, X)|
+%\footnote{in-place variant \texttt{vee!($\mathfrak g$, c, X)}}
 are implemented
@@
 \subsubsection*{Push forward and pull back of tangent vectors}
-To the representation of a tangent vector $X \in \frak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$.
+To the representation of a tangent vector $X \in \mathfrak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$.
 This is implemented as \verb|push_forward_tangent(G, g, X)|%
@@
 .
-Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\frak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$.
+Conversely, to represent a tangent vector $Y \in T_g\mathcal G$ back in the Lie algebra $\mathfrak g$, we have to use the pull back of the left multiplication with $g^{-1}$, i.e.\ $D\lambda_{g^{-1}}(g)[Y]$.
 This is implemented as \verb|pull_back_tangent(G, g, Y)|%
@@
 \verb|jacobian_exp(G, p, X, b)|
-%\footnote{in-place variant \texttt{jacobian\_exp!($\frak g$, J, p, X, b)}}
+%\footnote{in-place variant \texttt{jacobian\_exp!($\mathfrak g$, J, p, X, b)}}
 and \verb|jacobian_conjugate(G, g, X, b)|%
-%\footnote{in-place variant \texttt{jacobian\_conjugate!($\frak g$, J, g, X, b)}}
+%\footnote{in-place variant \texttt{jacobian\_conjugate!($\mathfrak g$, J, g, X, b)}}
 , where \verb|b| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.
@@
 %\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
-to compute \(D\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \frak g\).
+to compute \(D\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \mathfrak g\).
 
@@
 Though mathematically the case of a product Lie Group, where $G=H$ is also (just) a product, computationally we can benefit from the special structure.
-Therefore the \verb|Powermanifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(G\) with itself.
+Therefore the \verb|PowerManifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(G\) with itself.
 As a shortcut, you can use \verb|G^n| to construct the \(n\)-fold Cartesian product of the Lie group \(G\) with itself.
@@
 
-The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing funtionality.
+The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing functionality.
 For Lie groups there are currently two such decorators available
@@
 
-By implementing \verb|inner(|$\frak g$\verb|, X, Y)| and \verb|norm(|$\frak g$\verb|, X)| for a Lie algebra $\frak g$\verb| = LieAlgebra(G)|,
+By implementing \verb|inner(|$\mathfrak g$\verb|, X, Y)| and \verb|norm(|$\mathfrak g$\verb|, X)| for a Lie algebra $\mathfrak g$\verb| = LieAlgebra(G)|,
 the Lie group \verb|G| is equipped with a certain inner product on the Lie algebra and hence on all tangent spaces to the Lie group.
@@
 by wrapping the Lie group into a \verb|MetricLieGroup|, i.e.\ \verb|G2 = MetricLieGroup(G, m::M)|, where \verb|M| has to be a subtype of
-\verb|AbstractMetric| available from \verb|Manifolds.jl|\footnote{where the analogously a manifold of type~\texttt{MetricManifold} is defined.}.
+\verb|AbstractMetric| available from \verb|Manifolds.jl|\footnote{where analogously a manifold of type~\texttt{MetricManifold} is defined.}.
 All functions unrelated to the metric are forwarded from \verb|G2| to \verb|G|, while all functions
@@
 
-The special Euclidean group \(\mathcal G = \mathrm{SE}(n)\), there are three different representations avalable:
+The special Euclidean group \(\mathcal G = \mathrm{SE}(n)\), there are three different representations available:
 \\

### 2025-11-01 paper/paper.tex
Unify inline code to never use lstinline since it is very broken within jlcode
cf. https://github.com/wg030/jlcode/issues/57

@@
 
-The main type for Lie Groups is the \verb|LieGroup{|$\mathbb {F}$\verb|, O, M} <: AbstractLieGroup|, which contains a manifold \verb|M <: AbstractManifold{|$\mathbb F$\verb|}| as well as the group operation \verb| O <: AbstractGroupOperation|, where $\mathbb {F}$ is the set of scalars used in the representation of the manifold, usually \(\mathbb {F} = \mathbb{R}\) and in some cases \(\mathbb {F}\) is the field of complex numbers \(\mathbb{C}\) or the ring of quaternions \(\mathbb H\).
+The main type for Lie Groups is the \verb|LieGroup{|$\mathbb {F}$\verb|, O, M} <: AbstractLieGroup{|$\mathbb {F}$\verb|, O, M}|, which contains a manifold \verb|M <: AbstractManifold{|$\mathbb F$\verb|}| as well as the group operation \verb| O <: AbstractGroupOperation|, where $\mathbb {F}$ is the set of scalars used in the representation of the manifold, usually \(\mathbb {F} = \mathbb{R}\) and in some cases \(\mathbb {F}\) is the field of complex numbers \(\mathbb{C}\) or the ring of quaternions \(\mathbb H\).
 
@@
 
-Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE, where \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
-The function \lstinline{force_and_torque} calculates the force and torque affecting the rod, while \lstinline{dynamics} converts them to the Lie algebra element can be used by the solver.
+Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE, where \verb|E_field| is a function that calculates electric field strength from the point charges at the given point.
+The function \verb|force_and_torque| calculates the force and torque affecting the rod, while \verb|dynamics| converts them to the Lie algebra element can be used by the solver.
 \begin{lstlisting}[

### 2025-11-05 paper/paper.tex
Fixing a few minor LaTeX quirks and typos in the JuliaCon paper.

@@
+\documentclass{juliacon}
 % JuliaCon proceedings template
-\documentclass{juliacon}
 \setcounter{page}{1}
@@
 
-These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
+These are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
 They have applications in physics, robotics, stochastic processes, information geometry, and many other areas see~\cite{Chirikjian:2009,Chirikjian:2012},
 but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
-and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998,IserlesMunthe-KaasNoersettZanna:2000}.
+and their numerical aspects, for example when solving differential equations on Lie groups~\cite{IserlesMunthe-KaasNoersettZanna:2000, Munthe-Kaas:1998}.
 
@@
 
-This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}, whose logo is shown on the right.
+This paper provides an overview of the main features of \verb|LieGroups.jl|, whose logo is shown on the right, as of version \LieGroupsVersion{}.
 %
@@
     \centering
-    \includegraphics[width=0.12\textwidth]{logo.png}
+    \includegraphics[width=0.15\textwidth]{logo.png}
     \vspace{.5\baselineskip}
@@
 
-\subsection{Tangent spaces and the Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
+\subsection{The Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} $\frak g$ of the Lie group \(\mathcal{G}\).
 The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
 For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
-Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
-\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields.
+Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} defined by $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$%
+\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(e)$ to define right-invariant vector fields.
 \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
+.
 Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
@@
 
-An important tool to “move around” on the Lie group is the exponential map.
-The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
+An important tool to “move around” on the Lie group is the exponential.
+The (Lie group) exponential (function) \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined~\cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
 \begin{equation*}
@@
 
-For \((\mathbb{R}^n, +)\) the exponential map is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
-Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential map is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).
+For \((\mathbb{R}^n, +)\) the exponential is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
+Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential function is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).
+
+From the interpretation of the exponential function to follow a curve starting at the identity, we can define the (Lie group) \term{exponential map} to “start from an arbitrary point” \(g \in \mathcal{G}\) by \(\exp_g\colon \mathfrak{g} \to \mathcal{G}\), \(\exp_g(X) \coloneqq g \cdot \exp(X)\) for all \(X \in \mathfrak{g}\).
 
@@
         \bottomrule
-    \end{tabular}
+    \end{tabular}%
     \label{tab:GroupOperations}
@@
 
-In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G,X)|%
+In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G, X)|%
 %\footnote{in-place variant \texttt{exp!(G, g, X)}.}
 .
-If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\mathfrak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
+If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the exponential map $\exp_g\colon\mathfrak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
 In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
@@
 .
-As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map. The Riemannian exponential map \verb|exp(M, p, X)|
+Note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map.
+This usually differs from the Lie group exponential map.
+The Riemannian exponential map \verb|exp(M, p, X)|
 %\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.}
@@
 Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
-This allows to use either plain arrays or own structures to represent these in code.
-\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}
+This allows to use either plain arrays or own structures to represent these in code%
+\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}%
+.
 Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.
@@
 
-\subsubsection*{Topological basics.}
+\subsubsection*{Topological basics}
 Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\mathfrak g$ \verb|= LieAlgebra(G)|.
@@
         \bottomrule
-    \end{tabular}
+    \end{tabular}%
     \label{tab:GroupActionTypes}
@@
         \bottomrule
-    \end{tabular}
+    \end{tabular}%
     \label{tab:available_lie_groups}
@@
 Since $\times$ is a binary operator in Julia, given two Lie Groups \verb|G,H|, their product Lie group can be constructed via \verb|G|\,$\times$\,\verb|H|.
-The same applies for more than two groups, i.e. you can construct arbitrary long product Lie groups \verb|G1|\,$\times$\,\verb|G2|\,$\times$\,\verb|G3|\,$\times$\,\ldots.
+The same applies for more than two groups, i.e.\ you can construct arbitrary long product Lie groups \verb|G1|\,$\times$\,\verb|G2|\,$\times$\,\verb|G3|\,$\times$\,\ldots.
 
@@
 Note that inheriting from the power manifold, there are two different data types to represent data, per default a single large array.
-Alternatively, one can use the constructor \verb|PowerLieGroup(G, NestedPowerRepresentation(), n)| to represent points and tangent vectors as a vector of points.
+Alternatively, one can use the constructor \verb|PowerLieGroup(G, NestedPowerRepresentation(), n)| to use the other representation of points and tangent vectors, namely as a vector of points/tangent vectors.
 
@@
 For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([0, 2\pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
-These are constructed by specifying the representation of points, i.e.
-\verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively.
+These are constructed by specifying the representation of points, i.e.\ \verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively.
 Note that each of these representations uses a different underlying manifold, namely \verb|Circle(|$\mathbb R$\verb|)|, \verb|Circle(|$\mathbb C$\verb|)|, and \verb|Sphere(1)|, respectively.
@@
     \caption{Bézier curves on the Lie groups \(\mathrm{SE}(2)\) (teal) and \(\mathrm{SO}(2) \times \mathbb{R}^2\) (sand) based on the same control points
-    $g_1,\ldots,g_4$ (blue).}
+    $g_1,\ldots,g_4$ (blue).}%
     \label{fig:Bezier}
@@
 where \(g \in \mathcal G_1\) and \(f(g, \dot{g}, t)\) is derived from the above equations. We further require initial conditions for both \(g\) and \(\dot{g}\) at time \(t_0=0\).
-Then this can be solved using Lie group methods, such as the Runge-Kutta-Munthe-Kaas (RKMK) variant \cite{Munthe-Kaas:1998} of Heun's method~\cite[App.~A.1]{IserlesMunthe-KaasNoersettZanna:2000}.
+Then this can be solved using Lie group methods, such as the Runge-Kutta-Munthe-Kaas (RKMK) variant~\cite{Munthe-Kaas:1998} of Heun's method~\cite[App.~A.1]{IserlesMunthe-KaasNoersettZanna:2000}.
 
@@
     different end markers to indicate their orientation.
-    }
+    }%
     \label{fig:Heun}

### 2025-11-13 paper/header.tex
minor improvements

@@
 
-\keywords{Riemannian manifolds, Lie groups, differential geometry, numerical analysis, scientific computing}
+\keywords{Lie groups, differential geometry, Riemannian manifolds, numerical analysis, scientific computing}
 
@@
 pdfauthor = {Ronny Bergmann, Mateusz Baran},
-pdfkeywords = {Riemannian manifolds, Lie groups, differential geometry, numerical analysis, scientific computing},
+pdfkeywords = {Lie groups, differential geometry, Riemannian manifolds, numerical analysis, scientific computing},
 }

### 2025-11-13 paper/paper.tex
minor improvements

@@
     Julia package that provides an interface to define Lie groups as well as the corresponding Lie algebra and Lie group actions.
-The package also offers a well-documented, performant, and well-tested library of existing Lie groups, their algebra and corresponding group actions.
+The package also offers a well-documented, performant, and well-tested library of such objects, with a focus on numerical computations in engineering.
 
@@
 We denote a Lie group by \(\mathcal{G} = (\mathcal{M}, \cdot)\) where \(\mathcal{M}\) is a smooth manifold and \(\cdot\) is the group operation.
-A smooth manifold \(\mathcal{M}\) is a topological space that is locally isomorphic to an Euclidean space \(\mathbb{R}^n\) for some \(n \in \mathbb{N}\), but globally may have a different topology.
+A smooth manifold \(\mathcal{M}\) is a topological space that is locally homeomorphic to an Euclidean space \(\mathbb{R}^n\) for some \(n \in \mathbb{N}\), but globally may have a different topology.
 We call \(n\) the dimension of the manifold \(\mathcal{M}\), denoted by \(\dim(\mathcal{M}) = n\).
@@
 \\
-A group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) has to satisfy the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
+A group operation \(\cdot\colon\mathcal{G} \times \mathcal{G} \to \mathcal{G}\) is a function that satisfies the group axioms: associativity, existence of an identity element \(e \in \mathcal{G}\), and existence of inverses \(g^{-1} \in \mathcal{G}\) for all \(g \in \mathcal{G}\).
 Furthermore the group operation \(\cdot\) (on \( \mathcal{G}\times\mathcal{G} \)) and the inversion map \(\iota\colon\mathcal{G} \to \mathcal{G}, g \mapsto g^{-1}\) have to be smooth maps.
@@
 % chktex 9ok
-The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose $p^{\mathrm{T}}$ (or an angle $-\alpha$).
-% -(-\pi) would be out of the range
+The identity element is given by the identity matrix \(I\) (or the angle $\alpha=0$) and the inverse of a rotation matrix is given by its transpose $p^{\mathrm{T}}$ (or an angle $-\alpha$ when $\alpha \neq -\pi$ and $-\pi$ otherwise).
+%
 
@@
 We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
-Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
+Here, a tangent vector $X$ induces the constant vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
 As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
-The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
+The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ that consists of all skew-symmetric $n \times n$ matrices.
 For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$.
@@
 For example \verb|identity_element(G)| returns the identity element $e$ of a Lie group $G$ in a default representation, \verb|identity_element!(G, e)| writes the result into the pre-allocated variable \verb|e|, which can also be used with other representation types.
-To generate a point in a specific representation of type \verb|T|, one can use \verb|identity_element(G, T)|.
+To generate a point in a specific representation by objects of type \verb|T|, one can use \verb|identity_element(G, T)|.
+For example, a points on the special Euclidean group $\mathrm{SE}(n)$ can be represented either as a $(n+1) \times (n+1)$ matrix or a \verb|ArrayPartition| with rotation and translation stored separately.
 
@@
 
-Alternatively to points, also different manifolds can be used.
+Alternatively to points, the same group structure can be defined over different topological manifolds.
 This is for example the case for the circle group \(\mathbb{S}^1\), which can be represented as angles, points on the unit circle in \(\mathbb{R}^2\) or as complex numbers with unit norm, all three of which are different manifolds to be used internally.
@@
 \verb|manifold_dimension(G)| to get the dimension of the manifold,
-\verb|project(G, q)|, %
-%\footnote{in-place variant \texttt{project!(G, p, q)}.} to project a point onto the manifold, as well as
+\verb|project(G, q)|, to project a point onto the manifold, as well as %
+%\footnote{in-place variant \texttt{project!(G, p, q)}.} 
 \verb|rand(G)| and \verb|rand(rng, G)| %
@@
 If the actual value is needed, one can call
-
 \verb|identity_element(G)| or \verb|identity_element(G, typeof(g))|, where \verb|g| is an element of the group.
@@
 which computes \(g^{-1}\) for a point \(g \in \mathcal{G}\).
-
 Another function is $c_g(h) = g \cdot h \cdot g^{-1}$ called \term{conjugation}, which is available as \verb|conjugate(G, g, h)|%
@@
 
-Finally, there is a specific function for the differential of the conjugate at the identity $h=e$ called \verb|adjoint(G, g, X)|
+Finally, there is a specific function for the differential of the conjugate at the identity $h=e$, called \verb|adjoint(G, g, X)|,
 %\footnote{in-place variant \texttt{adjoint!(G, Y, g, X)}.}
@@
 .
-Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.
+Keep in mind, that the Lie algebra is a vector space, so that addition, subtraction and scalar multiplication, as well as their broadcasted variants, are assumed to be defined in case you use an individual data type.
 
@@
 
-Given a vector $c \in \mathbb R^n$, we obtain the tangent vector by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b)| where \verb|b| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
-When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\mathfrak g$\verb|, c, b, T)| specifying the tangent vector type%
-%\footnote{the in-place variant \texttt{get\_vector!($\mathfrak g$, X::T, c, b)} can infer the type \texttt{T} of \texttt{X} directly.}
+Given a vector $c \in \mathbb R^{\dim(\mathfrak g)}$, we obtain the corresponding tangent vector by calling \verb|get_vector(|$\mathfrak g$\verb|, c, B)| where \verb|B| is a basis of the tangent space, i.e.\ a subtype of \verb|AbstractBasis| and defaults within \verb|LieGroups.jl| to \verb|DefaultLieAlgebraOrthogonalBasis()|.
+When a Lie group has different representations of points and tangent vectors, these are distinguished by calling \verb|get_vector(|$\mathfrak g$\verb|, c, B, T)| specifying the tangent vector type%
+%\footnote{the in-place variant \texttt{get\_vector!($\mathfrak g$, X::T, c, B)} can infer the type \texttt{T} of \texttt{X} directly.}
 .
 Given a tangent vector $X \in \mathfrak g$, we obtain its coordinates by calling \verb|get_coordinates(|$\mathfrak g$\verb|, X, B)|%
-%\footnote{in-place variant \texttt{get\_coordinates!($\mathfrak g$, c, X, b)}}
-, where the Basis is again optional.
+%\footnote{in-place variant \texttt{get\_coordinates!($\mathfrak g$, c, X, B)}}
+, where the basis is again optional.
 
@@
 \subsubsection*{Push forward and pull back of tangent vectors}
-To the representation of a tangent vector $X \in \mathfrak g$ to be represented in the tangent space at $g \in \mathcal G$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$.
+To get a tangent vector $X_g \in T_g \mathcal G$ at point $g \in \mathcal G$ from its Lie algebra representation $X \in \mathfrak g$ we have to use the push forward of the left multiplication with $g$, i.e.\ $D\lambda_g(e)[X]$.
 This is implemented as \verb|push_forward_tangent(G, g, X)|%
@@
 \subsubsection*{Jacobians}
-On Euclidean space, the terms differential and Jacobian are often used interchangeably.
+On a Euclidean space, the terms differential and Jacobian are often used interchangeably.
 Within differential geometry, the differential (or push forward) of a smooth map $f\colon \mathcal M \to \mathcal N$ between two manifolds is a map between the corresponding tangent spaces.
@@
 The differential is a linear map between the tangent spaces.
-However, in the coordinate-free representation of tangent vectors, there is no matrix representation of this linear map.
+However, in the coordinate-free representation of tangent vectors, no matrix representation of this linear map is constructed.
 But as soon as we choose bases for the tangent spaces and we represent tangent vectors in coordinates (cf.~\verb|get_coordinates| previously), we can represent the differential as a matrix, called the Jacobian matrix.
@@
 The signatures are
-\verb|jacobian_exp(G, p, X, b)|
-%\footnote{in-place variant \texttt{jacobian\_exp!($\mathfrak g$, J, p, X, b)}}
-and \verb|jacobian_conjugate(G, g, X, b)|%
-%\footnote{in-place variant \texttt{jacobian\_conjugate!($\mathfrak g$, J, g, X, b)}}
-, where \verb|b| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.
+\verb|jacobian_exp(G, p, X, B)|
+%\footnote{in-place variant \texttt{jacobian\_exp!($\mathfrak g$, J, p, X, B)}}
+and \verb|jacobian_conjugate(G, g, X, B)|%
+%\footnote{in-place variant \texttt{jacobian\_conjugate!($\mathfrak g$, J, g, X, B)}}
+, where \verb|B| is again the basis to represent the tangent vectors in coordinates and defaults to \verb|DefaultLieAlgebraOrthogonalBasis()|.
 
@@
 
-the main functions then are \verb|apply(a::GroupAction, g, p)|
+The main functions then are \verb|apply(a::GroupAction, g, p)|
 %\footnote{in-place variant \texttt{apply!(a::GroupAction, q, g, p)}.}
@@
 %\footnote{in-place variant \texttt{diff\_apply!(a::GroupAction, Y, g, p, X)}.}
-to compute \(D\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|
+to compute \(D_{\mathcal M}\sigma(g, p)[X]\) for a tangent vector \(X \in T_p\mathcal M\) with respect to the manifold argument as well as \verb|diff_group_apply(a::GroupAction, g, p, Y)|
 %\footnote{in-place variant \texttt{diff\_group\_apply!(a::GroupAction, Z, g, p, Y)}.}
-to compute \(D\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \mathfrak g\).
+to compute \(D_{\mathcal G}\sigma(g, p)[Y]\) for a Lie algebra element \(Y \in \mathfrak g\).
 
@@
 Similarly for a right group action the inverse is a left group action.
-This is tanken into account when using \verb|apply|.
+This is taken into account when using \verb|apply|.
 
@@
         \verb|SymplecticGroup(n)| & \(\mathrm{Sp}(2n)\) & \\
-        \verb|TranslationGroup(n; field=|$\mathbb{F}$\verb|)| & \(\mathbb{F}^n\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}, \mathbb{H}\}$\\
-        \verb|UnitaryGroup(n)| & \(\mathrm{U}(n)\) & \\
+        \verb|TranslationGroup(n; field=|$\mathbb{F}$\verb|)| & \((\mathbb{F}^n, +)\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}, \mathbb{H}\}$\\
+        \verb|UnitaryGroup(n, |$\mathbb{F}$\verb|)| & \(\mathrm{U}(n)\) & $\mathbb{F} \in \{\mathbb{C}, \mathbb{H}\}$ \\
         \verb|ValidationLieGroup(G)| & \multicolumn{2}{l}{wraps \texttt{G} for numerical verification}\\
@@
 
-If you want to construct nested Products of Lie groups, e.g.\ \((\mathcal G_1 \times \mathcal G_2) \times \mathcal G_3\), the binary operator would not work, since it automatically flattens the input; use \verb|ProductLieGroup(ProductLieGroup(G1, G2), G3)| can be used instead.
+The binary operator $\times$ automatically flattens the input, so that the points are represented by non-nested \verb|ArrayPartition| objects.
+If you want to construct products of Lie groups where points are represented by nested \verb|ArrayPartition| objects, for example \((\mathcal G_1 \times \mathcal G_2) \times \mathcal G_3\), use \verb|ProductLieGroup(ProductLieGroup(G1, G2), G3)| instead.
 
 \subsubsection*{Power Lie groups}
-Though mathematically the case of a product Lie Group, where $G=H$ is also (just) a product, computationally we can benefit from the special structure.
-Therefore the \verb|PowerManifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(G\) with itself.
-As a shortcut, you can use \verb|G^n| to construct the \(n\)-fold Cartesian product of the Lie group \(G\) with itself.
+Mathematically power groups are product Lie groups where all manifolds are the same. Computationally we can sometimes benefit by exploiting this special structure.
+Therefore the \verb|PowerManifold(G, n1, n2, ...)| is provided to construct the \(n_1 \times n_2 \times \ldots\)-fold Cartesian product of the Lie group \(\mathcal G\) with itself.
+As a shortcut, you can use \verb|G^n| to construct the \(n\)-fold Cartesian product of the Lie group \(\mathcal G\) with itself.
 
@@
 %
-Similarly, the \term{right semidirect product Lie Group} $\mathcal{H} \rtimes \mathcal{G}$ is defined analogously with the positions of the groups swapped in the group operation and the default for the group action is given by \verb|default_right_action(G, H)|.
+Similarly, the \term{right semidirect product Lie Group} $\mathcal{H} \rtimes \mathcal{G}$ is defined analogously with the positions of the groups are swapped in the group operation and the default for the group action is given by \verb|default_right_action(G, H)|.
 This can be constructed via \verb|H|\,$\rtimes$\,\verb|G| then as well.
@@
 The decorator pattern is used within the \verb|JuliaManifolds| ecosystem to add or modify existing functionality.
-For Lie groups there are currently two such decorators available
+For Lie groups there are currently two such decorators available: one for adding a custom metric structure (inner product) to the group and one for numerical validation.
 
@@
 Most prominently,
-all input and output of a function is checked for validity, e.g.\ by calling \verb|is_point| and \verb|is_vector| on all points and tangent vectors, respectively.
-Similarly the inner manifold is wrapped into the similar concept from \verb|ManifoldsBase.jl|, the \verb|ValidationManifold|.
-While by default these checks do result in errors, this can be changed setting the \verb|error=| keyword to either \verb|:warn| or \verb|:info|.
+all input and output of group functions are checked for validity, e.g.\ by calling \verb|is_point| and \verb|is_vector| on all points and tangent vectors, respectively.
+Similarly, the inner manifold is wrapped into the similar decorator from \verb|ManifoldsBase.jl|, the \verb|ValidationManifold|.
+While by default failures in these checks result in errors, this can be changed by setting the \verb|error=| keyword of the \verb|ValidationLieGroup| constructor to either \verb|:warn| or \verb|:info|.
 
@@
 
-For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([0, 2\pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
+For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([-\pi, \pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
 These are constructed by specifying the representation of points, i.e.\ \verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively.
@@
 These two are available when \verb|RecursiveArrayTools.jl| is loaded and specifying the \verb|variant=:left| (default) or \verb|variant=:right|, respectively.
-The first, matrix representation work on either \verb|variant|.
+The matrix representation works on either \verb|variant|.
 
@@
             ymin = -3.25, ymax = 2.75,
+            align=center,
             title style={yshift=-.5\baselineskip,},
-            title={Trajectories of a charged rod between two point charges},
+            title={Trajectories of a charged rod in the electric field\\ generated by two point charges},
             xlabel={x [m]},
